### 2013-06-19 paper.tex
Initial commit of MACV SMD paper.

\documentclass[12pt]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{graphicx}
\usepackage{natbib}
%correct punctuation for MBE
\bibpunct[,]{(}{)}{;}{a}{}{,}

\usepackage{amsmath}

\renewcommand{\bottomfraction}{.9}
\renewcommand{\topfraction}{.9}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{.9}

\linespread{1.5}
\begin{document}

\title{\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}}
\author{Austin G. Meyer$^{1,2,4*}$, Sara L. Sawyer$^{3}$, Andrew D. Ellington$^{2}$,\\and Claus O. Wilke$^{1}$}
\date{}

\maketitle
\noindent
Address:\\
$^1$Section of Integrative Biology, Institute for Cellular and Molecular
Biology, and Center for Computational Biology and Bioinformatics. The University
of Texas at Austin, Austin, TX 78712, USA.\\
$^2$Department of Chemistry and Biochemistry, Institute for Cellular and Molecular 
Biology, The University of Texas at Austin, Austin, TX 78712, USA.\\
$^3$Section of Molecular Genetics and Microbiology, Institute for Cellular and Molecular 
Biology, The University of Texas at Austin, Austin, TX 78712, USA.\\
$^4$School of Medicine, Texas Tech University Health Sciences Center, 
Lubbock, TX 79430, USA.\\

\bigskip
\noindent
$^*$Corresponding author\\
$\phantom{^*}$Email: austin.meyer@utexas.edu\\
$\phantom{^*}$Phone: +1 512 971 0123\\

\bigskip
\noindent
Manuscript type: research article

\bigskip
\noindent Keywords: protein--protein interaction, molecular dynamics, arenavirus, machupo

\newpage
\begin{abstract}
Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases. In particular, the effects of multiple mutations, non-alanine substitutions, and flexible loops are difficult to predict with available tools and protocols. We present here a new method to interrogate affinity differences resulting from mutations in a host-virus protein--protein interface. Our method is based on extensive non-equilibrium all-atom simulations: We computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1) and estimate affinity using the maximum applied force during a pulling simulation and the area under the force-versus-distance curve. We find that these quantities provide novel biophysical insight into the GP1/hTfR1 interaction. First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes. Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that only one of them is critical for the binding interaction. Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-resistant hTfR1 mutants. Finally, our method provides an elegant framework to compare the effects of multiple mutations, individually and jointly, on protein--protein interactions.
\end{abstract}

\section*{Introduction}
The computational prediction of mutational effects on protein--protein interactions remains a challenging problem. Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure. For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}. However, such methods calculate the relative contribution to overall energy for each amino acid in a protein. They do not probe the interface directly, but rather calculate the total energy with each substitution \citep{Grant2011,Kortemme2004}. Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}. Unfortunately, such machine-learning approaches often suffer in novel applications, for which available training sets are small or non-existent. As such, these methods are poorly suited for most host-virus protein--protein systems. By contrast, first principles methods offer the chance to forgo training. However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model that phases directly from one three dimensional protein model to another with no intermediate transition states \citep{Gilson1997,Lu2004}. Thus, they can quickly lose physiological relevance as the two proteins become more dissimilar and as a result are largely limited to testing only sequential point mutants.

Here we propose a new method to probe protein--protein interactions directly, by pulling two proteins apart that are initially bound in a complex. We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place. Then, we measure the force required for dissociation. As a proxy for relative binding affinity, we use the maximum applied force (MAF) required for separation and the area under the force-versus-distance curve (AUC).

We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}. Machupo is a negative-sense RNA virus of the arenavirus family \citep{Charrel2003}. Worldwide arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}. Members of the arenavirus family include the lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the guanarito, junin and machupo viruses endemic to rodents in South America \citep{Charrel2003}. The South American arenaviruses typically infect humans after rodent contamination and cause a devastating hemorrhagic fever with very high mortality \citep{Charrel2003}.

The hTfR1 is the primary receptor used by MACV for binding its host cell prior to infection. The hTfR1 protein contains three extracellular domains with the two more basilar domains serving most of the endogenous, transferrin-binding function \citep{Abraham2010,Rad20112}. Cellular entry is initiated by GP1 binding to the apical domain of hTfR1. Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of viral infectivity \citep{Rad20111,Rad20112}. The co-crystal structure shows that a 15 amino acid flexible loop on GP1 is responsible for viral binding \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}. Furthermore, the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain. Experimental alanine-scanning and whole cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.

We applied our method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes. At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity is often not as simple as the static structure suggests. For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1. According to our simulations, this network likely does not affect binding affinity. Our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in host-virus protein--protein interfaces.

\section*{Materials and Methods}

\subsection*{System Modeling}
We used the protein visualization software PyMOL \citep{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1. No residues were removed from the viral protein. We also removed 10 sugar residues from the native structure.

After system reduction, the Visual Molecular Dynamics (VMD) \citep{Humphrey1996} package along with its system of back-ends was used for all subsequent modeling. The Orient add-on package allowed us to rotate the system axis such that the direction of steering was oriented directly down the z-axis. De-glycosylation simplified the system such that Autopsf could easily find the chain terminations and patch them appropriately. The Solvate package was used to generate a TIP3P water model with a 5 \AA ngstrom buffer (relative to the maximum dimensions of the proteins) on all sides except down the positive z-axis where a 20 \AA ngstrom buffer was created. Finally, we used the Autoionize package to place 150 millimolar NaCl and neutralize the total system charge. In the end, each modeled system had approximately 28,000 atoms.

\subsection*{Equilibration}
NAMD was used for all simulations in this study \citep{Phillips2005}. In addition to the modeled system, for equilibration we generated a configuration file that fixed the $\alpha$-carbon backbone. This was accomplished by setting $\beta = 1$. Further, we generated a configuration file with fixed $\alpha$-carbon atoms at residues 41-92 (numbered linearly, in this case, starting at 1 for the first amino acid as was required for VMD) in the hTfR1. The second file was used to affix a harmonic restraint, thus preventing any unfolding due to system reduction. More importantly, the harmonic restraint allowed the protein complex to equilibrate while preventing any drift from its predefined orientation. Finally, we calculated the system center and dimensions for use in molecular dynamics settings.

We used the Charmm27 \citep{Brooks1983} all-atom force field. The initial system temperature was set to 310K. Several typical MD settings were used including switching and cutoff. In addition, we used a 2 femtosecond time step with rigid bonds (the exact configuration is provided in the supplement). We used periodic boundary conditions with the particle mesh ewald (PME) method of computing full system electrostatics outside of the explicit box. Furthermore, we used a group pressure cell, flexible box, and langevin piston during equilibration. Finally, a harmonic restraint (called harmonic constraint in VMD) was set as stated previously.

To start the simulation, the langevin piston was switched off and the system was minimized for 1000 steps. Next, the fixed backbone was released, and the system was minimized for an additional 1000 time steps. Subsequently, the system was released into all-atom molecular dynamics for 3000 steps. Finally, the langevin piston was turned on and the system was simulated for 2 ns (1,000,000 steps) of chemical time. For each mutant, twenty independent replicates were run with an identical protocol.

\subsection*{Steered Molecular Dynamics}
Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1. Also, the $\alpha$-carbon of all residues (163-318 in linear numbering) in the GP1 were given an occupancy of 1, indicating they had an applied force during the simulation.

In every way possible the subsequent steering simulation had the same parameters as the equilibration simulation. Thus, the steered molecular dynamics (SMD) \citep{Is2001B} simulation was a typical restart with slightly different settings. Periodic boundary conditions were incorporated as part of the system and PME was again used to approximate full system electrostatics. In addition, we first tuned the pulling velocity for practical time considerations. Subsequently, the spring constant was adjusted to ensure sufficient signal to noise with the given velocity. As a result, the spring constant was set to 5 (units), velocity set to 2 (units), and direction down the positive $z$-axis.

Finally, SMD was run for 15 ns (7,500,000 time steps) of chemical time. For each simulation, we randomly selected one of the equilibration runs for restart. We ran 50 replicate simulations per mutant. All GP1/hTfR1 complexes separated by greater than 4 \AA ngstroms and many separated to 10 or more.

To leave the final trajectory of a tractable size, only 1000 evenly spaced frames were retained from each simulation, leaving a final trajectory size of 323 MB. See the supplemental movie for a representative unbinding trajectory. Initial development of the SMD protocol was carried out on the Lonestar cluster at the Texas Advanced Computing Center (TACC). All production SMD simulations were performed on the Hrothgar cluster at Texas Tech University, using NAMD 2.9. Each simulation was parallelized over 60 computational cores and utilized approximately 20 hours of computing time. The total chemical time simulated for this project was nearly 10 $\mu$s, requiring slightly over 1 million cpu-hours.

\subsection*{Post-processing}
The python packages MDAnalysis \citep{Agrawal2011} and ProDy \citep{Bakan2011} were both used at various points in post-processing. The molecular trajectory (comprising the atomic coordinates per time) was parsed to compute the center-of-mass for each of the two complexes. The starting center-of-mass distance was set to zero and the distance was re-computed at each time step relative to the starting distance. The force is output directly to the standard log file.

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. To reduce the noise, we averaged each point with a 75 point interpolation window. The window ensured that the averaged MAF was never cut off near the beginning of the simulation due to lost time steps (caused by averaging). The two primary descriptive statistics we used were maximum interpolated applied force (MAF) and total area under the interpolated curve (AUC). We used the pairwise.t.test function with the FDR p-value correction to test for significance. The ggplot \citep{ggplot} package was used to generate most of the figures.

\section*{Results}

\subsection*{The GP1/hTfR1 system}

The GP1/hTfR1 interface marks a particularly important and useful test system. There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV. Virtually all of the important sites have been mapped by \textit{in vitro} flow-cytometry based infectivity assays. The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals). In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \citep{Grant2011,Kortemme2004} only marginally useful. The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.

For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \citep{Abraham2010}. The apical domain of hTfR1 interacts directly with GP1 while the other two domains are closer to the cell membrane and have essentially no interaction with GP1. The biophysical independence of the apical domain allowed us to isolate it without significantly affecting the GP1/hTfR1 interaction. Figure~\ref{fig:protein_comparison} shows a model of the initial structure and that of the pared structure. Although GP1 has several glycosylatable residues, we opted to use the de-glycosylated protein for this study. The complexity of correctly parameterizing diverse sugar moieties is outside of the scope of this paper. Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the effect of those contacts \textit{in vivo} is still unresolved \citep{Abraham2010}.

\subsection*{Steered molecular dynamics simulations}
To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions. First, we generated a physically realistic simulation world around the protein complex. The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions. Next, with a fixed $\alpha$-carbon backbone we carried out successive rounds of energy minimization, followed by heating, and backbone release. Last,  with minimal restraints a short round of all-atom molecular dynamics \citep{Cuendet2008} was performed. We used the final state from each equilibrated system to restart another MD simulation. For SMD, at the centroid of several atoms a force was applied to a single point in GP1. By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging. The latter approach was used for statistical analysis. In addition, to minimize the computational time in unproductive molecular dynamics, we shortened simulation time such that the interaction just barely dropped to zero near the end of each run. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. \textbf{Supplement 1 ($\leftarrow$ what is this? figure? video? text? table?)} illustrates the separation distance between peptide domains. The quantities MAF and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

Before systematically applying SMD to the GP1/hTfR1 interaction, we needed to ensure the method was sufficiently sensitive to distinguish between relatively minor point mutations. While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape. For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach. Several residues on both proteins appear to mediate putative hydrogen-bonding interactions. We selected three sites for alanine substitutions: Tyr211 and Asn348 in the hTfR1 and vArg111 in MACV. For notation purposes, the viral site is always referred to with a preceding `v'. According to the static complex \citep{Abraham2010}, Tyr211 and vArg111 are involved in one hydrogen-bonding network and Asn348 is involved in a second, distinct hydrogen-bonding network.

We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the MAF and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:pairwise_differences}), while the other two did not (Table~\ref{tab:pairwise_differences}). From these results, we concluded that our method can indeed successfully distinguish individual mutants from WT, and that alanine substitutions at Asn348 and vArg111 may have little effect on binding affinity (see below for further discussion). When considering additional mutants (also discussed below), we found that MAF was generally sufficient to distinguish mutants (Tables~\ref{tab:pairwise_differences} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}). In general, however, MAF seemed to be the more sensitive statistic than AUC.

\subsection*{Comparative analysis of the GP1/hTfR1 interface}

Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. We made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause reduced binding. Then, we compared the size of those differences to reduced binding single mutants (Figure~\ref{fig:sensitivity}).

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_statistics}). The largest single-mutant impact was seen for a mutation to alanine at site Tyr211 in the hTfR1 ($ p < 0.001 $).  Several mutations at this site have proven capable of causing loss-of-infection, according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 in the GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure \citep{Abraham2010}. In agreement with the static structure, our molecular dynamics simulations show Tyr211 is involved in a large hydrogen-bonding network (results not shown).

Although Tyr211Ala appears to have a large impact on binding affinity, the single mutant provided little underlying biochemical rationale. Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for loss of binding affinity. In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding. We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}. Both mutations proved incapable of causing reduced binding affinity in our simulations. In fact, they caused a significant increase in relative binding affinity (Figure~\ref{fig:sensitivity} and Table~\ref{tab:MAF_p}). This outcome is not surprising, if we consider that Tyr211 participates in a bonded network rather than independently. Combined with some biochemical intuition, these results suggest that mutations to other polar residues would likely not affect the dynamics of binding directly.

We also simulated several point mutations at Asn348 in the hTfR1. Asn348Lys is reported in the literature to cause reduced infectivity \textit{in vivo} \citep{Rad2008,Abraham2010}. Further, in the crystal structure two residues on the GP1 and at least three on hTfR1 are involved in a fairly extensive hydrogen-bonding network. As discussed above, the alanine mutation at this site showed no significant difference in MAF or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, the Asn348Lys mutation showed no significant difference from WT either, and neither did Asn348Trp. (For both of these mutations, mean MAF and mean AUC was lower than for WT, however. See Table~\ref{tab:summary_statistics})  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. These results suggest that, in contrast to Tyr211, the critical difference at position 348 in hTfR1 is size exclusion rather than the destruction of a hydrogen-bonding network. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:pairwise_differences}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to MAF (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

Last, we further analyzed a single mutation in the GP1, vArg111Ala. It has been reported that vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}. As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either MAF or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}). This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity. To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala. This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:pairwise_differences}). Thus, vArg111Ala in the GP1 appeared capable of suppressing the loss-of-binding caused by Tyr211Ala. This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction leads to comparable binding affinity.

\section*{Discussion}
We have applied a method utilizing steering forces in all-atom molecular dynamics simulations to evaluate the effects of mutations at the GP1/hTfR1 interface. We modeled mutations at several sites in the GP1/hTfR1 interface, and verified that our computational protocol was sensitive enough to distinguish point mutants in hTfR1. Further, we identified two test statistics, MAF and AUC, that can be used as proxies for binding affinity. We systematically tested several point mutations to understand their contribution to the binding interaction. In the case of Asn348Lys, we have shown that the static structure provides little insight into why this mutation causes loss-of-infectivity \textit{in vivo}. While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, loss of binding affinity at that site is actually caused by size and charge restriction. Further, we have shown that the cause of diminished infectivity in the viral single-mutant vArg111Ala may not be due to binding-energy differences. We also found that a negatively polar residue at site 211 in hTfR1 is absolutely critical for a tight binding interaction. Any non-polar mutation at Tyr211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection. However, our data suggest that the viral mutation vArg111Ala may restore the infective potential of MACV that was previously destroyed by a Tyr211Ala hTfR1 mutation.

By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application. Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Buch2011,Giorgino2012}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}. Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2012}. And still others aimed at more esoteric problems apply driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}. Here, we have shown that SMD can provide some insight into the \textit{relative} strength of protein--protein interactions. In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application. Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanism. Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation. Thus, SMD may open avenues for subsequent experimental work.

Our findings rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}. We found that some substitutions at positions 211 and 348 did negatively affect the strength of viral binding. In the two cases, the computational data suggests that the reason and nature of the effects are very different. Congruent with infectivity data, mutations at Tyr211 often have a dramatic effect on receptor binding \citep{Rad2008,Rad20111}. Tyr211Ala alone produces such a large decrease in affinity that it can be easily distinguished from the WT complex. Structural analysis reveals that the biochemical cause involves the destruction of a critical hydrogen-bonding network with Ser113 and vArg111 in GP1. Further, Tyr211Asp and Tyr211Thr were easily resolved from the WT complex, but they appear to cause an increase in affinity. As a result, we predict experimental structural analysis will show Tyr211Asp and Tyr211Thr mutations cause a relatively dramatic tertiary structure deviation. Given that our model starts from an associated complex, it is not surprising that Tyr211Asp would cause increased rather than decreased affinity. Aspartate is charged rather than polar at physiological pH and its shorter side chain means the proteins would be pulled in closer apposition. Nevertheless, our results show Tyr211, or some other amino acid that is capable of hydrogen bonding, is necessary for a stable hydrogen-bonding network leading to efficient binding. Moreover, the cause of decreased infectivity seen in rodent populations with the Tyr211Asp mutation is not a simple change in binding affinity.

Previous work has shown that the hTfR1 mutation Asn348Lys is capable of producing decreased infectivity \emph{in vitro} \citep{Rad2008,Rad20111}. Moreover, the molecular model based on x-ray diffraction data suggests destruction of a second hydrogen-bonding network as the cause \citep{Abraham2010}. In our simulations no individual point mutation at Asn348 showed significantly different affinity than the WT complex. Moreover, the mutation Asn348Ala resulted in an average MAF that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}). Thus, the putative hydrogen-bonding network involving Asn348 is probably a crystallization artifact. Even when paired with the highly destructive Tyr211Ala mutation, Asn348Ala was refractory to changes in binding affinity. By contrast, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Moreover, although Asn348Trp and Asn348Lys individually failed to be distinguishable from WT hTfR1, they both caused a mean reduction in MAF and produced larger than expected synergistic effects. This data suggests that site 348 is size and charge restricted. Any mutation to a very large or positively charged side chain will likely produce decreased infectivity.

The GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations. We suspected that this result was primarily caused by a lack of sensitivity of our method, since vArg111's binding orientation near Tyr211 and involvement in the critical hydrogen-bonding network strongly suggested that mutations at that site should have an effect on binding affinity. As an alternative way to test this hypothesis, we evaluated whether we could introduce a suppressor mutant at site vArg111. We introduced vArg111Ala jointly with Tyr211Ala, thus effectively replacing a hydrogen-bonding network with a hydrophobic interaction. Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean MAF to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone. As mutations at hTfR1 site 211 are known to reduce infectivity \emph{in vitro} and \emph{in vivo}, it is possible that vArg111 in GP1 can serve as a supressor site for compensatory MACV mutations. Thus, this site may be critically important for surveillance in the rodent population.

Although our computational predictions largely agreed with the available literature, some mutations that are reported to cause reduced infectivity in experimental systems (most notably vArg111Ala and Asn348Lys, discussed above) were indistinguishable from the WT complex in our simulations. We would like to emphasize here that we cannot expect perfect agreement. First, while we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method. It is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method. More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph). Second, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays. There are several reasons other than binding due to which a viral mutation may cause a phenotypic difference in infectivity. \textbf{Shorten the following (till end of paragraph), make less critical to experimentalists. Just list the things other than binding that may matter.} First, across disconnected experiments there could be a gross difference in the GP1 concentration. Although western blots are a popular means of protein concentration normalization, when used in a quantitative manner they are notoriously difficult and unreliable. In addition, though western blots could show gross differences in protein concentration, viral binding may be very sensitive to small differences in fusion protein concentration.  Second, even if we trust protein expression normalization via western blot, it is not clear how much of the protein is grossly or partially misfolded. An additional analytical step with circular dichroism or an analogous technique may further clarify large-scale folding differences. Third, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method.

As with any structural approach, we are constrained to start our computational model system from a static structure. One unfortunate reality in structural biology is poor resolution/discrimination at particularly long and flexible amino acid side chains. In addition, the number of acceptable side-chain rotamers increases in large amino acids. As a consequence there could be many untested starting conformations. Although we believe we have minimized unsampled conformations here, such an issue could be overcome with a sufficiently large number of SMD starting models. Of course, increasing sampling in this case comes at significant computational cost. There are a few additional challenges for investigating host-virus interaction via molecular dynamics simulation. As with any atomistic simulation, there is going to be a fairly large noise to signal ratio. To reduce noise, one could further customize each simulation, e.g. by determining the optimal pulling speed. Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \citep{Shaw2012}. Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \citep{Shaw2011}. With improved resources, we could investigate thousands of individual permutations in the GP1/hTfR1 binding interface. Going forward, we believe that all-atom molecular dynamics simulations can become an increasingly viable approach for protein interaction studies.

Making accurate predictions regarding the effects of non-alanine mutations at protein--protein interfaces remains  challenging. The main obstacle is the enormous number of chemically allowed three dimensional configurations in proteins \citep{Gumbart2012}. In theory, one could carry out all-atom simulations from physical first principles \citep{Gumbart2012} for very long time periods. After equilibrium is reached in such simulations, one could then compute a binding constant by observing the fractional association/dissociation time among two proteins. Unfortunately, due to limited computational power and available methods, reaching equilibrium with a simulated protein system is truly a herculean task and is essentially impossible at this time \citep{Gumbart2012}. There are two broadly useful computational approaches that avoid first-principles equilibrium simulations. First, to reduce computational complexity several studies have used rigid body docking and machine learning \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}. While these approaches can work well in certain scenarios, they often underperform in applications to novel systems, due to their dependence on available training data. Second, hybrid approaches exist that exploit non-equilibrium calculations within the context of all atom simulations \citep{Dixit2001,Is2001A,Is2001B,Park2004,Grant2011,Kortemme2004,Gumbart2012}. However, to date most of these methods have not been applied to protein--protein interactions (but rather to simpler scenarios such as protein--ligand interactions). What we have shown here is that hybrid non-equilibrium approaches can provide a useful and practical tool in studying protein--protein interactions.

We have probed the biochemical effects of mutations at the host-virus interface between MACV and its human receptor. We applied an approach that imparts forces in an all-atom molecular dynamics (MD) simulation to drive the system through equilibrium. Our work has shown that with relatively modest computing resources it is possible to use MD simulations to investigate point mutations in the GP1/hTfR1 interface. In most cases, a double mutant was easily sufficient to resolve a difference in relative affinity. Moreover, relative to the WT complex one point mutant was capable of producing a large decrease in binding affinity. We applied our technique to the WT complex and 10 mutations in the GP1/hTfR1 interface. We were able to rationalize the results of infectivity data with a fundamental physical simulations. Our data has provided some direction for further biochemical intuition in this system. A polar residue is absolutely critical at position 211 in hTfR1. Any mutation to a nonpolar residue will probably lead to a rapid viral compensatory response. Also, although rodent populations may produce further mutations at Asn348 in hTfR1, unless a mutation is positively charged or very large it is unlikely to result in a large disruption to viral binding. We have also provided both a site (vArg111) and a mutation (vArg111Ala) that may be critical for future host-range adaptation of MACV, and thus should be monitored in surveillance efforts. Finally, our analysis offers a new means to systemically investigate host-virus interactions through computational molecular dynamics.

\section*{Acknowledgments}
This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu). The authors acknowledge Bryan Sutton for opening access to the Hrothgar cluster.

This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) to A.D.E., and the National Institutes of Health (R01-GM088344) to C.O.W.

\bibliographystyle{MBE}
\bibliography{paper}

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/protein_comparison.pdf}}
\caption{\label{fig:protein_comparison}\title{Comparison of GP1/hTfR1 complex.} Figure A shows the full, de-glycosylated GP1/hTfR1 co-crystal structure. For figure B, we show the reduced structure for SMD. GP1 is blue and hTfR1 is green.}
\end{figure}

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/force_curve.pdf}}
\caption{\label{fig:force_curve}\title{Force versus distance curve of Y211A mutant and WT.} The average force curve for 50 replicates of the WT complex was black, and the average of 50 replicates of the Y211A mutant was red. Notice there is a very large difference in MAF and AUC between the two complexes.}
\end{figure}

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/sensitivity_histogram.pdf}}
\caption{\label{fig:sensitivity}\title{Density plot of three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left. The large non-overlapping areas indicate a large and statistically significant difference in these three complexes with relatively minor differences.}
\end{figure}

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_statistics}Summary statistics for each mutation tested. $\mu_\text{MAF}$ is the mean and $\sigma_\text{MAF}$ is the standard deviation of MAF over all simulations. $\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations.}
\begin{center}
  \resizebox{11cm}{!} {
    \begin{tabular}{l r r r r}
    \hline
      Mutation &      $\mu_\text{MAF}$ & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ \\ \hline
            WT &                 772.00 & 154.13               & 198794.7    & 66765.64 \\
         N348A &                 799.33 & 154.36               & 202247.3    & 65175.78 \\
         N348K &                 731.36 & 151.01               & 179087.5    & 55799.49 \\
         N348W &                 732.13 & 131.21               & 192043.2    & 66916.36 \\
        vR111A &                 754.39 & 103.25               & 185896.8    & 46408.74 \\
         Y211A &                 684.03 & 126.69               & 153517.7    & 40889.12 \\ 
         Y211D &                 875.44 & 116.46               & 234909.3    & 50494.11 \\ 
         Y211T &                 861.17 & 158.48               & 231078.7    & 97569.29 \\
   N348A/Y211A &                 762.96 & 144.81               & 173059.2    & 59898.10 \\
   N348W/Y211A &                 617.21 & 163.33               & 147705.6    & 48548.61 \\
  vR111A/Y211A &                 780.71 & 135.65               & 189919.1    & 48868.16 \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\clearpage
\begin {table}[H]
\caption{\label{tab:pairwise_differences}Pairwise differences (row variable minus column variable) in mean MAF. Bolded values are statistically significant at $p<0.05$. Values marked with * and ** are significant at $p<0.01$ and $p<0.001$, respectively.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT                  & N348A              & N348W              & N348K              & Y211A             & Y211D              & Y211T              & vR111A             & N348A/Y211A        & N348W/Y211A        \\ \hline
    N348A       &  27.33              &                    &                    &                    &                   &                    &                    &                    &                    &                    \\
    N348W       & -39.87              & \textbf{-67.20}    &                    &                    &                   &                    &                    &                    &                    &                    \\
    N348K       & -40.64              & \textbf{-67.97}    & -0.77              &                    &                   &                    &                    &                    &                    &                    \\
    Y211A       & \textbf{-87.97*}    & \textbf{-115.30**} & -48.10             & -47.33             &                   &                    &                    &                    &                    &                    \\
    Y211D       & \textbf{103.44**}   & \textbf{76.12*}    & \textbf{143.31**}  & \textbf{144.08**}  & \textbf{191.42**} &                    &                    &                    &                    &                    \\
    Y211T       & \textbf{89.17*}     & \textbf{61.85}     & \textbf{129.04**}  & \textbf{129.81**}  & \textbf{177.15**} & -14.27             &                    &                    &                    &                    \\
   vR111A       & -17.76              & -44.94             & 22.26              & 23.03              & \textbf{70.36}    & \textbf{-121.05**} & \textbf{-106.78**} &                    &                    &                    \\
    N348A/Y211A & -9.04               & -36.37             & 30.83              & 31.60              & \textbf{78.93}    & \textbf{-112.48**} & \textbf{-98.21*}   & 8.57               &                    &                    \\
    N348W/Y211A & \textbf{-154.79**}  & \textbf{-182.12**} & \textbf{-114.92**} & \textbf{-114.15**} & \textbf{-66.2}    & \textbf{-258.23**} & \textbf{-243.96**} & \textbf{-137.18**} & \textbf{-145.75**} &                    \\
   vR111A/Y211A &  8.71               & -18.12             & 48.58              & 49.35              & \textbf{96.68*}   & \textbf{-94.74*}   & \textbf{-80.46*}   & 26.31              & 0.59               & \textbf{163.50**} \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\iffalse
\begin {table}[H]
\caption{Pairwise difference $p-$values for interpolated AUC. Bolded values are statistically significant at $p<0.05$.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT               & N348A                   & N348W           & N348K                   & Y211A                   & Y211D                   & Y211T                   & vR111A          & N348A/Y211A & N348W/Y211A \\ \hline
    N348A       & 0.77             &                         &                 &                         &                         &                         &                         &                 &             &        \\
    N348W       & 0.67             & 0.42                    &                 &                         &                         &                         &                         &                 &             &        \\
    N348K       & 0.17             & 0.056                   & 0.38            &                         &                         &                         &                         &                 &             &        \\
    Y211A       & \textbf{0.00094} & \textbf{3.5x10$^{-05}$} & \textbf{0.0045} & 0.063                   &                         &                         &                         &                 &             &        \\
    Y211D       & \textbf{0.0072}  & \textbf{0.0051}         & \textbf{0.0016} & \textbf{4.6x10$^{-05}$} & \textbf{1.2x10$^{-09}$} &                         &                         &                 &             &        \\
    Y211T       & \textbf{0.016}   & \textbf{0.014}          & \textbf{0.0041} & \textbf{0.00016}        & \textbf{6.1x10$^{-09}$} & 0.77                    &                         &                 &             &        \\
   vR111A       & 0.38             & 0.18                    & 0.68            & 0.67                    & \textbf{0.016}          & \textbf{0.00032}        & \textbf{0.00094}        &                 &             &        \\
    N348A/Y211A & 0.061            & \textbf{0.013}          & 0.18            & 0.68                    & 0.17                    & \textbf{5.2x10$^{-06}$} & \textbf{2.1$^{-05}$}    & 0.38            &             &        \\
    N348W/Y211A & \textbf{0.00017} & \textbf{3.9x10$^{-06}$} & \textbf{0.0011} & \textbf{0.021}          & 0.68                    & \textbf{1.6x10$^{-10}$} & \textbf{6.3x10$^{-10}$} & \textbf{0.0047} & 0.063       &        \\
   vR111A/Y211A & 0.56             & 0.34                    & 0.86            & 0.47                    & \textbf{0.0072}         & \textbf{0.0010}         & \textbf{0.0025}         & 0.77            & 0.24        & \textbf{0.0019}  \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}
\fi

% \clearpage
% \begin {table}[H]\label{tab:
% \caption{Pairwise differences in interpolated MAF. Bolded values are statistically significant at $p<0.05$. $\Delta$ Max Force is the mean MAF for the first mutant listed in a row minus that of the second mutant listed.}
% \begin{center}
%   \resizebox{14cm}{!} {
%     \begin{tabular}{l r}
%     \hline
%     Mutant Pair                &  $\Delta$ Max Force \\ \hline
%     N348A-WT                   &               27.33 \\
%     N348W-WT                   &              -39.87 \\
%     N348K-WT                   &              -40.64 \\
%     \textbf{Y211A-WT}          &      \textbf{-87.97} \\
%     \textbf{Y211D-WT}          &      \textbf{103.44} \\
%     \textbf{Y211T-WT}          &       \textbf{89.17} \\
%     vR111A-WT                  &              -17.61 \\
%     N348A/Y211A-WT             &               -9.04 \\
%     \textbf{N348W/Y211A-WT}    &     \textbf{-154.79} \\
%     vR111A/Y211A-WT            &                8.71 \\
%     \textbf{N348W-N348A}       &      \textbf{-67.20} \\
%     \textbf{N348K-N348A}       &      \textbf{-67.97} \\
%     \textbf{Y211A-N348A}       &     \textbf{-115.30} \\
%     \textbf{Y211D-N348A}       &       \textbf{76.12} \\
%     \textbf{Y211T-N348A}       &       \textbf{61.85} \\
%     vR111A-N348A               &              -44.94 \\
%     N348A/Y211A-N348A          &              -36.37 \\
%     \textbf{N348W/Y211A-N348A} &     \textbf{-182.12} \\
%     vR111A/Y211A-N348A         &              -18.62 \\
%     N348K-N348W                &               -0.77 \\
%     Y211A-N348W                &              -48.10 \\
%     \textbf{Y211D-N348W}       &      \textbf{143.31} \\
%     \textbf{Y211T-N348W}       &      \textbf{129.04} \\
%     vR111A-N348W               &               22.26 \\
%     N348A/Y211A-N348W          &               30.83 \\
%     \textbf{N348W/Y211A-N348W} &     \textbf{-114.92} \\
%     vR111A/Y211A-N348W         &               48.58 \\
%     Y211A-N348K                &              -47.33 \\
%     
%     \hline
%     \end{tabular}
%     
%     \begin{tabular}{l r}
%     \hline
%     Mutant Pair                       &  $\Delta$ Max Force  \\ \hline
%     \textbf{Y211D-N348K}              &      \textbf{144.08} \\
%     \textbf{Y211T-N348K}              &      \textbf{129.81} \\
%     vR111A-N348K                      &               23.03  \\
%     N348A/Y211A-N348K                 &               31.60  \\
%     \textbf{N348W/Y211A-N348K}        &     \textbf{-114.15} \\
%     vR111A/Y211A-N348K                &               49.35  \\
%     \textbf{Y211D-Y211A}              &      \textbf{191.42} \\
%     \textbf{Y211T-Y211A}              &      \textbf{177.15} \\
%     \textbf{vR111A-Y211A}             &       \textbf{70.36} \\
%     \textbf{N348A/Y211A-Y211A}        &       \textbf{78.93} \\
%     \textbf{N348W/Y211A-Y211A}        &      \textbf{-66.82} \\
%     \textbf{vR111A/Y211A-Y211A}       &       \textbf{96.68} \\
%     Y211T-Y211D                       &              -14.27  \\
%     \textbf{vR111A-Y211D}             &     \textbf{-121.05} \\
%     \textbf{N348A/Y211A-Y211D}        &     \textbf{-112.48} \\
%     \textbf{N348W/Y211A-Y211D}        &     \textbf{-258.23} \\
%     \textbf{vR111A/Y211A-Y211D}       &      \textbf{-94.74} \\
%     \textbf{vR111A-Y211T}             &     \textbf{-106.78} \\
%     \textbf{N348A/Y211A-Y211T}        &      \textbf{-98.21} \\
%     \textbf{N348W/Y211A-Y211T}        &     \textbf{-243.96} \\
%     \textbf{vR111A/Y211A-Y211T}       &      \textbf{-80.46} \\
%     N348A/Y211A-vR111A                &                8.57  \\
%     \textbf{N348W/Y211A-vR111A}       &     \textbf{-137.18} \\
%     vR111A/Y211A-R111A                &               26.31  \\
%     \textbf{N348W/Y211A-N348A/Y211A}  &     \textbf{-145.75} \\
%     vR111A/Y211A-N348A/Y211A          &               17.75  \\
%     \textbf{vR111A/Y211A-N348W/Y211A} &      \textbf{163.50} \\ 
%                                       &          \\
%     \hline
%     \end{tabular}
%   }
% \end{center}
% \end{table}

\clearpage
\begin {table}[H]
\caption{\label{tab:MAF_p}Pairwise difference $p-$values for MAF. Bolded values are statistically significant at $p<0.05$.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT                      & N348A                   & N348W                   & N348K                   & Y211A                   & Y211D                     & Y211T                   & vR111A                  & N348A/Y211A             & N348W/Y211A \\ \hline
    N348A       & 0.35                    &                         &                         &                         &                         &                           &                         &                         &                         &        \\
    N348W       & 0.22                    & \textbf{0.012}          &                         &                         &                         &                           &                         &                         &                         &        \\
    N348K       & 0.22                    & \textbf{0.012}          & 0.98                    &                         &                         &                           &                         &                         &                         &        \\
    Y211A       & \textbf{0.0045}         & \textbf{1.6x10$^{-05}$} & 0.14                    & 0.15                    &                         &                           &                         &                         &                         &        \\
    Y211D       & \textbf{0.00083}        & \textbf{0.0045}         & \textbf{3.9x10$^{-06}$} & \textbf{3.9x10$^{-06}$} & \textbf{5.7x10$^{-10}$} &                           &                         &                         &                         &        \\
    Y211T       & \textbf{0.0042}         & \textbf{0.022}          & \textbf{3.0x10$^{-05}$} & \textbf{3.0x10$^{-05}$} & \textbf{1.0x10$^{-08}$} & 0.67                      &                         &                         &                         &        \\
   vR111A       & 0.59                    & 0.11                    & 0.51                    & 0.51                    & \textbf{0.024}          & \textbf{8.9x10$^{-05}$}   & \textbf{0.00056}        &                         &                         &        \\
    N348A/Y211A & 0.78                    & 0.20                    & 0.35                    & 0.35                    & \textbf{0.011}          & \textbf{0.00026}          & \textbf{0.0016}         & 0.78                    &                         &        \\
    N348W/Y211A & \textbf{6.2x10$^{-07}$} & \textbf{9.3x10$^{-12}$} & \textbf{0.00021}        & \textbf{0.00026}        & \textbf{0.032}          & \textbf{$<$ 2x10$^{-16}$} & \textbf{2.4x10$^{-15}$} & \textbf{9.1x10$^{-06}$} & \textbf{2.8x10$^{-06}$} &        \\
   vR111A/Y211A & 0.78                    & 0.52                    & 0.14                    & 0.14                    & \textbf{0.0019}         & \textbf{0.0024}           & \textbf{0.010}          & 0.44                    & 0.59                    & \textbf{1.6x10$^{-07}$}  \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\begin {table}[H]
\caption{\label{tab:AUC_p}Pairwise difference $p-$values for interpolated AUC. Bolded values are statistically significant at $p<0.05$.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT               & N348A                   & N348W           & N348K                   & Y211A                   & Y211D                   & Y211T                   & vR111A          & N348A/Y211A & N348W/Y211A \\ \hline
    N348A       & 0.77             &                         &                 &                         &                         &                         &                         &                 &             &        \\
    N348W       & 0.67             & 0.42                    &                 &                         &                         &                         &                         &                 &             &        \\
    N348K       & 0.17             & 0.056                   & 0.38            &                         &                         &                         &                         &                 &             &        \\
    Y211A       & \textbf{0.00094} & \textbf{3.5x10$^{-05}$} & \textbf{0.0045} & 0.063                   &                         &                         &                         &                 &             &        \\
    Y211D       & \textbf{0.0072}  & \textbf{0.0051}         & \textbf{0.0016} & \textbf{4.6x10$^{-05}$} & \textbf{1.2x10$^{-09}$} &                         &                         &                 &             &        \\
    Y211T       & \textbf{0.016}   & \textbf{0.014}          & \textbf{0.0041} & \textbf{0.00016}        & \textbf{6.1x10$^{-09}$} & 0.77                    &                         &                 &             &        \\
   vR111A       & 0.38             & 0.18                    & 0.68            & 0.67                    & \textbf{0.016}          & \textbf{0.00032}        & \textbf{0.00094}        &                 &             &        \\
    N348A/Y211A & 0.061            & \textbf{0.013}          & 0.18            & 0.68                    & 0.17                    & \textbf{5.2x10$^{-06}$} & \textbf{2.1$^{-05}$}    & 0.38            &             &        \\
    N348W/Y211A & \textbf{0.00017} & \textbf{3.9x10$^{-06}$} & \textbf{0.0011} & \textbf{0.021}          & 0.68                    & \textbf{1.6x10$^{-10}$} & \textbf{6.3x10$^{-10}$} & \textbf{0.0047} & 0.063       &        \\
   vR111A/Y211A & 0.56             & 0.34                    & 0.86            & 0.47                    & \textbf{0.0072}         & \textbf{0.0010}         & \textbf{0.0025}         & 0.77            & 0.24        & \textbf{0.0019}  \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

\end{document}

### 2013-06-19 paper.tex
Claus's requested changes.

@@
 As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants.
-\textbf{Supplement 1 ($\leftarrow$ what is this?
-figure?
-video?
-text?
-table?)} illustrates the separation distance between peptide domains.
+The supplementary movie visually illustrates the separation distance between peptide domains.
 The quantities MAF and AUC were derived from the force versus distances curves.
@@
 We would like to emphasize here that we cannot expect perfect agreement.
-First, while we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method.
+While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method.
 It is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method.
 More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph).
-Second, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays.
+In addition, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays.
 There are several reasons other than binding due to which a viral mutation may cause a phenotypic difference in infectivity.
-\textbf{Shorten the following (till end of paragraph), make less critical to experimentalists.
-Just list the things other than binding that may matter.} First, across disconnected experiments there could be a gross difference in the GP1 concentration.
-Although western blots are a popular means of protein concentration normalization, when used in a quantitative manner they are notoriously difficult and unreliable.
-In addition, though western blots could show gross differences in protein concentration, viral binding may be very sensitive to small differences in fusion protein concentration.
-Second, even if we trust protein expression normalization via western blot, it is not clear how much of the protein is grossly or partially misfolded.
+First, from infectivity experiments it is not clear the extent to which a protein is grossly or partially misfolded.
 An additional analytical step with circular dichroism or an analogous technique may further clarify large-scale folding differences.
-Third, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method.
+Second, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method.
+Third, although it is tempting to think diminished infectivity should correlate strongly with decreased binding affinity, it is likely that dramatically increased binding affinity would prohibit cellular membrane detachment after viral invasion.
+As a result, the virus could not get to its replication compartment (which is the cytosol in the case of Machupo) and would not reproduce.
 

### 2013-06-20 paper.tex
Copy edits and final changes before sending back.

@@
 \newpage
+
 \begin{abstract}
+
 Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases.
@@
 \section*{Introduction}
+
 The computational prediction of mutational effects on protein--protein interactions remains a challenging problem.
@@
 
-Here we propose a new method to probe protein--protein interactions directly, by pulling two proteins apart that are initially bound in a complex.
+Here we propose a new method to probe mutational effects in protein--protein interactions directly, by pulling two proteins apart that are initially bound in a complex.
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
@@
 \subsection*{System Modeling}
+
 We used the protein visualization software PyMOL \citep{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1.
@@
 \subsection*{Equilibration}
+
 NAMD was used for all simulations in this study \citep{Phillips2005}.
@@
 \subsection*{Steered Molecular Dynamics}
+
 Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1.
@@
 \subsection*{Post-processing}
+
 The python packages MDAnalysis \citep{Agrawal2011} and ProDy \citep{Bakan2011} were both used at various points in post-processing.
@@
 The latter approach was used for statistical analysis.
-In addition, to minimize the computational time in unproductive molecular dynamics, we shortened simulation time such that the interaction just barely dropped to zero near the end of each run.
+In addition, to minimize the computational time in unproductive molecular dynamics, we shortened the simulation time such that the interaction just barely dropped to zero near the end of each run.
 As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants.
@@
 Several mutations at this site have proven capable of causing loss-of-infection, according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
-Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 in the GP1.
+Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 in GP1.
 The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure \citep{Abraham2010}.
@@
 Asn348Lys is reported in the literature to cause reduced infectivity \textit{in vivo} \citep{Rad2008,Abraham2010}.
-Further, in the crystal structure two residues on the GP1 and at least three on hTfR1 are involved in a fairly extensive hydrogen-bonding network.
+Further, in the crystal structure two residues on GP1 and at least three on hTfR1 are involved in a fairly extensive hydrogen-bonding network.
 As discussed above, the alanine mutation at this site showed no significant difference in MAF or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}).
-In addition, the Asn348Lys mutation showed no significant difference from WT either, and neither did Asn348Trp.
+In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp.
 (For both of these mutations, mean MAF and mean AUC was lower than for WT, however.
@@
 
-Last, we further analyzed a single mutation in the GP1, vArg111Ala.
+Last, we further analyzed a single mutation in GP1, vArg111Ala.
 It has been reported that vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}.
@@
 In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application.
-Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanism.
+Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanisms.
 Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation.
@@
 Moreover, the molecular model based on x-ray diffraction data suggests destruction of a second hydrogen-bonding network as the cause \citep{Abraham2010}.
-In our simulations no individual point mutation at Asn348 showed significantly different affinity than the WT complex.
+In our simulations no individual point mutation at Asn348 showed a significantly different affinity from the WT complex.
 Moreover, the mutation Asn348Ala resulted in an average MAF that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}).
@@
 Third, although it is tempting to think diminished infectivity should correlate strongly with decreased binding affinity, it is likely that dramatically increased binding affinity would prohibit cellular membrane detachment after viral invasion.
-As a result, the virus could not get to its replication compartment (which is the cytosol in the case of Machupo) and would not reproduce.
+As a result, the virus would be prevented from reaching its replication compartment (which is the cytosol in the case of Machupo) and could not reproduce.
 
@@
 We applied our technique to the WT complex and 10 mutations in the GP1/hTfR1 interface.
-We were able to rationalize the results of infectivity data with a fundamental physical simulations.
+We were able to rationalize the results of infectivity data with fundamental physical simulations.
 Our data has provided some direction for further biochemical intuition in this system.

### 2013-06-20 paper.tex
Copy edits and final changes before sending back... second.

@@
 Second, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method.
-Third, although it is tempting to think diminished infectivity should correlate strongly with decreased binding affinity, it is likely that dramatically increased binding affinity would prohibit cellular membrane detachment after viral invasion.
-As a result, the virus would be prevented from reaching its replication compartment (which is the cytosol in the case of Machupo) and could not reproduce.
+Third, it is tempting to think only decreased binding affinity should correlate strongly with diminished infectivity.
+However, dramatically increased binding affinity would also disrupt infectivity by prohibiting cellular membrane detachment after viral invasion.
+The virus would be prevented from reaching its replication compartment (which is the cytosol in the case of Machupo) and could not reproduce.
 

### 2013-06-20 paper.tex
Some copy editing.

@@
 By contrast, first principles methods offer the chance to forgo training.
-However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model that phases directly from one three dimensional protein model to another with no intermediate transition states \citep{Gilson1997,Lu2004}.
+However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model that phases directly from one three-dimensional protein model to another with no intermediate transition states \citep{Gilson1997,Lu2004}.
 Thus, they can quickly lose physiological relevance as the two proteins become more dissimilar and as a result are largely limited to testing only sequential point mutants.
@@
 Machupo is a negative-sense RNA virus of the arenavirus family \citep{Charrel2003}.
-Worldwide arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}.
-Members of the arenavirus family include the lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the guanarito, junin and machupo viruses endemic to rodents in South America \citep{Charrel2003}.
+Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}.
+Members of the arenavirus family include the lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the guanarito, junin, and machupo viruses endemic to rodents in South America \citep{Charrel2003}.
 The South American arenaviruses typically infect humans after rodent contamination and cause a devastating hemorrhagic fever with very high mortality \citep{Charrel2003}.
@@
 
-We applied our method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes.
+We applied our computational method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes.
 At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity is often not as simple as the static structure suggests.
@@
 In addition, we first tuned the pulling velocity for practical time considerations.
-Subsequently, the spring constant was adjusted to ensure sufficient signal to noise with the given velocity.
-As a result, the spring constant was set to 5 (units), velocity set to 2 (units), and direction down the positive $z$-axis.
+Subsequently, the spring constant was adjusted to ensure a sufficient signal-to-noise ratio with the given velocity.
+As a result, the spring constant was set to 5~kcal/mol/\AA$^2$, velocity set to 1~\AA/ns, and direction down the positive $z$-axis.
 
@@
 We ran 50 replicate simulations per mutant.
-All GP1/hTfR1 complexes separated by greater than 4 \AA ngstroms and many separated to 10 or more.
+All GP1/hTfR1 complexes separated by greater than 4 \AA\ and many separated to 10 or more.
 
@@
 The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions.
-Next, with a fixed $\alpha$-carbon backbone we carried out successive rounds of energy minimization, followed by heating, and backbone release.
-Last,  with minimal restraints a short round of all-atom molecular dynamics \citep{Cuendet2008} was performed.
+Next, we carried out successive rounds of energy minimization with a fixed $\alpha$-carbon backbone, followed by heating, and backbone release.
+Last, we performed a short round of all-atom molecular dynamics with minimal restraints \citep{Cuendet2008}.
 We used the final state from each equilibrated system to restart another MD simulation.
@@
 The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure \citep{Abraham2010}.
-In agreement with the static structure, our molecular dynamics simulations show Tyr211 is involved in a large hydrogen-bonding network (results not shown).
+In agreement with the static structure, our molecular dynamics simulations show Tyr211 to be involved in a large hydrogen-bonding network (results not shown).
 
@@
 Asn348Lys is reported in the literature to cause reduced infectivity \textit{in vivo} \citep{Rad2008,Abraham2010}.
-Further, in the crystal structure two residues on GP1 and at least three on hTfR1 are involved in a fairly extensive hydrogen-bonding network.
+Further, in the crystal structure two residues on GP1 and at least three on hTfR1, including Asn348, are involved in a fairly extensive hydrogen-bonding network.
 As discussed above, the alanine mutation at this site showed no significant difference in MAF or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}).
 In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp.
-(For both of these mutations, mean MAF and mean AUC was lower than for WT, however.
-See Table~\ref{tab:summary_statistics})  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder.
+(For both of these mutations, however, mean MAF and mean AUC was lower than for WT.
+See Table~\ref{tab:summary_statistics}.)  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder.
 Moreover, Asn348Trp showed nearly identical results to Asn348Lys.
@@
 Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2012}.
-And still others aimed at more esoteric problems apply driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}.
-Here, we have shown that SMD can provide some insight into the \textit{relative} strength of protein--protein interactions.
+And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}.
+Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions.
 In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application.
@@
 We found that some substitutions at positions 211 and 348 did negatively affect the strength of viral binding.
-In the two cases, the computational data suggests that the reason and nature of the effects are very different.
-Congruent with infectivity data, mutations at Tyr211 often have a dramatic effect on receptor binding \citep{Rad2008,Rad20111}.
+However, the computational data suggest that the reason and nature of the effects at these two sites are very different.
+Congruent with infectivity data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding.
 Tyr211Ala alone produces such a large decrease in affinity that it can be easily distinguished from the WT complex.
@@
 Further, Tyr211Asp and Tyr211Thr were easily resolved from the WT complex, but they appear to cause an increase in affinity.
-As a result, we predict experimental structural analysis will show Tyr211Asp and Tyr211Thr mutations cause a relatively dramatic tertiary structure deviation.
+As a result, we predict experimental structural analysis will show that Tyr211Asp and Tyr211Thr mutations cause a relatively dramatic tertiary structure deviation.
 Given that our model starts from an associated complex, it is not surprising that Tyr211Asp would cause increased rather than decreased affinity.
@@
 While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method.
-It is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method.
+First, it is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method.
 More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph).
-In addition, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays.
-There are several reasons other than binding due to which a viral mutation may cause a phenotypic difference in infectivity.
-First, from infectivity experiments it is not clear the extent to which a protein is grossly or partially misfolded.
-An additional analytical step with circular dichroism or an analogous technique may further clarify large-scale folding differences.
-Second, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method.
-Third, it is tempting to think only decreased binding affinity should correlate strongly with diminished infectivity.
-However, dramatically increased binding affinity would also disrupt infectivity by prohibiting cellular membrane detachment after viral invasion.
+Second, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays.
+A mutant may cause a phenotypic difference in infectivity without generating a signal by our method.
+For example, infectivity could be lost in the experimental system because the protein is grossly or partially misfolded.
+An additional analytical step with circular dichroism or an analogous technique could clarify such large-scale folding differences.
+Further, since our simulations start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method.
+Finally, it is tempting to think only decreased binding affinity should correlate strongly with diminished infectivity.
+However, dramatically increased binding affinity could also disrupt infectivity by prohibiting cellular membrane detachment after viral invasion.
 The virus would be prevented from reaching its replication compartment (which is the cytosol in the case of Machupo) and could not reproduce.
@@
 Second, hybrid approaches exist that exploit non-equilibrium calculations within the context of all atom simulations \citep{Dixit2001,Is2001A,Is2001B,Park2004,Grant2011,Kortemme2004,Gumbart2012}.
-However, to date most of these methods have not been applied to protein--protein interactions (but rather to simpler scenarios such as protein--ligand interactions).
+However, to date most of these methods have not been applied to protein--protein interactions, but rather to simpler scenarios such as protein--ligand interactions.
 What we have shown here is that hybrid non-equilibrium approaches can provide a useful and practical tool in studying protein--protein interactions.
@@
 \centerline{\includegraphics[width=6in]{figures/protein_comparison.pdf}}
-\caption{\label{fig:protein_comparison}\title{Comparison of GP1/hTfR1 complex.} Figure A shows the full, de-glycosylated GP1/hTfR1 co-crystal structure.
-For figure B, we show the reduced structure for SMD.
-GP1 is blue and hTfR1 is green.}
+\caption{\label{fig:protein_comparison}\title{The GP1/hTfR1 complex.} GP1 is shown in blue and hTfR1 is shown in green.
+(A) The full, de-glycosylated GP1/hTfR1 co-crystal structure.
+(B) The reduced structure used in SMD simulations.}
 \end{figure}
@@
 \centerline{\includegraphics[width=6in]{figures/force_curve.pdf}}
-\caption{\label{fig:force_curve}\title{Force versus distance curve of Y211A mutant and WT.} The average force curve for 50 replicates of the WT complex was black, and the average of 50 replicates of the Y211A mutant was red.
-Notice there is a very large difference in MAF and AUC between the two complexes.}
+\caption{\label{fig:force_curve}\title{Force versus distance curve of WT and the Y211A mutant.} The average force curve for 50 replicates of the WT complex is shown in black, and the average of 50 replicates of the Y211A mutant is shown in red.
+There is a large difference in both MAF and AUC between the two complexes.}
 \end{figure}
@@
 \centerline{\includegraphics[width=6in]{figures/sensitivity_histogram.pdf}}
-\caption{\label{fig:sensitivity}\title{Density plot of three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left.
-The large non-overlapping areas indicate a large and statistically significant difference in these three complexes with relatively minor differences.}
+\caption{\label{fig:sensitivity}\title{Distribution of interpolated maximum force (MAF) for three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left.
+The large non-overlapping areas indicate a large and statistically significant difference in these three complexes.}
 \end{figure}

### 2013-06-26 paper.tex
Major changes in line with Andy and Saras requests.

@@
 By contrast, first principles methods offer the chance to forgo training.
-However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model that phases directly from one three-dimensional protein model to another with no intermediate transition states \citep{Gilson1997,Lu2004}.
-Thus, they can quickly lose physiological relevance as the two proteins become more dissimilar and as a result are largely limited to testing only sequential point mutants.
+However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional protein model to another with no intermediate steps \citep{Gilson1997,Lu2004}.
+As a result, the accuracy of the measurement suffers as the two proteins become increasingly dissimilar; therefore, these techniques are largely limited to point mutant comparisons.
 
@@
 Then, we measure the force required for dissociation.
-As a proxy for relative binding affinity, we use the maximum applied force (MAF) required for separation and the area under the force-versus-distance curve (AUC).
+As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
 
 We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}.
-Machupo is a negative-sense RNA virus of the arenavirus family \citep{Charrel2003}.
+Machupo is an ambisense RNA virus of the arenavirus family \citep{Charrel2003}.
 Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}.
-Members of the arenavirus family include the lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the guanarito, junin, and machupo viruses endemic to rodents in South America \citep{Charrel2003}.
-The South American arenaviruses typically infect humans after rodent contamination and cause a devastating hemorrhagic fever with very high mortality \citep{Charrel2003}.
+Members of the arenavirus family include the Lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the Guanarito, Junin, and Machupo viruses endemic to rodents in South America \citep{Charrel2003}.
+The South American arenaviruses typically infect humans after rodent contamination and can cause a devastating hemorrhagic fever with high mortality \citep{Charrel2003}.
 
 The hTfR1 is the primary receptor used by MACV for binding its host cell prior to infection.
-The hTfR1 protein contains three extracellular domains with the two more basilar domains serving most of the endogenous, transferrin-binding function \citep{Abraham2010,Rad20112}.
-Cellular entry is initiated by GP1 binding to the apical domain of hTfR1.
-Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of viral infectivity \citep{Rad20111,Rad20112}.
-The co-crystal structure shows that a 15 amino acid flexible loop on GP1 is responsible for viral binding \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
-Furthermore, the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain.
+The primary role of hTfR1 $in~vivo$ is to bind transferrin for cellular iron uptake.
+The hTfR1 protein contains three extracellular domains: two basilar domains and an apical domain.
+The two basilar domains serve most of the transferrin-binding function \citep{Abraham2010,Rad20112}.
+Viral entry is initiated by GP1 binding to the apical domain of hTfR1.
+Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}.
+The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain.
+For GP1, several extended loops in mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
 Experimental alanine-scanning and whole cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.
@@
 For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1.
-According to our simulations, this network likely does not affect binding affinity.
-Our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in host-virus protein--protein interfaces.
+According to our simulations, this network likely does not affect binding affinity directly.
+Our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in protein--protein interfaces.
 
@@
 To reduce the noise, we averaged each point with a 75 point interpolation window.
-The window ensured that the averaged MAF was never cut off near the beginning of the simulation due to lost time steps (caused by averaging).
-The two primary descriptive statistics we used were maximum interpolated applied force (MAF) and total area under the interpolated curve (AUC).
+The window ensured that the averaged maximum applied force was never cut off near the beginning of the simulation due to lost time steps (caused by averaging).
+The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC).
 We used the pairwise.t.test function with the FDR p-value correction to test for significance.
@@
 \subsection*{The GP1/hTfR1 system}
-
 The GP1/hTfR1 interface marks a particularly important and useful test system.
 There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV.
-Virtually all of the important sites have been mapped by \textit{in vitro} flow-cytometry based infectivity assays.
+Virtually all of the important sites have been mapped by \textit{in vitro} flow-cytometry based entry assays.
 The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals).
@@
 Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the effect of those contacts \textit{in vivo} is still unresolved \citep{Abraham2010}.
+
+In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants}).
+Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
+Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1.
+The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}.
+In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1.
+In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}.
+Finally, an alanine mutation at site 111 in GP1 has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}.
+For notation purposes, the viral site is always referred to with a preceding `v'.
 
@@
 The supplementary movie visually illustrates the separation distance between peptide domains.
-The quantities MAF and AUC were derived from the force versus distances curves.
+The quantities maximum applied force and AUC were derived from the force versus distances curves.
 Their summary statistics are reported in Table~\ref{tab:summary_statistics}.
@@
 For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.
-Several residues on both proteins appear to mediate putative hydrogen-bonding interactions.
-We selected three sites for alanine substitutions: Tyr211 and Asn348 in the hTfR1 and vArg111 in MACV.
-For notation purposes, the viral site is always referred to with a preceding `v'.
-According to the static complex \citep{Abraham2010}, Tyr211 and vArg111 are involved in one hydrogen-bonding network and Asn348 is involved in a second, distinct hydrogen-bonding network.
 
-We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the MAF and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:pairwise_differences}), while the other two did not (Table~\ref{tab:pairwise_differences}).
+We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:pairwise_differences}), while the other two did not (Table~\ref{tab:pairwise_differences}).
 From these results, we concluded that our method can indeed successfully distinguish individual mutants from WT, and that alanine substitutions at Asn348 and vArg111 may have little effect on binding affinity (see below for further discussion).
-When considering additional mutants (also discussed below), we found that MAF was generally sufficient to distinguish mutants (Tables~\ref{tab:pairwise_differences} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}).
-In general, however, MAF seemed to be the more sensitive statistic than AUC.
+When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:pairwise_differences} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}).
+In general, however, maximum applied force seemed to be the more sensitive statistic than AUC.
 
 \subsection*{Comparative analysis of the GP1/hTfR1 interface}
-
 Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction.
 One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function.
-We made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine.
+Seen in Table~\ref{tab:summary_mutant}), we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine.
 Several mutations caused significant relative affinity changes.
@@
 
-In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_statistics}).
-The largest single-mutant impact was seen for a mutation to alanine at site Tyr211 in the hTfR1 ($ p < 0.001 $).
-Several mutations at this site have proven capable of causing loss-of-infection, according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
-Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 in GP1.
-The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure \citep{Abraham2010}.
-In agreement with the static structure, our molecular dynamics simulations show Tyr211 to be involved in a large hydrogen-bonding network (results not shown).
-
-Although Tyr211Ala appears to have a large impact on binding affinity, the single mutant provided little underlying biochemical rationale.
+Although Tyr211Ala appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism.
 Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for loss of binding affinity.
@@
 In fact, they caused a significant increase in relative binding affinity (Figure~\ref{fig:sensitivity} and Table~\ref{tab:MAF_p}).
-This outcome is not surprising, if we consider that Tyr211 participates in a bonded network rather than independently.
-Combined with some biochemical intuition, these results suggest that mutations to other polar residues would likely not affect the dynamics of binding directly.
 
 We also simulated several point mutations at Asn348 in the hTfR1.
-Asn348Lys is reported in the literature to cause reduced infectivity \textit{in vivo} \citep{Rad2008,Abraham2010}.
-Further, in the crystal structure two residues on GP1 and at least three on hTfR1, including Asn348, are involved in a fairly extensive hydrogen-bonding network.
-As discussed above, the alanine mutation at this site showed no significant difference in MAF or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}).
+As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}).
 In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp.
-(For both of these mutations, however, mean MAF and mean AUC was lower than for WT.
+(For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT.
 See Table~\ref{tab:summary_statistics}.)  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder.
@@
 As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study.
-Further, according to MAF (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}).
+Further, according to maximum applied force (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}).
 We suspect that the effect of Asn348Trp alone is near the limit of detection using our method.
@@
 Last, we further analyzed a single mutation in GP1, vArg111Ala.
-It has been reported that vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}.
-As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either MAF or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}).
+As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either maximum applied force or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}).
 This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity.
@@
 \section*{Discussion}
+
 We have applied a method utilizing steering forces in all-atom molecular dynamics simulations to evaluate the effects of mutations at the GP1/hTfR1 interface.
 We modeled mutations at several sites in the GP1/hTfR1 interface, and verified that our computational protocol was sensitive enough to distinguish point mutants in hTfR1.
-Further, we identified two test statistics, MAF and AUC, that can be used as proxies for binding affinity.
+Further, we identified two test statistics, maximum applied force and AUC, that can be used as proxies for binding affinity.
 We systematically tested several point mutations to understand their contribution to the binding interaction.
@@
 Our findings rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}.
-We found that some substitutions at positions 211 and 348 did negatively affect the strength of viral binding.
+First, we found that some substitutions at positions 211 and 348 did negatively affect the strength of receptor binding.
 However, the computational data suggest that the reason and nature of the effects at these two sites are very different.
-Congruent with infectivity data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding.
-Tyr211Ala alone produces such a large decrease in affinity that it can be easily distinguished from the WT complex.
-Structural analysis reveals that the biochemical cause involves the destruction of a critical hydrogen-bonding network with Ser113 and vArg111 in GP1.
+Congruent with entry data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding.
+Second, no individual point mutation at Asn348 showed a significantly different affinity from the WT complex.
+However, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference.
+That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually.
+Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
+Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.
+As mutations at hTfR1 site 211 are known to reduce infectivity \emph{in vitro} and \emph{in vivo}, it is possible that vArg111 in GP1 can serve as a site for compensatory MACV mutations.
+
+%Structural analysis reveals that the biochemical cause involves the destruction of a critical hydrogen-bonding network with Ser113 and vArg111 in GP1.
 Further, Tyr211Asp and Tyr211Thr were easily resolved from the WT complex, but they appear to cause an increase in affinity.
@@
 
-Previous work has shown that the hTfR1 mutation Asn348Lys is capable of producing decreased infectivity \emph{in vitro} \citep{Rad2008,Rad20111}.
+%Previous work has shown that the hTfR1 mutation Asn348Lys is capable of producing decreased infectivity \emph{in vitro} \citep{Rad2008,Rad20111}.
 Moreover, the molecular model based on x-ray diffraction data suggests destruction of a second hydrogen-bonding network as the cause \citep{Abraham2010}.
 In our simulations no individual point mutation at Asn348 showed a significantly different affinity from the WT complex.
-Moreover, the mutation Asn348Ala resulted in an average MAF that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}).
-Thus, the putative hydrogen-bonding network involving Asn348 is probably a crystallization artifact.
+Moreover, the mutation Asn348Ala resulted in an average maximum applied force that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}).
 Even when paired with the highly destructive Tyr211Ala mutation, Asn348Ala was refractory to changes in binding affinity.
@@
 That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually.
-Moreover, although Asn348Trp and Asn348Lys individually failed to be distinguishable from WT hTfR1, they both caused a mean reduction in MAF and produced larger than expected synergistic effects.
+Moreover, although Asn348Trp and Asn348Lys individually failed to be distinguishable from WT hTfR1, they both caused a mean reduction in maximum applied force and produced larger than expected synergistic effects.
 This data suggests that site 348 is size and charge restricted.
@@
 
-The GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
-We suspected that this result was primarily caused by a lack of sensitivity of our method, since vArg111's binding orientation near Tyr211 and involvement in the critical hydrogen-bonding network strongly suggested that mutations at that site should have an effect on binding affinity.
-As an alternative way to test this hypothesis, we evaluated whether we could introduce a suppressor mutant at site vArg111.
-We introduced vArg111Ala jointly with Tyr211Ala, thus effectively replacing a hydrogen-bonding network with a hydrophobic interaction.
-Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean MAF to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.
-As mutations at hTfR1 site 211 are known to reduce infectivity \emph{in vitro} and \emph{in vivo}, it is possible that vArg111 in GP1 can serve as a supressor site for compensatory MACV mutations.
-Thus, this site may be critically important for surveillance in the rodent population.
-
-Although our computational predictions largely agreed with the available literature, some mutations that are reported to cause reduced infectivity in experimental systems (most notably vArg111Ala and Asn348Lys, discussed above) were indistinguishable from the WT complex in our simulations.
 We would like to emphasize here that we cannot expect perfect agreement.
@@
 More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph).
-Second, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays.
+Second, our SMD method is fundamentally a measurement tool, and as a result we are dramatically limited by the accuracy of our starting structure.
+Third, the available experimental data for the GP1/hTfR1 system were generally obtained from entry assays or whole-cell binding assays rather than molecular binding assays.
 A mutant may cause a phenotypic difference in infectivity without generating a signal by our method.
-For example, infectivity could be lost in the experimental system because the protein is grossly or partially misfolded.
+For example, entry could be lost in the experimental system because the protein is grossly or partially misfolded.
 An additional analytical step with circular dichroism or an analogous technique could clarify such large-scale folding differences.
-Further, since our simulations start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method.
+Further, since our simulations start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) or relative orientation of the two proteins would be underestimated by our method.
 Finally, it is tempting to think only decreased binding affinity should correlate strongly with diminished infectivity.
@@
 
-As with any structural approach, we are constrained to start our computational model system from a static structure.
-One unfortunate reality in structural biology is poor resolution/discrimination at particularly long and flexible amino acid side chains.
-In addition, the number of acceptable side-chain rotamers increases in large amino acids.
-As a consequence there could be many untested starting conformations.
-Although we believe we have minimized unsampled conformations here, such an issue could be overcome with a sufficiently large number of SMD starting models.
-Of course, increasing sampling in this case comes at significant computational cost.
 There are a few additional challenges for investigating host-virus interaction via molecular dynamics simulation.
@@
 
-Making accurate predictions regarding the effects of non-alanine mutations at protein--protein interfaces remains  challenging.
+%Making accurate predictions regarding the effects of non-alanine mutations at protein--protein interfaces remains  challenging.
 The main obstacle is the enormous number of chemically allowed three dimensional configurations in proteins \citep{Gumbart2012}.
@@
 We were able to rationalize the results of infectivity data with fundamental physical simulations.
-Our data has provided some direction for further biochemical intuition in this system.
+Our data has provided some direction for further biochemical investigation in this system.
 A polar residue is absolutely critical at position 211 in hTfR1.
 Any mutation to a nonpolar residue will probably lead to a rapid viral compensatory response.
-Also, although rodent populations may produce further mutations at Asn348 in hTfR1, unless a mutation is positively charged or very large it is unlikely to result in a large disruption to viral binding.
+Also, although rodent populations may produce further mutations at Asn348 in hTfR1, unless a mutation is positively charged or very large it is unlikely to result in a large disruption to receptor binding.
 We have also provided both a site (vArg111) and a mutation (vArg111Ala) that may be critical for future host-range adaptation of MACV, and thus should be monitored in surveillance efforts.
@@
 
-This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) to A.D.E., and the National Institutes of Health (R01-GM088344) to C.O.W.
+This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) to A.D.E., the National Institutes of Health (R01-GM088344) to C.O.W., and the National Institutes of Health (R01-GM093086) to S.L.S.
 
@@
 \caption{\label{fig:force_curve}\title{Force versus distance curve of WT and the Y211A mutant.} The average force curve for 50 replicates of the WT complex is shown in black, and the average of 50 replicates of the Y211A mutant is shown in red.
-There is a large difference in both MAF and AUC between the two complexes.}
+There is a large difference in both maximum applied force and AUC between the two complexes.}
 \end{figure}
@@
 \begin {table}[H]
+\caption{\label{tab:summary_mutant}Summary of information for each mutation tested.
+Observed $in~vivo$ refers to mutations that have been observed in rodent populations.
+Phenotype $in~vitro$ refers to the observed phenotype in $in~vitro$ viral entry assays.}
+%Phenotype $in~silico$ refers to mutations that were statistically significant relative to wild-type only. 
+\begin{center}
+  \resizebox{11cm}{!} {
+    \begin{tabular}{l c r r}
+    \hline
+      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\%& Phenotype $in~silico$ \\ \hline
+            WT &                Yes &         Normal Entry \\%&        Normal Binding \\
+         N348A &                 No &                    - \\%&       Not Significant \\
+         N348K &                Yes &     Diminished Entry \\%&       Not Significant \\
+         N348W &                 No &                    - \\%&       Not Significant \\
+         Y211A &                 No &        No Expression \\%&     Decreased Binding \\ 
+         Y211D &                Yes &        No Expression \\%&     Increased Binding \\ 
+         Y211T &                 No &     Diminished Entry \\%&     Increased Binding \\
+        vR111A &                 No &     Diminished Entry \\%&       Not Significant \\
+   N348A/Y211A &                 No &                    - \\%&       Not Significant \\
+   N348W/Y211A &                 No &                    - \\%&     Decreased Binding \\
+  vR111A/Y211A &                 No &                    - \\%&       Not Significant \\
+    \hline
+    \end{tabular}
+  }
+\end{center}
+\end{table}
+
+\clearpage
+\begin {table}[H]
 \caption{\label{tab:summary_statistics}Summary statistics for each mutation tested.
-$\mu_\text{MAF}$ is the mean and $\sigma_\text{MAF}$ is the standard deviation of MAF over all simulations.
+$\mu_\text{MAF}$ is the mean and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations.
 $\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations.}
@@
          N348W &                732.13 &              131.21 &         192043.2 &            66916.36 \\
-        vR111A &                 754.39 & 103.25               & 185896.8    & 46408.74 \\
          Y211A &                684.03 &              126.69 &         153517.7 &            40889.12 \\ 
@@
          Y211T &                861.17 &              158.48 &         231078.7 &            97569.29 \\
+        vR111A &                754.39 &              103.25 &         185896.8 &            46408.74 \\
    N348A/Y211A &                762.96 &              144.81 &         173059.2 &            59898.10 \\
@@
 \begin {table}[H]
-\caption{\label{tab:pairwise_differences}Pairwise differences (row variable minus column variable) in mean MAF.
+\caption{\label{tab:pairwise_differences}Pairwise differences (row variable minus column variable) in mean maximum applied force.
 Bolded values are statistically significant at $p<0.05$.
@@
 % \begin {table}[H]\label{tab:
-% \caption{Pairwise differences in interpolated MAF.
+% \caption{Pairwise differences in interpolated maximum applied force.
 Bolded values are statistically significant at $p<0.05$.
-$\Delta$ Max Force is the mean MAF for the first mutant listed in a row minus that of the second mutant listed.}
+$\Delta$ Max Force is the mean maximum applied force for the first mutant listed in a row minus that of the second mutant listed.}
 % \begin{center}
@@
 \begin {table}[H]
-\caption{\label{tab:MAF_p}Pairwise difference $p-$values for MAF. Bolded values are statistically significant at $p<0.05$.}
+\caption{\label{tab:MAF_p}Pairwise difference $p-$values for maximum applied force. Bolded values are statistically significant at $p<0.05$.}
 \begin{center}

### 2013-07-01 paper.tex
Updated paper and added mutant comparison boxplot.

@@
 \documentclass[12pt]{article}
+\pdfoutput=1
 
@@
 Our method is based on extensive non-equilibrium all-atom simulations: We computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1) and estimate affinity using the maximum applied force during a pulling simulation and the area under the force-versus-distance curve.
-We find that these quantities provide novel biophysical insight into the GP1/hTfR1 interaction.
+We find that these quantities can provide novel biophysical insight into the GP1/hTfR1 interaction.
 First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes.
@@
 
-Here we propose a new method to probe mutational effects in protein--protein interactions directly, by pulling two proteins apart that are initially bound in a complex.
+Here we propose a new method to probe mutational effects in protein--protein interactions directly based on first principles simulation, by pulling two proteins apart that are initially bound in a complex.
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
@@
 We applied our computational method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes.
-At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity is often not as simple as the static structure suggests.
+At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity may not be as simple as the static structure suggests.
 For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1.
@@
 Thus, vArg111Ala in the GP1 appeared capable of suppressing the loss-of-binding caused by Tyr211Ala.
-This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction leads to comparable binding affinity.
+This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction could lead to comparable binding affinity.
 
@@
 
-Our findings rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}.
-First, we found that some substitutions at positions 211 and 348 did negatively affect the strength of receptor binding.
+Our findings may rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}.
+First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding.
 However, the computational data suggest that the reason and nature of the effects at these two sites are very different.
@@
 As mutations at hTfR1 site 211 are known to reduce infectivity \emph{in vitro} and \emph{in vivo}, it is possible that vArg111 in GP1 can serve as a site for compensatory MACV mutations.
-
-%Structural analysis reveals that the biochemical cause involves the destruction of a critical hydrogen-bonding network with Ser113 and vArg111 in GP1.
-Further, Tyr211Asp and Tyr211Thr were easily resolved from the WT complex, but they appear to cause an increase in affinity.
-As a result, we predict experimental structural analysis will show that Tyr211Asp and Tyr211Thr mutations cause a relatively dramatic tertiary structure deviation.
-Given that our model starts from an associated complex, it is not surprising that Tyr211Asp would cause increased rather than decreased affinity.
-Aspartate is charged rather than polar at physiological pH and its shorter side chain means the proteins would be pulled in closer apposition.
-Nevertheless, our results show Tyr211, or some other amino acid that is capable of hydrogen bonding, is necessary for a stable hydrogen-bonding network leading to efficient binding.
-Moreover, the cause of decreased infectivity seen in rodent populations with the Tyr211Asp mutation is not a simple change in binding affinity.
-
-%Previous work has shown that the hTfR1 mutation Asn348Lys is capable of producing decreased infectivity \emph{in vitro} \citep{Rad2008,Rad20111}.
-Moreover, the molecular model based on x-ray diffraction data suggests destruction of a second hydrogen-bonding network as the cause \citep{Abraham2010}.
-In our simulations no individual point mutation at Asn348 showed a significantly different affinity from the WT complex.
-Moreover, the mutation Asn348Ala resulted in an average maximum applied force that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}).
-Even when paired with the highly destructive Tyr211Ala mutation, Asn348Ala was refractory to changes in binding affinity.
-By contrast, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference.
-That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually.
-Moreover, although Asn348Trp and Asn348Lys individually failed to be distinguishable from WT hTfR1, they both caused a mean reduction in maximum applied force and produced larger than expected synergistic effects.
-This data suggests that site 348 is size and charge restricted.
-Any mutation to a very large or positively charged side chain will likely produce decreased infectivity.
 
@@
 
-%Making accurate predictions regarding the effects of non-alanine mutations at protein--protein interfaces remains  challenging.
-The main obstacle is the enormous number of chemically allowed three dimensional configurations in proteins \citep{Gumbart2012}.
-In theory, one could carry out all-atom simulations from physical first principles \citep{Gumbart2012} for very long time periods.
-After equilibrium is reached in such simulations, one could then compute a binding constant by observing the fractional association/dissociation time among two proteins.
-Unfortunately, due to limited computational power and available methods, reaching equilibrium with a simulated protein system is truly a herculean task and is essentially impossible at this time \citep{Gumbart2012}.
-There are two broadly useful computational approaches that avoid first-principles equilibrium simulations.
-First, to reduce computational complexity several studies have used rigid body docking and machine learning \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}.
-While these approaches can work well in certain scenarios, they often underperform in applications to novel systems, due to their dependence on available training data.
-Second, hybrid approaches exist that exploit non-equilibrium calculations within the context of all atom simulations \citep{Dixit2001,Is2001A,Is2001B,Park2004,Grant2011,Kortemme2004,Gumbart2012}.
-However, to date most of these methods have not been applied to protein--protein interactions, but rather to simpler scenarios such as protein--ligand interactions.
-What we have shown here is that hybrid non-equilibrium approaches can provide a useful and practical tool in studying protein--protein interactions.
-
 We have probed the biochemical effects of mutations at the host-virus interface between MACV and its human receptor.
@@
 We applied our technique to the WT complex and 10 mutations in the GP1/hTfR1 interface.
-We were able to rationalize the results of infectivity data with fundamental physical simulations.
 Our data has provided some direction for further biochemical investigation in this system.
@@
 Also, although rodent populations may produce further mutations at Asn348 in hTfR1, unless a mutation is positively charged or very large it is unlikely to result in a large disruption to receptor binding.
-We have also provided both a site (vArg111) and a mutation (vArg111Ala) that may be critical for future host-range adaptation of MACV, and thus should be monitored in surveillance efforts.
 Finally, our analysis offers a new means to systemically investigate host-virus interactions through computational molecular dynamics.

### 2013-07-08 paper.tex
Slight change in the mutant comparison figure.
I wanted to push changes before deleting more text.

@@
 First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes.
-Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that only one of them is critical for the binding interaction.
+Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that one of them may not be important for tight binding.
 Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-resistant hTfR1 mutants.
@@
 For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1.
-According to our simulations, this network likely does not affect binding affinity directly.
-Our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in protein--protein interfaces.
+According to our simulations, this network may not affect binding affinity directly.
+In addition, our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in protein--protein interfaces.
 
@@
 For each simulation, we randomly selected one of the equilibration runs for restart.
-We ran 50 replicate simulations per mutant.
+We ran 50 replicate simulations per mutant for a total of 550 SMD simulations.
 All GP1/hTfR1 complexes separated by greater than 4 \AA\ and many separated to 10 or more.
@@
 Each of the 50 independent trajectories per mutant produced a fairly noisy force curve.
-To reduce the noise, we averaged each point with a 75 point interpolation window.
+To reduce the noise, we averaged each point with a 50 point interpolation window.
 The window ensured that the averaged maximum applied force was never cut off near the beginning of the simulation due to lost time steps (caused by averaging).
@@
 
-In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants}).
+In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}).
 Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
@@
 
-We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:pairwise_differences}), while the other two did not (Table~\ref{tab:pairwise_differences}).
-From these results, we concluded that our method can indeed successfully distinguish individual mutants from WT, and that alanine substitutions at Asn348 and vArg111 may have little effect on binding affinity (see below for further discussion).
-When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:pairwise_differences} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}).
+We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:MAF_pairwise}), while the other two did not (Table~\ref{tab:MAF_pairwise}).
+When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:MAF_pairwise} and ~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}).
 In general, however, maximum applied force seemed to be the more sensitive statistic than AUC.
@@
 Several mutations caused significant relative affinity changes.
-In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause reduced binding.
-Then, we compared the size of those differences to reduced binding single mutants (Figure~\ref{fig:sensitivity}).
+In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause similar changes in binding.
+Then, we compared the size of those differences to single mutants (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}).
 
 Although Tyr211Ala appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism.
-Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for loss of binding affinity.
+Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for change in binding affinity.
 In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding.
 We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}.
-Both mutations proved incapable of causing reduced binding affinity in our simulations.
-In fact, they caused a significant increase in relative binding affinity (Figure~\ref{fig:sensitivity} and Table~\ref{tab:MAF_p}).
+Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity.
+(Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).
 
@@
 Moreover, Asn348Trp showed nearly identical results to Asn348Lys.
-These results suggest that, in contrast to Tyr211, the critical difference at position 348 in hTfR1 is size exclusion rather than the destruction of a hydrogen-bonding network.
-The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:pairwise_differences}).
+The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:MAF_pairwise}).
 To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding.
@@
 To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala.
-This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:pairwise_differences}).
-Thus, vArg111Ala in the GP1 appeared capable of suppressing the loss-of-binding caused by Tyr211Ala.
-This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction could lead to comparable binding affinity.
+This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:MAF_pairwises}).
+This result shows that the two sites do indeed interact, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction could lead to comparable binding affinity.
 
@@
 In the case of Asn348Lys, we have shown that the static structure provides little insight into why this mutation causes loss-of-infectivity \textit{in vivo}.
-While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, loss of binding affinity at that site is actually caused by size and charge restriction.
-Further, we have shown that the cause of diminished infectivity in the viral single-mutant vArg111Ala may not be due to binding-energy differences.
+While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, change in binding at that site may actually be caused by size and charge restriction.
 We also found that a negatively polar residue at site 211 in hTfR1 is absolutely critical for a tight binding interaction.
 Any non-polar mutation at Tyr211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection.
-However, our data suggest that the viral mutation vArg111Ala may restore the infective potential of MACV that was previously destroyed by a Tyr211Ala hTfR1 mutation.
 
 By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application.
-Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Buch2011,Giorgino2012}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
-Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2012}.
+Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
+Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}.
 And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}.
@@
 Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.
-As mutations at hTfR1 site 211 are known to reduce infectivity \emph{in vitro} and \emph{in vivo}, it is possible that vArg111 in GP1 can serve as a site for compensatory MACV mutations.
 
@@
 With improved resources, we could investigate thousands of individual permutations in the GP1/hTfR1 binding interface.
+In addition, with additional compute time it would be possible to incorporate equilibrium sampling approaches \citep{Buch2011} or use brute force equilibrium approaches \citep{Giorgino2012} to improve resolution.
 Going forward, we believe that all-atom molecular dynamics simulations can become an increasingly viable approach for protein interaction studies.
@@
 \centerline{\includegraphics[width=6in]{figures/sensitivity_histogram.pdf}}
-\caption{\label{fig:sensitivity}\title{Distribution of interpolated maximum force (MAF) for three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left.
+\caption{\label{fig:sensitivity}\title{Distribution of interpolated maximum force for three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left.
 The large non-overlapping areas indicate a large and statistically significant difference in these three complexes.}
@@
 \clearpage
+\begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/mutant_comparison_boxplot.pdf}}
+\caption{\label{fig:mutant_comparison}\title{Distribution of interpolated maximum force for all complexes tested.} The statistically significant difference relative to the WT complex are indicated by a star above the distribution.}
+\end{figure}
+
+\clearpage
 \begin {table}[H]
-\caption{\label{tab:summary_mutant}Summary of information for each mutation tested.
+\caption{\label{tab:summary_mutants}Summary of information for each mutation tested.
 Observed $in~vivo$ refers to mutations that have been observed in rodent populations.
 Phenotype $in~vitro$ refers to the observed phenotype in $in~vitro$ viral entry assays.}
-%Phenotype $in~silico$ refers to mutations that were statistically significant relative to wild-type only. 
 \begin{center}
@@
     \hline
-      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\%& Phenotype $in~silico$ \\ \hline
-            WT &                Yes &         Normal Entry \\%&        Normal Binding \\
-         N348A &                 No &                    - \\%&       Not Significant \\
-         N348K &                Yes &     Diminished Entry \\%&       Not Significant \\
-         N348W &                 No &                    - \\%&       Not Significant \\
-         Y211A &                 No &        No Expression \\%&     Decreased Binding \\ 
-         Y211D &                Yes &        No Expression \\%&     Increased Binding \\ 
-         Y211T &                 No &     Diminished Entry \\%&     Increased Binding \\
-        vR111A &                 No &     Diminished Entry \\%&       Not Significant \\
-   N348A/Y211A &                 No &                    - \\%&       Not Significant \\
-   N348W/Y211A &                 No &                    - \\%&     Decreased Binding \\
-  vR111A/Y211A &                 No &                    - \\%&       Not Significant \\
+      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\
+            WT &                Yes &         Normal Entry \\
+         N348A &                 No &                    - \\
+         N348K &                Yes &     Diminished Entry \\
+         N348W &                 No &                    - \\
+         Y211A &                 No &        No Expression \\
+         Y211D &                Yes &        No Expression \\
+         Y211T &                 No &     Diminished Entry \\
+        vR111A &                 No &     Diminished Entry \\
+   N348A/Y211A &                 No &                    - \\
+   N348W/Y211A &                 No &                    - \\
+  vR111A/Y211A &                 No &                    - \\
     \hline
@@
 \begin {table}[H]
-\caption{\label{tab:pairwise_differences}Pairwise differences (row variable minus column variable) in mean maximum applied force.
+\caption{\label{tab:MAF_pairwise}Pairwise differences (row variable minus column variable) in mean maximum applied force.
 Bolded values are statistically significant at $p<0.05$.
@@
 \end{table}
-
-\iffalse
-\begin {table}[H]
-\caption{Pairwise difference $p-$values for interpolated AUC. Bolded values are statistically significant at $p<0.05$.}
-\begin{center}
-  \resizebox{16cm}{!} {
-    \begin{tabular}{l l l l l l l l l l l}
-    \hline
-                & WT               & N348A                   & N348W           & N348K                   & Y211A                   & Y211D                   & Y211T                   & vR111A          & N348A/Y211A & N348W/Y211A \\ \hline
-    N348A       & 0.77             &                         &                 &                         &                         &                         &                         &                 &             &        \\
-    N348W       & 0.67             & 0.42                    &                 &                         &                         &                         &                         &                 &             &        \\
-    N348K       & 0.17             & 0.056                   & 0.38            &                         &                         &                         &                         &                 &             &        \\
-    Y211A       & \textbf{0.00094} & \textbf{3.5x10$^{-05}$} & \textbf{0.0045} & 0.063                   &                         &                         &                         &                 &             &        \\
-    Y211D       & \textbf{0.0072}  & \textbf{0.0051}         & \textbf{0.0016} & \textbf{4.6x10$^{-05}$} & \textbf{1.2x10$^{-09}$} &                         &                         &                 &             &        \\
-    Y211T       & \textbf{0.016}   & \textbf{0.014}          & \textbf{0.0041} & \textbf{0.00016}        & \textbf{6.1x10$^{-09}$} & 0.77                    &                         &                 &             &        \\
-   vR111A       & 0.38             & 0.18                    & 0.68            & 0.67                    & \textbf{0.016}          & \textbf{0.00032}        & \textbf{0.00094}        &                 &             &        \\
-    N348A/Y211A & 0.061            & \textbf{0.013}          & 0.18            & 0.68                    & 0.17                    & \textbf{5.2x10$^{-06}$} & \textbf{2.1$^{-05}$}    & 0.38            &             &        \\
-    N348W/Y211A & \textbf{0.00017} & \textbf{3.9x10$^{-06}$} & \textbf{0.0011} & \textbf{0.021}          & 0.68                    & \textbf{1.6x10$^{-10}$} & \textbf{6.3x10$^{-10}$} & \textbf{0.0047} & 0.063       &        \\
-   vR111A/Y211A & 0.56             & 0.34                    & 0.86            & 0.47                    & \textbf{0.0072}         & \textbf{0.0010}         & \textbf{0.0025}         & 0.77            & 0.24        & \textbf{0.0019}  \\
-    \hline
-    \end{tabular}
-  }
-\end{center}
-\end{table}
-\fi
-
-% \clearpage
-% \begin {table}[H]\label{tab:
-% \caption{Pairwise differences in interpolated maximum applied force.
-Bolded values are statistically significant at $p<0.05$.
-$\Delta$ Max Force is the mean maximum applied force for the first mutant listed in a row minus that of the second mutant listed.}
-% \begin{center}
-%   \resizebox{14cm}{!} {
-%     \begin{tabular}{l r}
-%     \hline
-%     Mutant Pair                &  $\Delta$ Max Force \\ \hline
-%     N348A-WT                   &               27.33 \\
-%     N348W-WT                   &              -39.87 \\
-%     N348K-WT                   &              -40.64 \\
-%     \textbf{Y211A-WT}          &      \textbf{-87.97} \\
-%     \textbf{Y211D-WT}          &      \textbf{103.44} \\
-%     \textbf{Y211T-WT}          &       \textbf{89.17} \\
-%     vR111A-WT                  &              -17.61 \\
-%     N348A/Y211A-WT             &               -9.04 \\
-%     \textbf{N348W/Y211A-WT}    &     \textbf{-154.79} \\
-%     vR111A/Y211A-WT            &                8.71 \\
-%     \textbf{N348W-N348A}       &      \textbf{-67.20} \\
-%     \textbf{N348K-N348A}       &      \textbf{-67.97} \\
-%     \textbf{Y211A-N348A}       &     \textbf{-115.30} \\
-%     \textbf{Y211D-N348A}       &       \textbf{76.12} \\
-%     \textbf{Y211T-N348A}       &       \textbf{61.85} \\
-%     vR111A-N348A               &              -44.94 \\
-%     N348A/Y211A-N348A          &              -36.37 \\
-%     \textbf{N348W/Y211A-N348A} &     \textbf{-182.12} \\
-%     vR111A/Y211A-N348A         &              -18.62 \\
-%     N348K-N348W                &               -0.77 \\
-%     Y211A-N348W                &              -48.10 \\
-%     \textbf{Y211D-N348W}       &      \textbf{143.31} \\
-%     \textbf{Y211T-N348W}       &      \textbf{129.04} \\
-%     vR111A-N348W               &               22.26 \\
-%     N348A/Y211A-N348W          &               30.83 \\
-%     \textbf{N348W/Y211A-N348W} &     \textbf{-114.92} \\
-%     vR111A/Y211A-N348W         &               48.58 \\
-%     Y211A-N348K                &              -47.33 \\
-%     
-%     \hline
-%     \end{tabular}
-%     
-%     \begin{tabular}{l r}
-%     \hline
-%     Mutant Pair                       &  $\Delta$ Max Force  \\ \hline
-%     \textbf{Y211D-N348K}              &      \textbf{144.08} \\
-%     \textbf{Y211T-N348K}              &      \textbf{129.81} \\
-%     vR111A-N348K                      &               23.03  \\
-%     N348A/Y211A-N348K                 &               31.60  \\
-%     \textbf{N348W/Y211A-N348K}        &     \textbf{-114.15} \\
-%     vR111A/Y211A-N348K                &               49.35  \\
-%     \textbf{Y211D-Y211A}              &      \textbf{191.42} \\
-%     \textbf{Y211T-Y211A}              &      \textbf{177.15} \\
-%     \textbf{vR111A-Y211A}             &       \textbf{70.36} \\
-%     \textbf{N348A/Y211A-Y211A}        &       \textbf{78.93} \\
-%     \textbf{N348W/Y211A-Y211A}        &      \textbf{-66.82} \\
-%     \textbf{vR111A/Y211A-Y211A}       &       \textbf{96.68} \\
-%     Y211T-Y211D                       &              -14.27  \\
-%     \textbf{vR111A-Y211D}             &     \textbf{-121.05} \\
-%     \textbf{N348A/Y211A-Y211D}        &     \textbf{-112.48} \\
-%     \textbf{N348W/Y211A-Y211D}        &     \textbf{-258.23} \\
-%     \textbf{vR111A/Y211A-Y211D}       &      \textbf{-94.74} \\
-%     \textbf{vR111A-Y211T}             &     \textbf{-106.78} \\
-%     \textbf{N348A/Y211A-Y211T}        &      \textbf{-98.21} \\
-%     \textbf{N348W/Y211A-Y211T}        &     \textbf{-243.96} \\
-%     \textbf{vR111A/Y211A-Y211T}       &      \textbf{-80.46} \\
-%     N348A/Y211A-vR111A                &                8.57  \\
-%     \textbf{N348W/Y211A-vR111A}       &     \textbf{-137.18} \\
-%     vR111A/Y211A-R111A                &               26.31  \\
-%     \textbf{N348W/Y211A-N348A/Y211A}  &     \textbf{-145.75} \\
-%     vR111A/Y211A-N348A/Y211A          &               17.75  \\
-%     \textbf{vR111A/Y211A-N348W/Y211A} &      \textbf{163.50} \\ 
-%                                       &          \\
-%     \hline
-%     \end{tabular}
-%   }
-% \end{center}
-% \end{table}
 

### 2013-07-08 paper.tex
Removed the final paragraph entirely.  Better to end of a paragraph about the method.

@@
 
-We have probed the biochemical effects of mutations at the host-virus interface between MACV and its human receptor.
-We applied an approach that imparts forces in an all-atom molecular dynamics (MD) simulation to drive the system through equilibrium.
-Our work has shown that with relatively modest computing resources it is possible to use MD simulations to investigate point mutations in the GP1/hTfR1 interface.
-In most cases, a double mutant was easily sufficient to resolve a difference in relative affinity.
-Moreover, relative to the WT complex one point mutant was capable of producing a large decrease in binding affinity.
-We applied our technique to the WT complex and 10 mutations in the GP1/hTfR1 interface.
-Our data has provided some direction for further biochemical investigation in this system.
-A polar residue is absolutely critical at position 211 in hTfR1.
-Any mutation to a nonpolar residue will probably lead to a rapid viral compensatory response.
-Also, although rodent populations may produce further mutations at Asn348 in hTfR1, unless a mutation is positively charged or very large it is unlikely to result in a large disruption to receptor binding.
-Finally, our analysis offers a new means to systemically investigate host-virus interactions through computational molecular dynamics.
-
 \section*{Acknowledgments}

### 2013-07-09 smd_pro/data_wt_old/distance/first_analysis.R
Adding all of the data and the analysis scripts.

@@ -1,0 +1,6 @@
+setwd("~/Desktop/data")
+
+for(i in 0:50) {
+  distance <- read.table(paste("distances/distance_", i, ".dat", sep=""))[2]
+  #force <- read.table(paste("forces/force_", i, ".dat", sep=""))[2]
+}

### 2013-07-09 smd_pro/old_analysis.R

@@ -1,0 +1,686 @@
+setwd('~/Desktop/MD_Misc/smd_pro/')
+
+diff.plot <- function(diff.df) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=22
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
+                            axis.line=element_line(),
+                            panel.grid.minor=element_blank(),
+                            panel.grid.major=element_blank(),
+                            panel.background=element_blank(),
+                            panel.border=element_blank(),
+                            axis.line=element_line())
+  
+  g <- ggplot(rbind(diff.df[diff.df$id=='awt', ], diff.df[diff.df$id=='jnw_ya', ], diff.df[diff.df$id=='fyd', ]), 
+              aes(y2, fill = id)) + 
+    geom_density(alpha = 0.2, size=1) +  #+ 
+    theme(strip.background=element_blank()) +
+    ylab('Density') +
+    xlab('Interpolated Max Force (pN)') +
+    scale_x_continuous(breaks=seq(100, 1300, 200), limits=c(50, 1350)) +
+    theme(panel.border=element_blank(), axis.line=element_line()) +
+    theme(axis.title.x = element_text(size=the_pointsize, vjust=-.2)) +
+    theme(axis.title.y = element_text(angle=90, size=the_pointsize, vjust=.3)) +
+    theme(axis.line = element_line(colour = 'black', size = 1)) +
+    theme(axis.ticks = element_line(colour = 'black', size = 1)) +
+    theme(plot.margin=unit(c(1, 1, 1, 1), "lines")) +
+    theme(axis.ticks.margin = unit(0.15, "cm")) +
+    theme(legend.position = "none")
+  
+  ggsave(g, file='~/Google Drive/Documents/Research/SMD_manuscript/MACV_SMD/figures/test_sensitivity.pdf', width=7, height=5)
+}
+
+
+smd.boxplot <- function(boxplot.df) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=20
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
[…]

### 2013-07-09 smd_pro/smd_analysis.R

@@ -1,0 +1,222 @@
+setwd('~/Google Drive/Documents/Research/SMD_manuscript/MACV_SMD/smd_pro/')
+
+diff.plot <- function(diff.df) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=22
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
+                            axis.line=element_line(),
+                            panel.grid.minor=element_blank(),
+                            panel.grid.major=element_blank(),
+                            panel.background=element_blank(),
+                            panel.border=element_blank(),
+                            axis.line=element_line())
+  
+  g <- ggplot(rbind(diff.df[diff.df$id=='awt', ], diff.df[diff.df$id=='jnw_ya', ], diff.df[diff.df$id=='fyd', ]), 
+              aes(y2, fill = id)) + 
+    geom_density(alpha = 0.2, size=1) +  #+ 
+    theme(strip.background=element_blank()) +
+    ylab('Density') +
+    xlab('Interpolated Max Force (pN)') +
+    scale_x_continuous(breaks=seq(100, 1300, 200), limits=c(50, 1350)) +
+    theme(panel.border=element_blank(), axis.line=element_line()) +
+    theme(axis.title.x = element_text(size=the_pointsize, vjust=-.2)) +
+    theme(axis.title.y = element_text(angle=90, size=the_pointsize, vjust=.3)) +
+    theme(axis.line = element_line(colour = 'black', size = 1)) +
+    theme(axis.ticks = element_line(colour = 'black', size = 1)) +
+    theme(plot.margin=unit(c(1, 1, 1, 1), "lines")) +
+    theme(axis.ticks.margin = unit(0.15, "cm")) +
+    theme(legend.position = "none")
+  
+  ggsave(g, file='~/Google Drive/Documents/Research/SMD_manuscript/MACV_SMD/figures/test_sensitivity.pdf', width=7, height=5)
+}
+
+
+smd.boxplot <- function(boxplot.df) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=20
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
[…]

### 2013-07-09 paper.tex
Updated everything to current text.
Changes to manuscript.

@@
 \begin{abstract}
-
 Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases.
@@
 By contrast, first principles methods offer the chance to forgo training.
-However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional protein model to another with no intermediate steps \citep{Gilson1997,Lu2004}.
+However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional model to another with no intermediate steps \citep{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
 As a result, the accuracy of the measurement suffers as the two proteins become increasingly dissimilar; therefore, these techniques are largely limited to point mutant comparisons.
@@
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
-Then, we measure the force required for dissociation.
+Then, we measure the force required for dissociation \citep{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
 As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
@@
 In every way possible the subsequent steering simulation had the same parameters as the equilibration simulation.
-Thus, the steered molecular dynamics (SMD) \citep{Is2001B} simulation was a typical restart with slightly different settings.
+Thus, the steered molecular dynamics (SMD) \citep{Is2001A,Is2001B} simulation was a typical restart with slightly different settings.
 Periodic boundary conditions were incorporated as part of the system and PME was again used to approximate full system electrostatics.
@@
 One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function.
-Seen in Table~\ref{tab:summary_mutant}), we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine.
+Seen in Table~\ref{tab:summary_mutants}), we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine.
 Several mutations caused significant relative affinity changes.
@@
 To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala.
-This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:MAF_pairwises}).
+This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:MAF_pairwise}).
 This result shows that the two sites do indeed interact, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction could lead to comparable binding affinity.
@@
 
-This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) to A.D.E., the National Institutes of Health (R01-GM088344) to C.O.W., and the National Institutes of Health (R01-GM093086) to S.L.S.
+This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) and the Welch Foundation (F-1654) to A.D.E., the National Institutes of Health (R01-GM088344) to C.O.W., and the National Institutes of Health (R01-GM093086) to S.L.S.
 

### 2013-07-10 smd_pro/smd_analysis.R
Commiting changes.

@@ -213,10 +213,10 @@ df <- data.frame(id=c(rep('awt', length(max.wt[, 1])), rep('eya', length(max.ya[
 #                                'Y211D', 'Y211T', 'Y211A', 'N348W/Y211A'))
 # smd.boxplot(bp.df)
 # 
-fit <- aov(y1/mean(y1) + y2/mean(y2) ~ id, data = df)
+fit <- aov(y1 + y2 ~ id, data = df)
 print(TukeyHSD(fit))
 
 
-print(pairwise.t.test(df$y1/mean(df$y1) + df$y2/mean(df$y2), df$id, p.adjust.method="fdr"))
+#print(pairwise.t.test(scale(df$y1) + scale(df$y2), df$id, p.adjust.method="fdr"))
 
 

### 2013-07-16 paper.tex
Small change in discussion.

@@
 One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function.
-Seen in Table~\ref{tab:summary_mutants}), we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine.
+Seen in Table~\ref{tab:summary_mutants}, we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine.
 Several mutations caused significant relative affinity changes.
@@
 We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}.
-Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity.
-(Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).
+Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).
 
@@
 As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}).
-In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp.
-(For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT.
-See Table~\ref{tab:summary_statistics}.)  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder.
+In addition, neither the Asn348Lys nor the Asn348Trp mutation showed a significant difference from WT.
+For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT (See Table~\ref{tab:summary_statistics}).
+On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder.
 Moreover, Asn348Trp showed nearly identical results to Asn348Lys.
@@
 By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application.
-Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
+Traditionally this technique has been either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
 Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}.
@@
 However, the computational data suggest that the reason and nature of the effects at these two sites are very different.
-Congruent with entry data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding.
-Second, no individual point mutation at Asn348 showed a significantly different affinity from the WT complex.
-However, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference.
+At position 211, mutations to non-polar residues causes a large change in binding.
+This is congruent with what is known from viral entry data \citep{Rad2008,Rad20111}.
+By contrast, mutations at position 348 need only be small to maintain WT binding.
+The ability to hydrogen bond appears to be insignificant.
+This can be inferred from the fact that Tyr211Ala paired with large (Trp) and positively charged (Lys) substitutions at position 348 results in a larger than expected synergistic difference.
 That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually.

### 2013-07-17 paper.tex
last minor edits.

@@
 The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain.
-For GP1, several extended loops in mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
-Experimental alanine-scanning and whole cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.
+For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
+Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.
 
@@
 The starting center-of-mass distance was set to zero and the distance was re-computed at each time step relative to the starting distance.
-The force is output directly to the standard log file.
+The force was output directly to the standard log file.
 
@@
 The complexity of correctly parameterizing diverse sugar moieties is outside of the scope of this paper.
-Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the effect of those contacts \textit{in vivo} is still unresolved \citep{Abraham2010}.
+Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \citep{Abraham2010}.
 
@@
 In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}.
-Finally, an alanine mutation at site 111 in GP1 has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}.
+Finally, an alanine mutation at site 111 in GP1 (mutation vArg111Ala) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}.
 For notation purposes, the viral site is always referred to with a preceding `v'.
@@
 \subsection*{Steered molecular dynamics simulations}
+
+We proceeded to analyze this system using molecular dynamics.
 To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions.
@@
 We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:MAF_pairwise}), while the other two did not (Table~\ref{tab:MAF_pairwise}).
-When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:MAF_pairwise} and ~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}).
+When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:MAF_pairwise} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}).
 In general, however, maximum applied force seemed to be the more sensitive statistic than AUC.
@@
 One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function.
-Seen in Table~\ref{tab:summary_mutants}, we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine.
+As shown in Table~\ref{tab:summary_mutants}, we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine.
 Several mutations caused significant relative affinity changes.
 In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause similar changes in binding.
-Then, we compared the size of those differences to single mutants (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}).
+Then, we compared the size of those differences to single mutants (Figure~\ref{fig:sensitivity} and~\ref{fig:mutant_comparison}).
 
@@
 We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}.
-Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).
+Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity (Figures~\ref{fig:sensitivity} and~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).
 
 We also simulated several point mutations at Asn348 in the hTfR1.
-As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}).
+As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}).
 In addition, neither the Asn348Lys nor the Asn348Trp mutation showed a significant difference from WT.
@@
 To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding.
-As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study.
+As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $, Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study.
 Further, according to maximum applied force (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}).
@@
 While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, change in binding at that site may actually be caused by size and charge restriction.
-We also found that a negatively polar residue at site 211 in hTfR1 is absolutely critical for a tight binding interaction.
+We also found that a negatively polar residue at site 211 in hTfR1 seem critical for a tight binding interaction.
 Any non-polar mutation at Tyr211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection.
@@
 However, the computational data suggest that the reason and nature of the effects at these two sites are very different.
-At position 211, mutations to non-polar residues causes a large change in binding.
+At position 211, mutations to non-polar residues cause a large change in binding.
 This is congruent with what is known from viral entry data \citep{Rad2008,Rad20111}.
@@
 
-We would like to emphasize here that we cannot expect perfect agreement.
+We would like to emphasize here that we cannot expect perfect agreement between our simulations and the available experimental data.
 While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method.
@@
 More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph).
-Second, our SMD method is fundamentally a measurement tool, and as a result we are dramatically limited by the accuracy of our starting structure.
+Second, our SMD method is fundamentally limited by the accuracy of our starting structure.
 Third, the available experimental data for the GP1/hTfR1 system were generally obtained from entry assays or whole-cell binding assays rather than molecular binding assays.
@@
 \centerline{\includegraphics[width=6in]{figures/mutant_comparison_boxplot.pdf}}
-\caption{\label{fig:mutant_comparison}\title{Distribution of interpolated maximum force for all complexes tested.} The statistically significant difference relative to the WT complex are indicated by a star above the distribution.}
+\caption{\label{fig:mutant_comparison}\title{Distribution of interpolated maximum force for all complexes tested.} Stars above the boxplots indicate statistically significant difference in mean maximum force relative to the WT complex.}
 \end{figure}
@@
 \begin {table}[H]
-\caption{\label{tab:summary_mutants}Summary of information for each mutation tested.
+\caption{\label{tab:summary_mutants}Summary of prior information available for each mutation tested.
 Observed $in~vivo$ refers to mutations that have been observed in rodent populations.

### 2013-07-18 paper.tex
Change format for final submission to JCTC.

@@
-\documentclass[12pt]{article}
-\pdfoutput=1
+\documentclass[journal=jacsat,manuscript=article]{achemso}
 
-\usepackage[margin=2.5cm]{geometry}
+\usepackage{mathptmx} 
 \usepackage{graphicx}
-\usepackage{natbib}
-%correct punctuation for MBE
-\bibpunct[,]{(}{)}{;}{a}{}{,}
-
+\usepackage{lscape}
 \usepackage{amsmath}
-
+\usepackage{verbatim}
 \renewcommand{\bottomfraction}{.9}
@@
 
-\linespread{1.5}
+\title[Predicting mutant binding]{Predicting virus-receptor mutant binding by molecular dynamics simulation}
+
+\author{Austin G. Meyer}
+\email{austin.meyer@utexas.edu}
+\phone{+1 512 971 0123}
+\affiliation{Section of Integrative Biology, Institute for Cellular and Molecular Biology, and Center for Computational Biology and Bioinformatics, The University of Texas at Austin, Austin, TX 78731, USA}
+\altaffiliation{Current address: School of Medicine, Texas Tech University Health Sciences Center, Lubbock, TX, 79430, USA}
+\altaffiliation{Department of Molecular Biosciences, Institute for Cellular and Molecular Biology, The University of Texas at Austin, Austin, TX 78731, USA}
+
+\author{Sara L. Sawyer}
+
+\author{Andrew D. Ellington}
+\affiliation{Department of Molecular Biosciences, Institute for Cellular and Molecular Biology, The University of Texas at Austin, Austin, TX 78731, USA}
+
+\author{Claus O. Wilke}
+\affiliation{Section of Integrative Biology, Institute for Cellular and Molecular Biology, and Center for Computational Biology and Bioinformatics, The University of Texas at Austin, Austin, TX 78731, USA}
+
 \begin{document}
-
-\title{\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}}
-\author{Austin G. Meyer$^{1,2,4*}$, Sara L. Sawyer$^{3}$, Andrew D. Ellington$^{2}$,\\and Claus O. Wilke$^{1}$}
-\date{}
-
-\maketitle
-\noindent
-Address:\\
-$^1$Section of Integrative Biology, Institute for Cellular and Molecular
-Biology, and Center for Computational Biology and Bioinformatics. The University
-of Texas at Austin, Austin, TX 78712, USA.\\
-$^2$Department of Chemistry and Biochemistry, Institute for Cellular and Molecular 
-Biology, The University of Texas at Austin, Austin, TX 78712, USA.\\
-$^3$Section of Molecular Genetics and Microbiology, Institute for Cellular and Molecular 
-Biology, The University of Texas at Austin, Austin, TX 78712, USA.\\
-$^4$School of Medicine, Texas Tech University Health Sciences Center, 
-Lubbock, TX 79430, USA.\\
-
-\bigskip
-\noindent
-$^*$Corresponding author\\
-$\phantom{^*}$Email: austin.meyer@utexas.edu\\
-$\phantom{^*}$Phone: +1 512 971 0123\\
-
-\bigskip
-\noindent
-Manuscript type: research article
-
-\bigskip
-\noindent Keywords: protein--protein interaction, molecular dynamics, arenavirus, machupo
 
@@
 
-\section*{Introduction}
+\section{Introduction}
 
@@
 Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure.
-For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}.
+For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \cite{Grant2011,Kortemme2004}.
 However, such methods calculate the relative contribution to overall energy for each amino acid in a protein.
-They do not probe the interface directly, but rather calculate the total energy with each substitution \citep{Grant2011,Kortemme2004}.
-Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}.
+They do not probe the interface directly, but rather calculate the total energy with each substitution \cite{Grant2011,Kortemme2004}.
+Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \cite{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}.
 Unfortunately, such machine-learning approaches often suffer in novel applications, for which available training sets are small or non-existent.
@@
 By contrast, first principles methods offer the chance to forgo training.
-However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional model to another with no intermediate steps \citep{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
+However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional model to another with no intermediate steps \cite{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
 As a result, the accuracy of the measurement suffers as the two proteins become increasingly dissimilar; therefore, these techniques are largely limited to point mutant comparisons.
@@
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
-Then, we measure the force required for dissociation \citep{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
+Then, we measure the force required for dissociation \cite{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
 As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
 
-We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}.
-Machupo is an ambisense RNA virus of the arenavirus family \citep{Charrel2003}.
-Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}.
-Members of the arenavirus family include the Lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the Guanarito, Junin, and Machupo viruses endemic to rodents in South America \citep{Charrel2003}.
-The South American arenaviruses typically infect humans after rodent contamination and can cause a devastating hemorrhagic fever with high mortality \citep{Charrel2003}.
+We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \cite{Abraham2010,Charrel2003}.
+Machupo is an ambisense RNA virus of the arenavirus family \cite{Charrel2003}.
+Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \cite{Charrel2003}.
+Members of the arenavirus family include the Lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the Guanarito, Junin, and Machupo viruses endemic to rodents in South America \cite{Charrel2003}.
+The South American arenaviruses typically infect humans after rodent contamination and can cause a devastating hemorrhagic fever with high mortality \cite{Charrel2003}.
 
@@
 The hTfR1 protein contains three extracellular domains: two basilar domains and an apical domain.
-The two basilar domains serve most of the transferrin-binding function \citep{Abraham2010,Rad20112}.
+The two basilar domains serve most of the transferrin-binding function \cite{Abraham2010,Rad20112}.
 Viral entry is initiated by GP1 binding to the apical domain of hTfR1.
-Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}.
+Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \cite{Rad20111,Rad20112}.
 The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain.
-For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
-Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.
+For GP1, several extended loops mediate binding to hTfR1 \cite{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \cite{Abraham2010}.
+Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \cite{Rad20111,Rad20112}.
 
@@
 
-\section*{Materials and Methods}
+\section{Materials and Methods}
 
-\subsection*{System Modeling}
+\subsection{System Modeling}
 
-We used the protein visualization software PyMOL \citep{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1.
+We used the protein visualization software PyMOL \cite{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1.
 No residues were removed from the viral protein.
@@
 
-After system reduction, the Visual Molecular Dynamics (VMD) \citep{Humphrey1996} package along with its system of back-ends was used for all subsequent modeling.
+After system reduction, the Visual Molecular Dynamics (VMD) \cite{Humphrey1996} package along with its system of back-ends was used for all subsequent modeling.
 The Orient add-on package allowed us to rotate the system axis such that the direction of steering was oriented directly down the z-axis.
@@
 
-\subsection*{Equilibration}
+\subsection{Equilibration}
 
-NAMD was used for all simulations in this study \citep{Phillips2005}.
+NAMD was used for all simulations in this study \cite{Phillips2005}.
 In addition to the modeled system, for equilibration we generated a configuration file that fixed the $\alpha$-carbon backbone.
@@
 
-We used the Charmm27 \citep{Brooks1983} all-atom force field.
+We used the Charmm27 \cite{Brooks1983} all-atom force field.
 The initial system temperature was set to 310K.
@@
 
-\subsection*{Steered Molecular Dynamics}
+\subsection{Steered Molecular Dynamics}
 
@@
 In every way possible the subsequent steering simulation had the same parameters as the equilibration simulation.
-Thus, the steered molecular dynamics (SMD) \citep{Is2001A,Is2001B} simulation was a typical restart with slightly different settings.
+Thus, the steered molecular dynamics (SMD) \cite{Is2001A,Is2001B} simulation was a typical restart with slightly different settings.
 Periodic boundary conditions were incorporated as part of the system and PME was again used to approximate full system electrostatics.
@@
 
-\subsection*{Post-processing}
+\subsection{Post-processing}
 
-The python packages MDAnalysis \citep{Agrawal2011} and ProDy \citep{Bakan2011} were both used at various points in post-processing.
+The python packages MDAnalysis \cite{Agrawal2011} and ProDy \cite{Bakan2011} were both used at various points in post-processing.
 The molecular trajectory (comprising the atomic coordinates per time) was parsed to compute the center-of-mass for each of the two complexes.
@@
 We used the pairwise.t.test function with the FDR p-value correction to test for significance.
-The ggplot \citep{ggplot} package was used to generate most of the figures.
+The ggplot \cite{ggplot} package was used to generate most of the figures.
 
-\section*{Results}
+\section{Results}
 
-\subsection*{The GP1/hTfR1 system}
+\subsection{The GP1/hTfR1 system}
 The GP1/hTfR1 interface marks a particularly important and useful test system.
@@
 The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals).
-In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \citep{Grant2011,Kortemme2004} only marginally useful.
+In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \cite{Grant2011,Kortemme2004} only marginally useful.
 The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.
 
-For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \citep{Abraham2010}.
+For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \cite{Abraham2010}.
 The apical domain of hTfR1 interacts directly with GP1 while the other two domains are closer to the cell membrane and have essentially no interaction with GP1.
@@
 The complexity of correctly parameterizing diverse sugar moieties is outside of the scope of this paper.
-Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \citep{Abraham2010}.
+Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \cite{Abraham2010}.
 
 In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}).
-Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
+Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \cite{Rad2008,Rad20111,Rad20112}.
 Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1.
-The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}.
-In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1.
-In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}.
-Finally, an alanine mutation at site 111 in GP1 (mutation vArg111Ala) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}.
+The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \cite{Abraham2010}.
+In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \cite{Rad2008,Abraham2010} to GP1.
+In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \cite{Rad2008,Abraham2010}.
+Finally, an alanine mutation at site 111 in GP1 (mutation vArg111Ala) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \cite{Rad20112}.
 For notation purposes, the viral site is always referred to with a preceding `v'.
 
-\subsection*{Steered molecular dynamics simulations}
+\subsection{Steered molecular dynamics simulations}
 
@@
 Next, we carried out successive rounds of energy minimization with a fixed $\alpha$-carbon backbone, followed by heating, and backbone release.
-Last, we performed a short round of all-atom molecular dynamics with minimal restraints \citep{Cuendet2008}.
+Last, we performed a short round of all-atom molecular dynamics with minimal restraints \cite{Cuendet2008}.
 We used the final state from each equilibrated system to restart another MD simulation.
 For SMD, at the centroid of several atoms a force was applied to a single point in GP1.
-By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
+By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \cite{Cuendet2008,Cuendet2011}.
 A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}.
@@
 Their summary statistics are reported in Table~\ref{tab:summary_statistics}.
-As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.
+As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \cite{Is2001A,Is2001B}.
 
 Before systematically applying SMD to the GP1/hTfR1 interaction, we needed to ensure the method was sufficiently sensitive to distinguish between relatively minor point mutations.
-While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
+While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \cite{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
 For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.
@@
 
-\subsection*{Comparative analysis of the GP1/hTfR1 interface}
+\subsection{Comparative analysis of the GP1/hTfR1 interface}
 Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction.
@@
 In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding.
-We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}.
+We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \cite{Rad2008,Rad20111,Rad20112}.
 Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity (Figures~\ref{fig:sensitivity} and~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).
@@
 
-\section*{Discussion}
+\section{Discussion}
 
@@
 By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application.
-Traditionally this technique has been either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
-Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}.
-And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}.
+Traditionally this technique has been either applied to compute equilibrium quantities via a non-equilibrium approximation \cite{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \cite{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \cite{Dixit2001}.
+Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \cite{Cuendet2011,Giorgino2011}.
+And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \cite{Sulkowska2010}.
 Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions.
@@
 
-Our findings may rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}.
+Our findings may rationalize several effects observed in both infectivity data and rodent populations \cite{Rad2008,Rad20111}.
 First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding.
@@
 At position 211, mutations to non-polar residues cause a large change in binding.
-This is congruent with what is known from viral entry data \citep{Rad2008,Rad20111}.
+This is congruent with what is known from viral entry data \cite{Rad2008,Rad20111}.
 By contrast, mutations at position 348 need only be small to maintain WT binding.
@@
 That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually.
-Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
+Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \cite{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
 Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.
@@
 by determining the optimal pulling speed.
-Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \citep{Shaw2012}.
-Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \citep{Shaw2011}.
+Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \cite{Shaw2012}.
+Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \cite{Shaw2011}.
 With improved resources, we could investigate thousands of individual permutations in the GP1/hTfR1 binding interface.
-In addition, with additional compute time it would be possible to incorporate equilibrium sampling approaches \citep{Buch2011} or use brute force equilibrium approaches \citep{Giorgino2012} to improve resolution.
+In addition, with additional compute time it would be possible to incorporate equilibrium sampling approaches \cite{Buch2011} or use brute force equilibrium approaches \cite{Giorgino2012} to improve resolution.
 Going forward, we believe that all-atom molecular dynamics simulations can become an increasingly viable approach for protein interaction studies.
 
-\section*{Acknowledgments}
+\section{Author Information}
+
+\subsection{Corresponding Author}
+
+austin.meyer@utexas.edu
+
+\subsection{Notes}
+
+The authors declare no competing financial interest.
+
+\begin{acknowledgement}
 This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu).
@@
 This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) and the Welch Foundation (F-1654) to A.D.E., the National Institutes of Health (R01-GM088344) to C.O.W., and the National Institutes of Health (R01-GM093086) to S.L.S.
+\end{acknowledgement}
 
-\bibliographystyle{MBE}
 \bibliography{paper}
@@
   \resizebox{11cm}{!} {
-    \begin{tabular}{l r r r r}
+    \begin{tabular}{l c c c c}
     \hline

### 2013-07-19 paper.tex
Submitted version to JCTC.

@@
     \hline
-      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\
+      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\ \hline
             WT &                Yes &         Normal Entry \\

### 2013-07-19 paper.tex
Fixed the two formatting mistakes from the JCTC submission.

@@
 \affiliation{Section of Integrative Biology, Institute for Cellular and Molecular Biology, and Center for Computational Biology and Bioinformatics, The University of Texas at Austin, Austin, TX 78731, USA}
-\altaffiliation{Current address: School of Medicine, Texas Tech University Health Sciences Center, Lubbock, TX, 79430, USA}
-\altaffiliation{Department of Molecular Biosciences, Institute for Cellular and Molecular Biology, The University of Texas at Austin, Austin, TX 78731, USA}
+\alsoaffiliation{School of Medicine, Texas Tech University Health Sciences Center, Lubbock, TX, 79430, USA}
+\alsoaffiliation{Department of Molecular Biosciences, Institute for Cellular and Molecular Biology, The University of Texas at Austin, Austin, TX 78731, USA}
 
@@
 
-\begin{acknowledgement}
-This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu).
+\section{Acknowledgements}
+This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock \\ (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu).
 The authors acknowledge Bryan Sutton for opening access to the Hrothgar cluster.
@@
 This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) and the Welch Foundation (F-1654) to A.D.E., the National Institutes of Health (R01-GM088344) to C.O.W., and the National Institutes of Health (R01-GM093086) to S.L.S.
-\end{acknowledgement}
 

### 2013-12-03 full_paper.tex
Adding everything necessary for submissiong that is in my local repo. This includes data and the new PeerJ manuscript.

--- 2013-07-19 paper.tex
@@
-\documentclass[journal=jacsat,manuscript=article]{achemso}
+\documentclass[12pt]{article}
 
-\usepackage{mathptmx} 
+\usepackage[margin=2.5cm]{geometry}
 \usepackage{graphicx}
-\usepackage{lscape}
+\usepackage{natbib}
+\usepackage{lineno}
+
+\usepackage{times}
+%correct punctuation for MBE
+\bibpunct[,]{(}{)}{;}{a}{}{,}
+
 \usepackage{amsmath}
-\usepackage{verbatim}
+
 \renewcommand{\bottomfraction}{.9}
@@
 
-\title[Predicting mutant binding]{Predicting virus-receptor mutant binding by molecular dynamics simulation}
-
-\author{Austin G. Meyer}
-\email{austin.meyer@utexas.edu}
-\phone{+1 512 971 0123}
-\affiliation{Section of Integrative Biology, Institute for Cellular and Molecular Biology, and Center for Computational Biology and Bioinformatics, The University of Texas at Austin, Austin, TX 78731, USA}
-\alsoaffiliation{School of Medicine, Texas Tech University Health Sciences Center, Lubbock, TX, 79430, USA}
-\alsoaffiliation{Department of Molecular Biosciences, Institute for Cellular and Molecular Biology, The University of Texas at Austin, Austin, TX 78731, USA}
-
-\author{Sara L. Sawyer}
-
-\author{Andrew D. Ellington}
-\affiliation{Department of Molecular Biosciences, Institute for Cellular and Molecular Biology, The University of Texas at Austin, Austin, TX 78731, USA}
-
-\author{Claus O. Wilke}
-\affiliation{Section of Integrative Biology, Institute for Cellular and Molecular Biology, and Center for Computational Biology and Bioinformatics, The University of Texas at Austin, Austin, TX 78731, USA}
-
+\linespread{1.5}
 \begin{document}
 
+\title{\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}}
+\author{Austin G. Meyer$^{1,2,3*}$, Sara L. Sawyer$^{2}$, Andrew D. Ellington$^{2}$,\\and Claus O. Wilke$^{1}$}
+\date{}
+
+\maketitle
+\noindent
+Address:\\
+$^1$Department of Integrative Biology, Institute for Cellular and Molecular
+Biology, and Center for\\Computational Biology and Bioinformatics. The University
+of Texas at Austin, Austin, TX 78712, USA.\\
+$^2$Department of Molecular Biosciences, Institute for Cellular and Molecular 
+Biology, The University of Texas at Austin, Austin, TX 78712, USA.\\
+$^3$School of Medicine, Texas Tech University Health Sciences Center, 
+Lubbock, TX 79430, USA.\\
+
+\bigskip
+\noindent
+$^*$Corresponding author\\
+$\phantom{^*}$Email: austin.meyer@utexas.edu\\
+$\phantom{^*}$Phone: +1 512 971 0123\\
+
+\bigskip
+\noindent
+Manuscript type: research article
+
+\bigskip
+\noindent Keywords: protein--protein interaction, molecular dynamics, arenavirus, machupo
+
+\linenumbers
 \newpage
-
 \begin{abstract}
@@
 Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure.
-For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \cite{Grant2011,Kortemme2004}.
+For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}.
 However, such methods calculate the relative contribution to overall energy for each amino acid in a protein.
-They do not probe the interface directly, but rather calculate the total energy with each substitution \cite{Grant2011,Kortemme2004}.
-Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \cite{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}.
+They do not probe the interface directly, but rather calculate the total energy with each substitution \citep{Grant2011,Kortemme2004}.
+Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}.
 Unfortunately, such machine-learning approaches often suffer in novel applications, for which available training sets are small or non-existent.
@@
 By contrast, first principles methods offer the chance to forgo training.
-However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional model to another with no intermediate steps \cite{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
+However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional model to another with no intermediate steps \citep{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
 As a result, the accuracy of the measurement suffers as the two proteins become increasingly dissimilar; therefore, these techniques are largely limited to point mutant comparisons.
@@
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
-Then, we measure the force required for dissociation \cite{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
+Then, we measure the force required for dissociation \citep{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
 As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
 
-We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \cite{Abraham2010,Charrel2003}.
-Machupo is an ambisense RNA virus of the arenavirus family \cite{Charrel2003}.
-Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \cite{Charrel2003}.
-Members of the arenavirus family include the Lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the Guanarito, Junin, and Machupo viruses endemic to rodents in South America \cite{Charrel2003}.
-The South American arenaviruses typically infect humans after rodent contamination and can cause a devastating hemorrhagic fever with high mortality \cite{Charrel2003}.
+We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}.
+Machupo is an ambisense RNA virus of the arenavirus family \citep{Charrel2003}.
+Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}.
+Members of the arenavirus family include the Lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the Guanarito, Junin, and Machupo viruses endemic to rodents in South America \citep{Charrel2003}.
+The South American arenaviruses typically infect humans after rodent contamination and can cause a devastating hemorrhagic fever with high mortality \citep{Charrel2003}.
 
@@
 The hTfR1 protein contains three extracellular domains: two basilar domains and an apical domain.
-The two basilar domains serve most of the transferrin-binding function \cite{Abraham2010,Rad20112}.
+The two basilar domains serve most of the transferrin-binding function \citep{Abraham2010,Rad20112}.
 Viral entry is initiated by GP1 binding to the apical domain of hTfR1.
-Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \cite{Rad20111,Rad20112}.
+Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}.
 The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain.
-For GP1, several extended loops mediate binding to hTfR1 \cite{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \cite{Abraham2010}.
-Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \cite{Rad20111,Rad20112}.
+For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
+Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.
 
@@
 
-We used the protein visualization software PyMOL \cite{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1.
+We used the protein visualization software PyMOL \citep{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1.
 No residues were removed from the viral protein.
@@
 
-After system reduction, the Visual Molecular Dynamics (VMD) \cite{Humphrey1996} package along with its system of back-ends was used for all subsequent modeling.
+After system reduction, the Visual Molecular Dynamics (VMD) \citep{Humphrey1996} package along with its system of back-ends was used for all subsequent modeling.
 The Orient add-on package allowed us to rotate the system axis such that the direction of steering was oriented directly down the z-axis.
@@
 
-NAMD was used for all simulations in this study \cite{Phillips2005}.
+NAMD was used for all simulations in this study \citep{Phillips2005}.
 In addition to the modeled system, for equilibration we generated a configuration file that fixed the $\alpha$-carbon backbone.
@@
 
-We used the Charmm27 \cite{Brooks1983} all-atom force field.
+We used the Charmm27 \citep{Brooks1983} all-atom force field.
 The initial system temperature was set to 310K.
@@
 In every way possible the subsequent steering simulation had the same parameters as the equilibration simulation.
-Thus, the steered molecular dynamics (SMD) \cite{Is2001A,Is2001B} simulation was a typical restart with slightly different settings.
+Thus, the steered molecular dynamics (SMD) \citep{Is2001A,Is2001B} simulation was a typical restart with slightly different settings.
 Periodic boundary conditions were incorporated as part of the system and PME was again used to approximate full system electrostatics.
@@
 
-The python packages MDAnalysis \cite{Agrawal2011} and ProDy \cite{Bakan2011} were both used at various points in post-processing.
+The python packages MDAnalysis \citep{Agrawal2011} and ProDy \citep{Bakan2011} were both used at various points in post-processing.
 The molecular trajectory (comprising the atomic coordinates per time) was parsed to compute the center-of-mass for each of the two complexes.
@@
 We used the pairwise.t.test function with the FDR p-value correction to test for significance.
-The ggplot \cite{ggplot} package was used to generate most of the figures.
+The ggplot \citep{ggplot} package was used to generate most of the figures.
 
@@
 The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals).
-In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \cite{Grant2011,Kortemme2004} only marginally useful.
+In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \citep{Grant2011,Kortemme2004} only marginally useful.
 The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.
 
-For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \cite{Abraham2010}.
+For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \citep{Abraham2010}.
 The apical domain of hTfR1 interacts directly with GP1 while the other two domains are closer to the cell membrane and have essentially no interaction with GP1.
@@
 The complexity of correctly parameterizing diverse sugar moieties is outside of the scope of this paper.
-Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \cite{Abraham2010}.
+Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \citep{Abraham2010}.
 
 In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}).
-Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \cite{Rad2008,Rad20111,Rad20112}.
+Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
 Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1.
-The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \cite{Abraham2010}.
-In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \cite{Rad2008,Abraham2010} to GP1.
-In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \cite{Rad2008,Abraham2010}.
-Finally, an alanine mutation at site 111 in GP1 (mutation vArg111Ala) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \cite{Rad20112}.
+The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}.
+In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1.
+In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}.
+Finally, an alanine mutation at site 111 in GP1 (mutation vArg111Ala) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}.
 For notation purposes, the viral site is always referred to with a preceding `v'.
@@
 Next, we carried out successive rounds of energy minimization with a fixed $\alpha$-carbon backbone, followed by heating, and backbone release.
-Last, we performed a short round of all-atom molecular dynamics with minimal restraints \cite{Cuendet2008}.
+Last, we performed a short round of all-atom molecular dynamics with minimal restraints \citep{Cuendet2008}.
 We used the final state from each equilibrated system to restart another MD simulation.
 For SMD, at the centroid of several atoms a force was applied to a single point in GP1.
-By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \cite{Cuendet2008,Cuendet2011}.
+By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
 A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}.
@@
 Their summary statistics are reported in Table~\ref{tab:summary_statistics}.
-As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \cite{Is2001A,Is2001B}.
+As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.
 
 Before systematically applying SMD to the GP1/hTfR1 interaction, we needed to ensure the method was sufficiently sensitive to distinguish between relatively minor point mutations.
-While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \cite{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
+While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
 For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.
@@
 In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding.
-We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \cite{Rad2008,Rad20111,Rad20112}.
+We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}.
 Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity (Figures~\ref{fig:sensitivity} and~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).
@@
 By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application.
-Traditionally this technique has been either applied to compute equilibrium quantities via a non-equilibrium approximation \cite{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \cite{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \cite{Dixit2001}.
-Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \cite{Cuendet2011,Giorgino2011}.
-And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \cite{Sulkowska2010}.
+Traditionally this technique has been either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
+Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}.
 Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions.
 In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application.
-Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanisms.
 Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation.
@@
 
-Our findings may rationalize several effects observed in both infectivity data and rodent populations \cite{Rad2008,Rad20111}.
+Our findings may rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}.
 First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding.
@@
 At position 211, mutations to non-polar residues cause a large change in binding.
-This is congruent with what is known from viral entry data \cite{Rad2008,Rad20111}.
+This is congruent with what is known from viral entry data \citep{Rad2008,Rad20111}.
 By contrast, mutations at position 348 need only be small to maintain WT binding.
@@
 That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually.
-Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \cite{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
+Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
 Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.
@@
 Further, since our simulations start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) or relative orientation of the two proteins would be underestimated by our method.
-Finally, it is tempting to think only decreased binding affinity should correlate strongly with diminished infectivity.
-However, dramatically increased binding affinity could also disrupt infectivity by prohibiting cellular membrane detachment after viral invasion.
-The virus would be prevented from reaching its replication compartment (which is the cytosol in the case of Machupo) and could not reproduce.
 
@@
 by determining the optimal pulling speed.
-Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \cite{Shaw2012}.
-Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \cite{Shaw2011}.
+Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \citep{Shaw2012}.
+Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \citep{Shaw2011}.
 With improved resources, we could investigate thousands of individual permutations in the GP1/hTfR1 binding interface.
-In addition, with additional compute time it would be possible to incorporate equilibrium sampling approaches \cite{Buch2011} or use brute force equilibrium approaches \cite{Giorgino2012} to improve resolution.
-Going forward, we believe that all-atom molecular dynamics simulations can become an increasingly viable approach for protein interaction studies.
+In addition, with additional compute time it would be possible to incorporate equilibrium sampling approaches \citep{Buch2011} or use brute force equilibrium approaches \citep{Giorgino2012} to improve resolution.
 
-\section{Author Information}
-
-\subsection{Corresponding Author}
-
-austin.meyer@utexas.edu
+For future studies, although our approach offers the simplicity of not requiring prior knowledge about a system of interest (other than a bound model), all atom, unrestrained SMD is likely not the best approach for most relative affinity calculations.
+To ensure one's results are independent of the dissociation path one selects would require computing the work of separation for all likely paths.
+Such an approach eventually requires using the Jarzynski inequality \citep{Jar1997} to establish a lower limit for binding energy and would quickly become computationally inefficient for evaluating a large number of mutations in most systems.
+By contrast, another all atom approach that makes calculations via short simulations of spatially restrained complexes has proven capable of generating relatively accurate binding affinities with less compute time than is required from our steering strategy \citep{Gumbart2013b}.
 
@@
 
+\bibliographystyle{MBE}
 \bibliography{paper}
@@
 \centerline{\includegraphics[width=6in]{figures/mutant_comparison_boxplot.pdf}}
-\caption{\label{fig:mutant_comparison}\title{Distribution of interpolated maximum force for all complexes tested.} Stars above the boxplots indicate statistically significant difference in mean maximum force relative to the WT complex.}
+\caption{\label{fig:mutant_comparison}\title{Distribution of interpolated maximum force for all bound complexes tested.} Stars above the boxplots indicate a statistically significant difference in mean maximum force relative to the WT complex.}
 \end{figure}

### 2013-12-03 smd_pro/smd_analysis.R

@@ -1,4 +1,6 @@
-setwd('~/Google Drive/Documents/Research/SMD_manuscript/MACV_SMD/smd_pro/')
+rm(list = ls())
+
+setwd('~/Google Drive/Documents/Biochem/Research/SMD_manuscript/MACV_SMD/smd_pro/')
 
 diff.plot <- function(diff.df) {
   require(ggplot2)
@@ -31,7 +33,7 @@ diff.plot <- function(diff.df) {
     theme(axis.ticks.margin = unit(0.15, "cm")) +
     theme(legend.position = "none")
   
-  ggsave(g, file='~/Google Drive/Documents/Research/SMD_manuscript/MACV_SMD/figures/test_sensitivity.pdf', width=7, height=5)
+  ggsave(g, file='../figures/test_sensitivity.pdf', width=7, height=5)
 }
 
 
@@ -50,14 +52,14 @@ smd.boxplot <- function(boxplot.df) {
                             panel.border=element_blank(),
                             axis.line=element_line())
   
-  g <- ggplot(boxplot.df, aes(factor(id), y2)) + geom_boxplot()
+  g <- ggplot(boxplot.df, aes(factor(id), y1)) + geom_boxplot()
   g <- g + theme(strip.background=element_blank())
-  g <- g + ylab('Max Force')
+  g <- g + ylab('Distribution of Max Force (pN)')
   g <- g + xlab('Mutant')
   g <- g + theme(panel.border=element_blank(), axis.line=element_line())
-  g <- g + theme(axis.title.x = element_text(size=the_pointsize, vjust=0))
+  g <- g + theme(axis.title.x = element_text(size=18, vjust=0))
   g <- g + theme(axis.text.x = element_text(angle=45, vjust = 1.05, hjust = 1, size=12))
-  g <- g + theme(axis.title.y = element_text(angle=90, size=the_pointsize, vjust=.3))
+  g <- g + theme(axis.title.y = element_text(angle=90, size=18, vjust=.3))
   g <- g + theme(axis.text.y = element_text(size=12))
   g <- g + theme(axis.line = element_line(colour = 'black', size = 1))
   g <- g + theme(axis.ticks = element_line(colour = 'black', size = 1))
@@ -69,13 +71,11 @@ smd.boxplot <- function(boxplot.df) {
   g <- g + geom_text(aes(10, 1300, label="*"), colour='black', size=8)
   g <- g + geom_text(aes(11, 1300, label="*"), colour='black', size=8)
   
[…]

### 2014-01-05 paper.tex
Pausing point.

@@
 
-\title[Predicting mutant binding]{Predicting virus-receptor mutant binding by molecular dynamics simulation}
-
-\author{Austin G. Meyer}
-\email{austin.meyer@utexas.edu}
-\phone{+1 512 971 0123}
-\affiliation{Section of Integrative Biology, Institute for Cellular and Molecular Biology, and Center for Computational Biology and Bioinformatics, The University of Texas at Austin, Austin, TX 78731, USA}
-\alsoaffiliation{School of Medicine, Texas Tech University Health Sciences Center, Lubbock, TX, 79430, USA}
-\alsoaffiliation{Department of Molecular Biosciences, Institute for Cellular and Molecular Biology, The University of Texas at Austin, Austin, TX 78731, USA}
-
-\author{Sara L. Sawyer}
-
-\author{Andrew D. Ellington}
-\affiliation{Department of Molecular Biosciences, Institute for Cellular and Molecular Biology, The University of Texas at Austin, Austin, TX 78731, USA}
-
-\author{Claus O. Wilke}
-\affiliation{Section of Integrative Biology, Institute for Cellular and Molecular Biology, and Center for Computational Biology and Bioinformatics, The University of Texas at Austin, Austin, TX 78731, USA}
-
+\linespread{1.5}
 \begin{document}
 
+\title{\textbf{Analyzing Machupo virus-receptor binding by molecular dynamics simulations}}
+\author{Austin G. Meyer$^{1,2,3*}$, Sara L. Sawyer$^{2}$, Andrew D. Ellington$^{2}$,\\and Claus O. Wilke$^{1}$}
+\date{}
+
+\maketitle
+\noindent
+Address:\\
+$^1$Department of Integrative Biology, Institute for Cellular and Molecular
+Biology, and Center for\\Computational Biology and Bioinformatics. The University
+of Texas at Austin, Austin, TX 78712, USA.\\
+$^2$Department of Molecular Biosciences, Institute for Cellular and Molecular 
+Biology, The University of Texas at Austin, Austin, TX 78712, USA.\\
+$^3$School of Medicine, Texas Tech University Health Sciences Center, 
+Lubbock, TX 79430, USA.\\
+
+\bigskip
+\noindent
+$^*$Corresponding author\\
+$\phantom{^*}$Email: austin.meyer@utexas.edu\\
+$\phantom{^*}$Phone: +1 512 971 0123\\
+
+\bigskip
+\noindent
+Manuscript type: research article
+
+\bigskip
+\noindent Keywords: protein--protein interaction, molecular dynamics, arenavirus, machupo
+
+\linenumbers
 \newpage
-
-\begin{abstract}
-Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases.
-In particular, the effects of multiple mutations, non-alanine substitutions, and flexible loops are difficult to predict with available tools and protocols.
-We present here a new method to interrogate affinity differences resulting from mutations in a host-virus protein--protein interface.
-Our method is based on extensive non-equilibrium all-atom simulations: We computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1) and estimate affinity using the maximum applied force during a pulling simulation and the area under the force-versus-distance curve.
-We find that these quantities can provide novel biophysical insight into the GP1/hTfR1 interaction.
-First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes.
-Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that one of them may not be important for tight binding.
-Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-resistant hTfR1 mutants.
-Finally, our method provides an elegant framework to compare the effects of multiple mutations, individually and jointly, on protein--protein interactions.
-\end{abstract}
 
@@
 Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure.
-For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \cite{Grant2011,Kortemme2004}.
-However, such methods calculate the relative contribution to overall energy for each amino acid in a protein.
-They do not probe the interface directly, but rather calculate the total energy with each substitution \cite{Grant2011,Kortemme2004}.
-Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \cite{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}.
+For example, end point methods can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}.
+However, such methods can suffer from various simplifying assumptions.
+For example, they generally use an implicit solvent approximation and assume the end state difference with minimal structural rearrangement is sufficient to discriminate energetic differences \citep{Grant2011,Kortemme2004}.
+Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}.
 Unfortunately, such machine-learning approaches often suffer in novel applications, for which available training sets are small or non-existent.
 As such, these methods are poorly suited for most host-virus protein--protein systems.
-By contrast, first principles methods offer the chance to forgo training.
-However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional model to another with no intermediate steps \cite{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
-As a result, the accuracy of the measurement suffers as the two proteins become increasingly dissimilar; therefore, these techniques are largely limited to point mutant comparisons.
+By contrast, first principles methods can forgo training, but currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a transitional model (where one state may be wild-type and the other may be a mutant) to make rigorous free energy calculations  \citep{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
+While these may be considered two of the gold standard techniques for calculating affinity differences, there are a huge number theoretical and technical complexities that must all be properly managed to ensure a converged solution \citep{Gumbart2013b}.
+Such considerations quickly come to dominate the protocol, and the necessary book keeping introduces the possibility of human error \citep{Gumbart2013b}.
+Moreover, as the two ending states look ever more dissimilar the chances of convergence falls rapidly.
+To ensure convergence, these techniques are typically limited to small differences (such as point mutant comparisons) with a few, very impressive exceptions \citep{Wang2006,Gumbart2013,Gumbart2013b}.
+For most, larger differences quickly become intractable as the number of intermediate steps required to compute a converged solution grows or the complexity of adding restraining potentials and computing approximations expands \citep{Wang2006,Gumbart2013,Gumbart2013b}.
 
-Here we propose a new method to probe mutational effects in protein--protein interactions directly based on first principles simulation, by pulling two proteins apart that are initially bound in a complex.
+Here we propose applying an existing method to a new test case to probe mutational effects in protein--protein interactions based on first principles simulation.
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
-Then, we measure the force required for dissociation \cite{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
+Then, we measure the force required for dissociation \citep{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
 As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
+Although such biasing techniques are common in protein-ligand binding problems, they are less commonly applied to protein--protein interactions, and almost never to mutational analysis in a protein--protein system.
+This is largely the result of free energy convergence difficulties and computational limitations \citep{Cuendet2008,Cuendet2011}; using maximum force or simple AUC can solve these problems.
+We also used the traditional dual topology FEP technique to compute relative free energy differences for comparison to our simplified force biased approach.
 
-We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \cite{Abraham2010,Charrel2003}.
-Machupo is an ambisense RNA virus of the arenavirus family \cite{Charrel2003}.
-Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \cite{Charrel2003}.
-Members of the arenavirus family include the Lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the Guanarito, Junin, and Machupo viruses endemic to rodents in South America \cite{Charrel2003}.
-The South American arenaviruses typically infect humans after rodent contamination and can cause a devastating hemorrhagic fever with high mortality \cite{Charrel2003}.
+We used SMD and FEP to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}.
+Machupo virus is an ambisense RNA virus of the arenavirus family \citep{Charrel2003}.
+Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}.
+Members of the arenavirus family include the Lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the Guanarito, Junin, and Machupo viruses endemic to rodents in South America \citep{Charrel2003}.
+The South American arenaviruses typically infect humans after rodent contamination and can cause a devastating hemorrhagic fever with high mortality \citep{Charrel2003}.
 
@@
 The hTfR1 protein contains three extracellular domains: two basilar domains and an apical domain.
-The two basilar domains serve most of the transferrin-binding function \cite{Abraham2010,Rad20112}.
+The two basilar domains serve most of the transferrin-binding function \citep{Abraham2010,Rad20112}.
 Viral entry is initiated by GP1 binding to the apical domain of hTfR1.
-Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \cite{Rad20111,Rad20112}.
-The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain.
-For GP1, several extended loops mediate binding to hTfR1 \cite{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \cite{Abraham2010}.
-Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \cite{Rad20111,Rad20112}.
+Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}.
+The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid $\beta$-pleated sheet domain.
+For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
+Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.
 
 We applied our computational method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes.
+Importantly, the affinity changes predicted using only max force or AUC show a strong correlation with rigorous relative free energy differences computed by FEP.
 At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity may not be as simple as the static structure suggests.
@@
 According to our simulations, this network may not affect binding affinity directly.
-In addition, our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in protein--protein interfaces.
+In addition, our study offers an all-atom steered molecular dynamic approach to avoid some of the pitfalls of several existing methods used to evaluate mutations in protein--protein interfaces.
 
@@
 
-We used the protein visualization software PyMOL \cite{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1.
+For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \citep{Abraham2010}.
+The apical domain of hTfR1 interacts directly with GP1 while the other two domains are closer to the cell membrane and have essentially no interaction with GP1.
+The biophysical independence of the apical domain allowed us to isolate it without significantly affecting the GP1/hTfR1 interaction.
+
+We used the protein visualization software PyMOL \citep{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1.
 No residues were removed from the viral protein.
-We also removed 10 sugar residues from the native structure.
+Figure~\ref{fig:protein_comparison} shows a model of the initial structure and that of the pared structure.
+Although GP1 has several glycosylatable residues, we opted to use the de-glycosylated protein for this study.
+The complexity of correctly parameterizing diverse sugar moieties is outside of the scope of this paper.
+Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \citep{Abraham2010}.
+In total we removed 10 sugars from the crystal structure for this study.
 
-After system reduction, the Visual Molecular Dynamics (VMD) \cite{Humphrey1996} package along with its system of back-ends was used for all subsequent modeling.
+After system reduction, the Visual Molecular Dynamics (VMD) \citep{Humphrey1996} package along with its system of back-ends was used for all subsequent modeling.
 The Orient add-on package allowed us to rotate the system axis such that the direction of steering was oriented directly down the z-axis.
@@
 
-NAMD was used for all simulations in this study \cite{Phillips2005}.
+NAMD was used for all simulations in this study \citep{Phillips2005}.
 In addition to the modeled system, for equilibration we generated a configuration file that fixed the $\alpha$-carbon backbone.
-This was accomplished by setting $\beta = 1$.
-Further, we generated a configuration file with fixed $\alpha$-carbon atoms at residues 41-92 (numbered linearly, in this case, starting at 1 for the first amino acid as was required for VMD) in the hTfR1.
+This was accomplished by setting the B-factor column to 1 for the fixed atoms and to zero for all other atoms.
+Further, we generated a configuration file with fixed $\alpha$-carbon atoms at residues 41-92 (numbered linearly, in this case, starting at 1 for the first amino acid as was required for NAMD) in the hTfR1.
 The second file was used to affix a harmonic restraint, thus preventing any unfolding due to system reduction.
-More importantly, the harmonic restraint allowed the protein complex to equilibrate while preventing any drift from its predefined orientation.
+More importantly, the harmonic restraint allowed the protein complex to equilibrate while preventing any drift from its predefined position; the restraint did not constrain the structure of each protein, or the relative position or orientation of the two proteins to each other.
 Finally, we calculated the system center and dimensions for use in molecular dynamics settings.
 
-We used the Charmm27 \cite{Brooks1983} all-atom force field.
+The exact NAMD configuration file is made available as a supplement and a github repository with all necessary set up files is provided.
+We used the Charmm27 \citep{Brooks1983} all-atom force field.
 The initial system temperature was set to 310K.
-Several typical MD settings were used including switching and cutoff.
-In addition, we used a 2 femtosecond time step with rigid bonds (the exact configuration is provided in the supplement).
+Several typical MD settings were used including switching and cutoff distances (see provided configuration file).
+In addition, we used a 2 femtosecond time step with rigid bonds.
 We used periodic boundary conditions with the particle mesh ewald (PME) method of computing full system electrostatics outside of the explicit box.
-Furthermore, we used a group pressure cell, flexible box, and langevin piston during equilibration.
-Finally, a harmonic restraint (called harmonic constraint in VMD) was set as stated previously.
+Furthermore, we used a group pressure cell, flexible box, langevin barostat and lavegin thermostat during equilibration.
+A harmonic restraint (called harmonic constraint in VMD) was set as stated previously.
 
-To start the simulation, the langevin piston was switched off and the system was minimized for 1000 steps.
+To start the simulation, the barostat was switched off and the system was minimized for 1000 steps.
 Next, the fixed backbone was released, and the system was minimized for an additional 1000 time steps.
@@
 Finally, the langevin piston was turned on and the system was simulated for 2 ns (1,000,000 steps) of chemical time.
-For each mutant, twenty independent replicates were run with an identical protocol.
+For each mutant, twenty independent equilibration replicates were run with an identical protocol.
 
@@
 
+We used the final state from each equilibrated system to restart another MD simulation.
+Our steering protocol is fundamentally similar to \citet{Cuendet2008} with slightly different parameter choices.
+Perhaps the one significant difference lies in our choosing to not use a thermostat or barostat.
+However, this should not be a problem since we are not trying to calculate the binding free energy by any physically rigorous approach (the Jarzynski inequality being one example).
 Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1.
-Also, the $\alpha$-carbon of all residues (163-318 in linear numbering) in the GP1 were given an occupancy of 1, indicating they had an applied force during the simulation.
+These residues were selected as they are far from the binding interface and sufficiently distributed to prevent any orientational motion of the receptor relative to the viral spike protein.
+The $\alpha$-carbon of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation.
+The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms, but rather uses the selection to compute an initial center of mass.
+Then, during the steering run, the single center of mass point is pulled with the parameters desribed below.
+We used the same force field parameters (exclude, cutoff, switching, etc.), the same integrator parameters (time step, rigidbonds on, all molecular being wrapped, etc.), and the same particle mesh ewald parameters as in equilibration.
+Periodic boundary conditions were incorporated as part of the system (as is the convention in NAMD restart) and PME was again used to approximate full system electrostatics.
 
-In every way possible the subsequent steering simulation had the same parameters as the equilibration simulation.
-Thus, the steered molecular dynamics (SMD) \cite{Is2001A,Is2001B} simulation was a typical restart with slightly different settings.
-Periodic boundary conditions were incorporated as part of the system and PME was again used to approximate full system electrostatics.
-In addition, we first tuned the pulling velocity for practical time considerations.
-Subsequently, the spring constant was adjusted to ensure a sufficient signal-to-noise ratio with the given velocity.
-As a result, the spring constant was set to 5~kcal/mol/\AA$^2$, velocity set to 1~\AA/ns, and direction down the positive $z$-axis.
+We ran test simulations at several force constants, and visually inspected the results.
+A force constant of 5~kcal/mol/\AA$^2$ was chosen due to its relatively low signal-to-noise ratio.
+This is slightly lower than the more common 7~kcal/mol/\AA$^2$ found in several recent studies; however, it seems that value is so commonly selected primarily because it is the force constant found in the NAMD, SMD tutorial available through the developers.
+Moreover, the force constant could very likely be set to a range of nearby values with little loss in predictive power.
+In SMD experiments the pulling velocity should be as low as possible for the available computational time \citet{Cuendet2008,Cuendet2011}.
+Thus we choose a velocity of 0.000001~\AA/fs = 1~\AA/ns, and direction down the positive $z$-axis.
+This number could be set to pull faster if the computing time must be reduced, but slower than necessary pulling speeds are not typically considered problematic.
 
-Finally, SMD was run for 15 ns (7,500,000 time steps) of chemical time.
+SMD was run for 15 ns (7,500,000 time steps) of chemical time.
 For each simulation, we randomly selected one of the equilibration runs for restart.
@@
 
+\subsection{Free Energy Perturbation}
+
+Briefly, we used the traditional dual topology approach to FEP.
+This involves a thermodynamic cycle where a set of atoms are progressively decoupled from the environment while another set of atoms are progressively coupled.
+To compute the relative free energy difference requires knowing the free energy change when the transformation is carried out for the bound complex and the individual protein.
+Then, one can compute the relative free energy difference between a WT and mutant complex by taking the difference between the energy required to decouple/couple the atoms in solution from the energy required to decouple/couple the atoms in the bound complex.
+
+Again, the NAMD configuration file is made available as a supplemental file.
+We used a similar configuration to that in equilibration.
+One significant difference was to make a cubic water box with a side length equal to the long axis of the complex plus a 10~\AA buffer on either side, and simply restrict center of mass motion with the NAMD setting.
+This was done to avoid affecting the system energy while calculating free energy differences.
+
+The transition protocol for bound and free protein systems were identical.
+They started with 1000 steps of minimization and 250,000 steps of equilibration in the starting state for the forward and reverse directions.
+Phase transitions were carried out in steps of $\lambda$=0.05.
+Each transition was carried out for 250,000 steps.
+The first 100,000 steps after phase transition were reserved for equilibration and the final 150,000 steps were used for data collection.
+
+The FEP analysis tool in VMD was used for subsequent analysis.
+We used it to perform error analysis and compute the Bennett acceptance ratio as the maximum likelihood free energy difference of the two states under consideration.
+Though the larger transitions presented difficulty in a small number of windows, forward and reverse hysteresis was generally in good agreement for all complexes.
+The double mutants were performed by first doing the Y211A mutation followed by the other of the two mutants.
+Then, the numbers were simply added together to get the total energetic difference.
+
 \subsection{Post-processing}
 
-The python packages MDAnalysis \cite{Agrawal2011} and ProDy \cite{Bakan2011} were both used at various points in post-processing.
+The python packages MDAnalysis \citep{Agrawal2011} and ProDy \citep{Bakan2011} were both used at various points in post-processing.
 The molecular trajectory (comprising the atomic coordinates per time) was parsed to compute the center-of-mass for each of the two complexes.
 The starting center-of-mass distance was set to zero and the distance was re-computed at each time step relative to the starting distance.
-The force was output directly to the standard log file.
 
@@
 Each of the 50 independent trajectories per mutant produced a fairly noisy force curve.
-To reduce the noise, we averaged each point with a 50 point interpolation window.
-The window ensured that the averaged maximum applied force was never cut off near the beginning of the simulation due to lost time steps (caused by averaging).
+The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting a 50 frame interpolation window for averaging.
+The latter approach was used for statistical analysis.
 The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC).
 We used the pairwise.t.test function with the FDR p-value correction to test for significance.
-The ggplot \cite{ggplot} package was used to generate most of the figures.
+The ggplot \citep{ggplot} package was used to generate most of the figures.
 
@@
 There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV.
-Virtually all of the important sites have been mapped by \textit{in vitro} flow-cytometry based entry assays.
+Many of the important sites have been mapped by \textit{in vitro} flow-cytometry based entry assays.
 The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals).
-In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \cite{Grant2011,Kortemme2004} only marginally useful.
+In addition, much of the binding domain on hTfR1 is on a loop that is flexible prior to viral binding, but organizes to become a strand of a $\beta$-sheet on binding.
+As a result,  many other computational techniques \citep{Grant2011,Kortemme2004} only marginally useful.
 The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.
 
-For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \cite{Abraham2010}.
-The apical domain of hTfR1 interacts directly with GP1 while the other two domains are closer to the cell membrane and have essentially no interaction with GP1.
-The biophysical independence of the apical domain allowed us to isolate it without significantly affecting the GP1/hTfR1 interaction.
-Figure~\ref{fig:protein_comparison} shows a model of the initial structure and that of the pared structure.
-Although GP1 has several glycosylatable residues, we opted to use the de-glycosylated protein for this study.
-The complexity of correctly parameterizing diverse sugar moieties is outside of the scope of this paper.
-Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \cite{Abraham2010}.
-
 In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}).
-Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \cite{Rad2008,Rad20111,Rad20112}.
+All of the mutations are within 5~\AA of the protein--protein interface.
+Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
 Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1.
-The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \cite{Abraham2010}.
-In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \cite{Rad2008,Abraham2010} to GP1.
-In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \cite{Rad2008,Abraham2010}.
-Finally, an alanine mutation at site 111 in GP1 (mutation vArg111Ala) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \cite{Rad20112}.
+The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}.
+In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1.
+In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}.
+Finally, an alanine mutation at site 111 in GP1 (mutation vArg111Ala) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}.
 For notation purposes, the viral site is always referred to with a preceding `v'.
 
-\subsection{Steered molecular dynamics simulations}
+Despite, the fact that viral binding occurs at the site of a flexible loop in the free hTfR structure, our data shows after binding the strand is extremely rigid.
+None of the sites in the interfacial strand have root mean squared fluctuation (RMSF) values in the top half of all sites in the complex during equilibration (figure).
+This is unsurprising considering the high degree of burial that occurs as a result of viral binding.
+In addition, computing the root mean squared deviation (RMSD) of the entire structure over the trajectory shows that none of the mutations are so deleterious as the cause rapid unbinding.
+In fact, the RMSD over trajectory looks incredibly invariant across mutants (figure).
 
-We proceeded to analyze this system using molecular dynamics.
-To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions.
-First, we generated a physically realistic simulation world around the protein complex.
-The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions.
-Next, we carried out successive rounds of energy minimization with a fixed $\alpha$-carbon backbone, followed by heating, and backbone release.
-Last, we performed a short round of all-atom molecular dynamics with minimal restraints \cite{Cuendet2008}.
-We used the final state from each equilibrated system to restart another MD simulation.
-For SMD, at the centroid of several atoms a force was applied to a single point in GP1.
-By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \cite{Cuendet2008,Cuendet2011}.
+\subsection{Molecular dynamics simulations}
+
+We proceeded to analyze this system using two molecular dynamics techniques.
+First, in SMD by using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
 A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}.
-The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging.
-The latter approach was used for statistical analysis.
-In addition, to minimize the computational time in unproductive molecular dynamics, we shortened the simulation time such that the interaction just barely dropped to zero near the end of each run.
 As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants.
@@
 Their summary statistics are reported in Table~\ref{tab:summary_statistics}.
-As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \cite{Is2001A,Is2001B}.
+As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.
+Indeed, our set up would not have allowed rigorous calculations as we omitted several tools such as a barostat and a thermostat that would be required for an NPT ensemble.
 
 Before systematically applying SMD to the GP1/hTfR1 interaction, we needed to ensure the method was sufficiently sensitive to distinguish between relatively minor point mutations.
-While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \cite{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
+While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
 For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.
+We proceeded to compare our SMD results to that of the standard dual topology FEP approach to calculate relative free energy differences.
+The correlation between the energetically rigorous FEP and our statistical approach is high.
+For all 11 complexes tested, the correlation between max force and FEP was r=0.767 at p=0.00576, and the correlation between AUC and FEP was r=0.72 at p=0.01 (figure).
+As a result of the strong correlation, we will generally refer to the SMD differences only.
 
@@
 Further, we identified two test statistics, maximum applied force and AUC, that can be used as proxies for binding affinity.
+Both of these statistics correlate well with FEP, but offer the simplicity of not requiring a large commitment to planning for the theoretical issues inherent to free energy methods.
 We systematically tested several point mutations to understand their contribution to the binding interaction.
@@
 
-By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application.
-Traditionally this technique has been either applied to compute equilibrium quantities via a non-equilibrium approximation \cite{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \cite{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \cite{Dixit2001}.
-Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \cite{Cuendet2011,Giorgino2011}.
-And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \cite{Sulkowska2010}.
+Traditionally SMD has been either applied to compute equilibrium free energies via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
+Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}.
 Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions.
-In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application.
-Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanisms.
-Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation.
+Via SMD, one can separate mutations whose likely effect is altered binding affinity with simple statistics like maximum force of separation.
 Thus, SMD may open avenues for subsequent experimental work.
 
-Our findings may rationalize several effects observed in both infectivity data and rodent populations \cite{Rad2008,Rad20111}.
+Our findings rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}.
 First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding.
@@
 That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually.
-Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \cite{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
+Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
 Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.
 
-We would like to emphasize here that we cannot expect perfect agreement between our simulations and the available experimental data.
+We would like to emphasize here that we cannot expect perfect agreement between our simulations and the available experimental data, but the correspondence to a well established free energy method is bolsters our conclusions.
 While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method.
-First, it is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method.
+First, it is possible that some mutants display measurable phenotypic effects in experiments yet appear identical in simulation.
 More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph).
-Second, our SMD method is fundamentally limited by the accuracy of our starting structure.
+Second, the SMD method is fundamentally limited by the accuracy of our starting structure.
 Third, the available experimental data for the GP1/hTfR1 system were generally obtained from entry assays or whole-cell binding assays rather than molecular binding assays.
@@
 by determining the optimal pulling speed.
-Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \cite{Shaw2012}.
-Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \cite{Shaw2011}.
+Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \citep{Shaw2012}.
+Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \citep{Shaw2011}.
 With improved resources, we could investigate thousands of individual permutations in the GP1/hTfR1 binding interface.
-In addition, with additional compute time it would be possible to incorporate equilibrium sampling approaches \cite{Buch2011} or use brute force equilibrium approaches \cite{Giorgino2012} to improve resolution.
-Going forward, we believe that all-atom molecular dynamics simulations can become an increasingly viable approach for protein interaction studies.
+In addition, with additional compute time it would be possible to incorporate equilibrium sampling approaches \citep{Buch2011} or use brute force equilibrium approaches \citep{Giorgino2012} to improve resolution.
 
-\section{Author Information}
-
-\subsection{Corresponding Author}
-
-austin.meyer@utexas.edu
-
-\subsection{Notes}
-
-The authors declare no competing financial interest.
+For future studies, although our approach offers the simplicity of not requiring prior knowledge about a system of interest (other than a bound model), all atom, SMD is likely not the best approach for most relative affinity calculations.
+To ensure one's results are independent of the dissociation path one selects would require computing the work of separation for all likely paths.
+Such an approach eventually requires using the Jarzynski inequality \citep{Jar1997} to establish a lower limit for binding energy and would quickly become computationally inefficient for evaluating a large number of mutations in most systems.
+By contrast, another all atom approach that makes calculations via short simulations of spatially restrained complexes has proven capable of generating relatively accurate binding affinities with less compute time than is required from our steering strategy \citep{Gumbart2013b}.
 
@@
 
-This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) and the Welch Foundation (F-1654) to A.D.E., the National Institutes of Health (R01-GM088344) to C.O.W., and the National Institutes of Health (R01-GM093086) to S.L.S.
-
+\bibliographystyle{MBE}
 \bibliography{paper}

### 2014-01-05 paper.tex
Preliminary final

@@
 For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.
+
 We proceeded to compare our SMD results to that of the standard dual topology FEP approach to calculate relative free energy differences.
@@
 Via SMD, one can separate mutations whose likely effect is altered binding affinity with simple statistics like maximum force of separation.
-Thus, SMD may open avenues for subsequent experimental work.
+Thus, SMD may open avenues for subsequent experimental work in some situations where FEP may be prohibitively difficult.
 
@@
 
-For future studies, although our approach offers the simplicity of not requiring prior knowledge about a system of interest (other than a bound model), all atom, SMD is likely not the best approach for most relative affinity calculations.
+For future studies, although our approach offers the simplicity of not requiring prior knowledge about a system of interest (other than a bound model), at this point SMD may not the best approach for many relative affinity calculations.
 To ensure one's results are independent of the dissociation path one selects would require computing the work of separation for all likely paths.
 Such an approach eventually requires using the Jarzynski inequality \citep{Jar1997} to establish a lower limit for binding energy and would quickly become computationally inefficient for evaluating a large number of mutations in most systems.
-By contrast, another all atom approach that makes calculations via short simulations of spatially restrained complexes has proven capable of generating relatively accurate binding affinities with less compute time than is required from our steering strategy \citep{Gumbart2013b}.
+However, considering the strong correlation between FEP and SMD in this system, it may not be important to ensure one's results are path independent for relative affinity calculations.
+
+More importantly, with no \textit{a priori} knowledge of the appropriate number of equilibration samples, the during of equilibration, the appropriate number of pulling runs, or the best pulling speed means the computational expense in our SMD protocol may not be commensurate with the information provided.
+For example, another all atom approach that makes calculations via short simulations of spatially restrained complexes has proven capable of generating relatively accurate binding affinities with less compute time than is required from our steering strategy \citep{Gumbart2013,Gumbart2013b}.
+That being said, there is no reason to believe this SMD approach to mutagenic studies could not be optimized to reduce computational expense.
+Further analysis would be needed to understand the lower limits of resources required for accurate predictions.
 

### 2014-01-05 paper.tex
Prelim final.

@@
 We used a similar configuration to that in equilibration.
-One significant difference was to make a cubic water box with a side length equal to the long axis of the complex plus a 10~\AA buffer on either side, and simply restrict center of mass motion with the NAMD setting.
+One significant difference was to make a cubic water box with a side length equal to the long axis of the complex plus a 10~\AA\  buffer on either side, and simply restrict center of mass motion with the NAMD setting.
 This was done to avoid affecting the system energy while calculating free energy differences.
@@
 In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}).
-All of the mutations are within 5~\AA of the protein--protein interface.
+All of the mutations are within 5~\AA\ of the protein--protein interface.
 Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.

### 2014-01-05 paper.tex
Added two figures.

@@
 Despite, the fact that viral binding occurs at the site of a flexible loop in the free hTfR structure, our data shows after binding the strand is extremely rigid.
-None of the sites in the interfacial strand have root mean squared fluctuation (RMSF) values in the top half of all sites in the complex during equilibration (figure).
+None of the sites in the interfacial strand have root mean squared fluctuation (RMSF) values in the top half of all sites in the complex during equilibration (Figure~\ref{fig:rmsf}).
 This is unsurprising considering the high degree of burial that occurs as a result of viral binding.
 In addition, computing the root mean squared deviation (RMSD) of the entire structure over the trajectory shows that none of the mutations are so deleterious as the cause rapid unbinding.
-In fact, the RMSD over trajectory looks incredibly invariant across mutants (figure).
+In fact, the RMSD over trajectory looks incredibly invariant across mutants (Figure~\ref{fig:rmsd}).
 
@@
 (B) The reduced structure used in SMD simulations.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/rmsfs.pdf}}
+\caption{\label{fig:rmsf}\title{RMSF values.} The RMSF values for every site in the bound complex computed during the equilibration phase of the protocol.
+Each color represents the average over 20 trajectories of a single mutant.
+The plot shows the flexibility of each site is essentially independent of mutation, and none of the sites above 0.72~\AA\ are found in the interface.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/rmsds.pdf}}
+\caption{\label{fig:rmsd}\title{RMSD values.} The RMSD values over the time of the trajectory computed during the equilibration phase of the protocol.
+Each color represents the average over 20 trajectories of a single mutant.
+The plot shows none of the mutants causes immediate unbinding of the protein--protein complex.
+In addition, the universal upward trend near the end of the equilibration trajectories may indicate the crystal is more tightly packed than would normally occur in solution.}
 \end{figure}

### 2014-01-05 paper.tex
Updates for Claus's suggestions and added the final figure and data.

@@
 These residues were selected as they are far from the binding interface and sufficiently distributed to prevent any orientational motion of the receptor relative to the viral spike protein.
-The $\alpha$-carbon of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation.
+The center of mass of the $\alpha$-carbons of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation.
 The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms, but rather uses the selection to compute an initial center of mass.
@@
 We proceeded to analyze this system using two molecular dynamics techniques.
-First, in SMD by using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
+First, in SMD by using a known force constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
 A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}.
@@
 As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.
-Indeed, our set up would not have allowed rigorous calculations as we omitted several tools such as a barostat and a thermostat that would be required for an NPT ensemble.
 
@@
 The correlation between the energetically rigorous FEP and our statistical approach is high.
-For all 11 complexes tested, the correlation between max force and FEP was r=0.767 at p=0.00576, and the correlation between AUC and FEP was r=0.72 at p=0.01 (figure).
+For all 11 complexes tested, the correlation between max force and FEP was r=-0.767 at p=0.005857 (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was r=--0.729 at p=0.011.
 As a result of the strong correlation, we will generally refer to the SMD differences only.
@@
 The authors acknowledge Bryan Sutton for opening access to the Hrothgar cluster.
+The authors also acknowledge the reviewers Ilan Samish and Matteo Masetti for their helpful comments in preparing this manuscript.
 
@@
 \begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/fep_mutants.pdf}}
+\caption{\label{fig:fep}\title{Max force versus free energy perturbation} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex.
+The WT complex for FEP was simply set to 0.0.
+The correlation between the two is r=-0.767 with p=0.005857.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
 \centerline{\includegraphics[width=6in]{figures/mutant_comparison_boxplot.pdf}}
@@
 \caption{\label{tab:summary_statistics}Summary statistics for each mutation tested.
-$\mu_\text{MAF}$ is the mean and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations.
-$\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations.}
+$\mu_\text{MAF}$ is the mean in piconewtons and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations.
+$\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations.
+$\Delta G$ is the free energy difference in kcal/mol calculated via FEP by the dual topology paradigm.}
 \begin{center}
-  \resizebox{11cm}{!} {
-    \begin{tabular}{l c c c c}
+  \resizebox{15cm}{!} {
+    \begin{tabular}{l c c c c c}
     \hline
-      Mutation &      $\mu_\text{MAF}$ & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ \\ \hline
-            WT &                772.00 &              154.13 &         198794.7 &            66765.64 \\
-         N348A &                799.33 &              154.36 &         202247.3 &            65175.78 \\
-         N348K &                731.36 &              151.01 &         179087.5 &            55799.49 \\
-         N348W &                732.13 &              131.21 &         192043.2 &            66916.36 \\
-         Y211A &                684.03 &              126.69 &         153517.7 &            40889.12 \\ 
-         Y211D &                875.44 &              116.46 &         234909.3 &            50494.11 \\ 
-         Y211T &                861.17 &              158.48 &         231078.7 &            97569.29 \\
-        vR111A &                754.39 &              103.25 &         185896.8 &            46408.74 \\
-   N348A/Y211A &                762.96 &              144.81 &         173059.2 &            59898.10 \\
-   N348W/Y211A &                617.21 &              163.33 &         147705.6 &            48548.61 \\
-  vR111A/Y211A &                780.71 &              135.65 &         189919.1 &            48868.16 \\
+      Mutation & $\mu_\text{MAF}$ (pN) & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ & $\Delta G$ (kcal/mol)\\ \hline
+            WT &                772.00 &              154.13 &         198794.7 &            66765.64 &  0.000\\
+         N348A &                799.33 &              154.36 &         202247.3 &            65175.78 & -2.149\\
+         N348K &                731.36 &              151.01 &         179087.5 &            55799.49 & +3.184\\
+         N348W &                732.13 &              131.21 &         192043.2 &            66916.36 & +3.033\\
+         Y211A &                684.03 &              126.69 &         153517.7 &            40889.12 & +2.526\\ 
+         Y211D &                875.44 &              116.46 &         234909.3 &            50494.11 & -2.760\\ 
+         Y211T &                861.17 &              158.48 &         231078.7 &            97569.29 & +0.876\\
+        vR111A &                754.39 &              103.25 &         185896.8 &            46408.74 & +0.467\\
+   N348A/Y211A &                762.96 &              144.81 &         173059.2 &            59898.10 & +5.203\\
+   N348W/Y211A &                617.21 &              163.33 &         147705.6 &            48548.61 & +8.207\\
+  vR111A/Y211A &                780.71 &              135.65 &         189919.1 &            48868.16 & -2.441\\
     \hline

### 2014-01-05 fep_data/all_mutants.R

@@ -1,0 +1,42 @@
+all.mutants <- data.frame(id=c('WT','Y211A', 'Y211D', 'Y211T', 'N348A', 'N348K', 'N348W', 'vR111A','211A_348A','211A_348W','211A_111A'), 
+                          smd=c(772.00,684.03, 875.44, 861.17, 799.33, 731.36, 732.13, 754.39,762.96,617.21,780.71),
+                          auc=c(198794.7,153517.7,234909.3,231078.7,202247.3,179087.5,192043.2,185896.8,173059.2,144705.6,189919.1), 
+                          fep=c(0,2.525525,-2.760073,0.875825,-2.149291,3.184248,3.033377,0.4666468,5.203034,8.206784,-2.440771))
+
+smd.points <- function(df) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=18
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
+                            axis.line=element_line(),
+                            panel.grid.minor=element_blank(),
+                            panel.grid.major=element_blank(),
+                            panel.background=element_blank(),
+                            panel.border=element_blank(),
+                            axis.line=element_line())
+  
+  g <- ggplot(df, aes(x=fep, y=smd)) + geom_point(size=3) + geom_smooth(method=lm)
+  g <- g + theme(strip.background=element_blank())
+  g <- g + ylab('Interpolated Maximum Force (pN)')
+  g <- g + xlab('Free Energy Perturbation (kcal/mol)')
+  g <- g + scale_x_continuous(breaks=seq(-10, 10, 2), limits=c(-3.5, 8.5))
+  g <- g + scale_y_continuous(breaks=seq(500, 1000, 100), limits=c(550, 900))
+  g <- g + theme(panel.border=element_blank(), axis.line=element_line())
+  g <- g + theme(axis.title.x = element_text(size=18, vjust=0))
+  g <- g + theme(axis.text.x = element_text(size=18))
+  g <- g + theme(axis.title.y = element_text(size=18, vjust=0.25))
+  g <- g + theme(axis.text.y = element_text(size=18))
+  g <- g + theme(axis.line = element_line(colour = 'black', size = 1))
+  g <- g + theme(axis.ticks = element_line(colour = 'black', size = 1))
[…]

### 2014-01-05 paper.tex
Added a lot of data and scripts as well as many manuscript updates.

@@
 Despite, the fact that viral binding occurs at the site of a flexible loop in the free hTfR structure, our data shows after binding the strand is extremely rigid.
-None of the sites in the interfacial strand have root mean squared fluctuation (RMSF) values in the top half of all sites in the complex during equilibration (Figure~\ref{fig:rmsf}).
+All two sites had receptor strand have root mean squared fluctuation (RMSF) values in the top half of all receptor sites during equilibration (Figure~\ref{fig:rmsf}), and those are almost completely exposed to solvent.
 This is unsurprising considering the high degree of burial that occurs as a result of viral binding.
-In addition, computing the root mean squared deviation (RMSD) of the entire structure over the trajectory shows that none of the mutations are so deleterious as the cause rapid unbinding.
+Computing the root mean squared deviation (RMSD) of the entire structure over the trajectory shows that none of the mutations are so deleterious as the cause rapid unbinding.
 In fact, the RMSD over trajectory looks incredibly invariant across mutants (Figure~\ref{fig:rmsd}).
+By comparison to the equilibration RMSF calculation, near the end of the SMD trajectory all of the residues in the WT receptor interfacial strand are in the top half of RMSF over all receptor sites (Figure~\ref{fig:rmsf_comparison}).
+Thus, if sufficient simulation time is not dedicated to allowing this unfolding process, standard free energy techniques may miss the energetic contributions that result from ordering the flexible loop in the hTfR apical domain.
 
@@
 \centerline{\includegraphics[width=6in]{figures/rmsfs.pdf}}
-\caption{\label{fig:rmsf}\title{RMSF values.} The RMSF values for every site in the bound complex computed during the equilibration phase of the protocol.
+\caption{\label{fig:rmsf}\title{RMSF values during equilibration.} The RMSF values for every site in the bound complex computed during the equilibration phase of the protocol.
 Each color represents the average over 20 trajectories of a single mutant.
-The plot shows the flexibility of each site is essentially independent of mutation, and none of the sites above 0.72~\AA\ are found in the interface.}
+Indices 17-25 are the hTfR flexible loop.
+The plot shows the flexibility of each site is essentially independent of mutation, and two sites above 0.72~\AA\ are found on the receptor in the interface.}
 \end{figure}
@@
 \centerline{\includegraphics[width=6in]{figures/rmsds.pdf}}
-\caption{\label{fig:rmsd}\title{RMSD values.} The RMSD values over the time of the trajectory computed during the equilibration phase of the protocol.
+\caption{\label{fig:rmsd}\title{RMSD values during equilibration.} The RMSD values over the time of the trajectory computed during the equilibration phase of the protocol.
 Each color represents the average over 20 trajectories of a single mutant.
@@
 In addition, the universal upward trend near the end of the equilibration trajectories may indicate the crystal is more tightly packed than would normally occur in solution.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/rmsf_comparison.pdf}}
+\caption{\label{fig:rmsf_comparison}\title{RMSF values of WT hTfR in equilibration and SMD.} The RMSF values for every site in the WT receptor were computed during the equilibration phase and during final 50 frames of the SMD trajectories.
+The black line was computed over equilibration and the red line during SMD.
+The plot shows the solution mobility of the hTfR flexible loop increases more than the average during the unbinding process.}
 \end{figure}

### 2014-01-05 eq_structural_analysis/eq_structural_summary.R

@@ -1,0 +1,116 @@
+setwd('~/MolecularDynamics/3kas_hrothgar_eq/')
+
+list.dirs <- function(path=".", pattern=NULL, all.dirs=FALSE,
+                      full.names=FALSE, ignore.case=FALSE) {
+  
+  all <- list.files(path, pattern, all.dirs,
+                    full.names, recursive=FALSE, ignore.case)
+  all[file.info(all)$isdir]
+}
+
+folders <- list.dirs()
+folders <- folders[-8]
+folders <- folders[-10]
+
+frames <- dim(read.table(paste(folders[1],'/structural_analysis/rmsd_mean.txt',sep='')))[1]
+atoms <- dim(read.table(paste(folders[1],'/structural_analysis/rmsf_mean.txt',sep='')))[1]
+rmsd.data <- data.frame(matrix(ncol = length(folders), nrow=frames))
+rmsf.data <- data.frame(matrix(ncol = length(folders), nrow=atoms))
+colnames(rmsd.data) <- folders
+colnames(rmsf.data) <- folders
+
+for(folder in folders){
+  rmsd.data[[folder]] <- read.table(paste(folder,'/structural_analysis/rmsd_mean.txt',sep=''))[,1]
+  rmsf.data[[folder]] <- read.table(paste(folder,'/structural_analysis/rmsf_mean.txt',sep=''))[,1]
+}
+
+pdf('structural_analysis/rmsds.pdf', width=12, height=11, useDingbats=F)
+par(mgp=c(5.5,1.5,0))
+par(mar=c(7,8,1,1))
+par(oma=c(0,0,0,0))
+
+plot(rmsd.data[,1],
+     type='l', 
+     lwd=3,
+     ylim=c(0.6, 1.5),
+     xlab='Frames',
+     ylab='RMSD (Angstroms)',
+     axes=F,
+     xaxt="n",
+     yaxt="n",
+     cex.lab=3)
+
+axis(1,
+     col.axis="black",
+     lwd=4,
+     at=seq(0,2000,200),
+     line=1.0,
+     tck=0.01,
+     cex.axis=3)
+
+axis(2,
+     col.axis='black',
+     at=c(0.6, 0.8, 1.0, 1.2, 1.4, 1.6),
+     lwd=4,
+     line=-.6,
+     tck=0.01,
+     las=1,
+     cex.axis=3)
+
[…]

### 2014-01-05 smd_rmsf_data/smd_structural_summary.R

@@ -1,0 +1,53 @@
+setwd('~/Desktop/MACV_SMD/smd_rmsf_data/data/')
+
+files <- list.files()
+
+atoms <- dim(read.table(paste(files[1])))[1]
+
+rmsf.data <- data.frame(matrix(ncol = length(files), nrow=atoms))
+
+colnames(rmsf.data) <- files
+
+for(file in files){
+  rmsf.data[[file]] <- read.table(paste(file))[,1]
+}
+
+smd.rmsf <- rowMeans(rmsf.data)
+eq.rmsf <- read.table('../eq_wt_rmsf.dat')[,1]
+
+pdf('../rmsf_comparison.pdf', width=12, height=11, useDingbats=F)
+par(mgp=c(5.5,1.5,0))
+par(mar=c(7,8,1,1))
+par(oma=c(0,0,0,0))
+
+plot(eq.rmsf[1:160],
+     type='l', 
+     lwd=3,
+     ylim=c(0.4, 2.2),
+     xlim=c(0, 165),
+     xlab='Index',
+     ylab='RMSF (Angstroms)',
+     axes=F,
+     xaxt="n",
+     yaxt="n",
+     cex.lab=3)
+
+axis(1,
+     col.axis="black",
+     at=seq(0,350,25),
+     line=1.0,
+     lwd=4,
+     tck=0.01,
+     cex.axis=3)
+
+axis(2,
+     col.axis='black',
+     at=c(seq(0.4,20,0.4)),
+     line=-.6,
+     lwd=4,
+     tck=0.01,
+     las=1,
+     cex.axis=3)
+
+lines(smd.rmsf[1:160], col='red', lwd='3')
+dev.off()

### 2014-01-05 paper.tex
Claus's edits.

@@
 By contrast, first principles methods can forgo training, but currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a transitional model (where one state may be wild-type and the other may be a mutant) to make rigorous free energy calculations  \citep{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
-While these may be considered two of the gold standard techniques for calculating affinity differences, there are a huge number theoretical and technical complexities that must all be properly managed to ensure a converged solution \citep{Gumbart2013b}.
+While these may be considered two of the gold standard techniques for calculating affinity differences, there are a huge number of theoretical and technical complexities that must all be properly managed to ensure a converged solution \citep{Gumbart2013b}.
 Such considerations quickly come to dominate the protocol, and the necessary book keeping introduces the possibility of human error \citep{Gumbart2013b}.
-Moreover, as the two ending states look ever more dissimilar the chances of convergence falls rapidly.
+Moreover, as the two ending states look ever more dissimilar the chances of convergence fall rapidly.
 To ensure convergence, these techniques are typically limited to small differences (such as point mutant comparisons) with a few, very impressive exceptions \citep{Wang2006,Gumbart2013,Gumbart2013b}.
-For most, larger differences quickly become intractable as the number of intermediate steps required to compute a converged solution grows or the complexity of adding restraining potentials and computing approximations expands \citep{Wang2006,Gumbart2013,Gumbart2013b}.
+For most investigators, larger differences quickly become intractable as the number of intermediate steps required to compute a converged solution grows or the complexity of adding restraining potentials and computing approximations expands \citep{Wang2006,Gumbart2013,Gumbart2013b}.
 
-Here we propose applying an existing method to a new test case to probe mutational effects in protein--protein interactions based on first principles simulation.
+Here we propose that much of these complexities can be avoided if all we are interested in is a relative comparison of the effects of different mutations on protein-protein interactions, rather than measuring an absolute or relative binding affinity with experimentally realistic units.
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
@@
 As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
-Although such biasing techniques are common in protein-ligand binding problems, they are less commonly applied to protein--protein interactions, and almost never to mutational analysis in a protein--protein system.
+Although such biasing techniques are commonly used in protein-ligand binding problems, they are less commonly applied to protein--protein interactions, and almost never to mutational analysis in a protein--protein system.
 This is largely the result of free energy convergence difficulties and computational limitations \citep{Cuendet2008,Cuendet2011}; using maximum force or simple AUC can solve these problems.
-We also used the traditional dual topology FEP technique to compute relative free energy differences for comparison to our simplified force biased approach.
+For comparison, we also calculate relative free energy differences using the traditional dual topology FEP paradigm, and we show that the two approaches yield congruent results.
 
@@
 
-For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \citep{Abraham2010}.
+For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDB-ID: 3KAS) \citep{Abraham2010}.
 The apical domain of hTfR1 interacts directly with GP1 while the other two domains are closer to the cell membrane and have essentially no interaction with GP1.
@@
 
-The exact NAMD configuration file is made available as a supplement and a github repository with all necessary set up files is provided.
+The exact NAMD configuration file is made available as a github repository with all necessary set up files is provided.
 We used the Charmm27 \citep{Brooks1983} all-atom force field.
 The initial system temperature was set to 310K.
-Several typical MD settings were used including switching and cutoff distances (see provided configuration file).
+Several typical MD settings were used including switching and cutoff distances (see provided configuration files).
 In addition, we used a 2 femtosecond time step with rigid bonds.
 We used periodic boundary conditions with the particle mesh ewald (PME) method of computing full system electrostatics outside of the explicit box.
-Furthermore, we used a group pressure cell, flexible box, langevin barostat and lavegin thermostat during equilibration.
+Furthermore, we used a group pressure cell, flexible box, langevin barostat, and lavegin thermostat during equilibration.
 A harmonic restraint (called harmonic constraint in VMD) was set as stated previously.
@@
 Subsequently, the system was released into all-atom molecular dynamics for 3000 steps.
-Finally, the langevin piston was turned on and the system was simulated for 2 ns (1,000,000 steps) of chemical time.
+Finally, the langevin barostat was turned on and the system was simulated for 2 ns (1,000,000 steps) of chemical time.
 For each mutant, twenty independent equilibration replicates were run with an identical protocol.
@@
 Perhaps the one significant difference lies in our choosing to not use a thermostat or barostat.
-However, this should not be a problem since we are not trying to calculate the binding free energy by any physically rigorous approach (the Jarzynski inequality being one example).
+We can make this choice because we are not trying to calculate the binding free energy by any physically rigorous approach (the Jarzynski inequality being one example).
 Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1.
@@
 
-We ran test simulations at several force constants, and visually inspected the results.
+We ran test simulations at several force constants and visually inspected the results.
 A force constant of 5~kcal/mol/\AA$^2$ was chosen due to its relatively low signal-to-noise ratio.
-This is slightly lower than the more common 7~kcal/mol/\AA$^2$ found in several recent studies; however, it seems that value is so commonly selected primarily because it is the force constant found in the NAMD, SMD tutorial available through the developers.
+This constant is slightly lower than the more common 7~kcal/mol/\AA$^2$ found in several recent studies; that value is commonly selected primarily because it is the force constant found in the NAMD, SMD tutorial available through the developers.
 Moreover, the force constant could very likely be set to a range of nearby values with little loss in predictive power.
@@
 
-Again, the NAMD configuration file is made available as a supplemental file.
+Again, the NAMD configuration file is made available via github.
 We used a similar configuration to that in equilibration.
@@
 The double mutants were performed by first doing the Y211A mutation followed by the other of the two mutants.
-Then, the numbers were simply added together to get the total energetic difference.
+Then, the $\Delta G$'s were simply added together to get the total energetic difference.
 
@@
 In addition, much of the binding domain on hTfR1 is on a loop that is flexible prior to viral binding, but organizes to become a strand of a $\beta$-sheet on binding.
-As a result,  many other computational techniques \citep{Grant2011,Kortemme2004} only marginally useful.
+As a result,  many other computational techniques \citep{Grant2011,Kortemme2004} are only marginally useful.
 The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.
@@
 All of the mutations are within 5~\AA\ of the protein--protein interface.
-Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
+Mutations in hTfR1 at site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
 Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1.
@@
 
-Despite, the fact that viral binding occurs at the site of a flexible loop in the free hTfR structure, our data shows after binding the strand is extremely rigid.
-All two sites had receptor strand have root mean squared fluctuation (RMSF) values in the top half of all receptor sites during equilibration (Figure~\ref{fig:rmsf}), and those are almost completely exposed to solvent.
+Despite the fact that viral binding occurs at the site of a flexible loop in the free hTfR structure, our data shows after binding the strand is extremely rigid.
+In the bound conformation, two sites of the loop have root mean squared fluctuation (RMSF) values in the top half of all receptor sites during equilibration (Figure~\ref{fig:rmsf}), and those are almost completely exposed to solvent.
 This is unsurprising considering the high degree of burial that occurs as a result of viral binding.
-Computing the root mean squared deviation (RMSD) of the entire structure over the trajectory shows that none of the mutations are so deleterious as the cause rapid unbinding.
-In fact, the RMSD over trajectory looks incredibly invariant across mutants (Figure~\ref{fig:rmsd}).
+Computing the root mean squared deviation (RMSD) of the entire structure over the trajectory shows that none of the mutations are so deleterious as to cause rapid unbinding.
+In fact, the RMSD over trajectory looks highly invariant across mutants (Figure~\ref{fig:rmsd}).
 By comparison to the equilibration RMSF calculation, near the end of the SMD trajectory all of the residues in the WT receptor interfacial strand are in the top half of RMSF over all receptor sites (Figure~\ref{fig:rmsf_comparison}).
@@
 The correlation between the energetically rigorous FEP and our statistical approach is high.
-For all 11 complexes tested, the correlation between max force and FEP was r=-0.767 at p=0.005857 (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was r=--0.729 at p=0.011.
-As a result of the strong correlation, we will generally refer to the SMD differences only.
+For all 11 complexes tested, the correlation between max force and FEP was $r=-0.767$ at $p=0.005857$ (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was $r=--0.729$ at $p=0.011$.
+Because of this strong correlation, we refer exclusively to the SMD results for the remainder of this work.
 
@@
 
-We would like to emphasize here that we cannot expect perfect agreement between our simulations and the available experimental data, but the correspondence to a well established free energy method is bolsters our conclusions.
+We would like to emphasize here that we cannot expect perfect agreement between our simulations and the available experimental data, but the correspondence to a well established free energy method bolsters our conclusions.
 While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method.
@@
 
-More importantly, with no \textit{a priori} knowledge of the appropriate number of equilibration samples, the during of equilibration, the appropriate number of pulling runs, or the best pulling speed means the computational expense in our SMD protocol may not be commensurate with the information provided.
+More importantly, with no \textit{a priori} knowledge of the appropriate number of equilibration samples, the best duration of equilibration, the appropriate number of pulling runs, or the best pulling speed means the computational expense in our SMD protocol may not be commensurate with the information provided.
 For example, another all atom approach that makes calculations via short simulations of spatially restrained complexes has proven capable of generating relatively accurate binding affinities with less compute time than is required from our steering strategy \citep{Gumbart2013,Gumbart2013b}.
@@
 Indices 17-25 are the hTfR flexible loop.
-The plot shows the flexibility of each site is essentially independent of mutation, and two sites above 0.72~\AA\ are found on the receptor in the interface.}
+The plot shows the flexibility of each site is essentially independent of mutation, and two sites (indices 17 and 18) above 0.72~\AA\ are a part of the flexible loop in the free receptor.
+However, these two residues are not actually found in the protein--protein interface, but rather are almost completely solvent exposed with the virus bound.}
 \end{figure}
@@
 \centerline{\includegraphics[width=6in]{figures/fep_mutants.pdf}}
-\caption{\label{fig:fep}\title{Max force versus free energy perturbation} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex.
+\caption{\label{fig:fep}\title{Max force versus free energy perturbation.} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex.
 The WT complex for FEP was simply set to 0.0.
-The correlation between the two is r=-0.767 with p=0.005857.}
+The correlation between the two is $r=-0.767$ with $p=0.005857$.}
 \end{figure}

### 2014-01-06 paper.tex
Changed pretty much every table and figure as well as all of the scripts.

@@
 Each of the 50 independent trajectories per mutant produced a fairly noisy force curve.
-The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting a 50 frame interpolation window for averaging.
+The force curves for each mutant were smoothed either by using the smooth.spline and predict functions in R with default settings over all replicates as in Figure~\ref{fig:force_curve}.
 The latter approach was used for statistical analysis.
@@
 The correlation between the energetically rigorous FEP and our statistical approach is high.
-For all 11 complexes tested, the correlation between max force and FEP was $r=-0.767$ at $p=0.005857$ (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was $r=--0.729$ at $p=0.011$.
+For all 11 complexes tested, the correlation between max force and FEP was $r=-0.795$ at $p=0.0034$ (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was $r=-0.593$ at $p=0.055$.
 Because of this strong correlation, we refer exclusively to the SMD results for the remainder of this work.
@@
 The WT complex for FEP was simply set to 0.0.
-The correlation between the two is $r=-0.767$ with $p=0.005857$.}
+The correlation between the two is $r=-0.795$ with $p=0.0034$.}
 \end{figure}
@@
          N348W &                 No &                    - \\
-         Y211A &                 No &        No Expression \\
+        vR111A &                 No &     Diminished Entry \\
+   N348A/Y211A &                 No &                    - \\
+  vR111A/Y211A &                 No &                    - \\
          Y211D &                Yes &        No Expression \\
          Y211T &                 No &     Diminished Entry \\
-        vR111A &                 No &     Diminished Entry \\
-   N348A/Y211A &                 No &                    - \\
+         Y211A &                 No &        No Expression \\
    N348W/Y211A &                 No &                    - \\
-  vR111A/Y211A &                 No &                    - \\
+
     \hline
@@
       Mutation & $\mu_\text{MAF}$ (pN) & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ & $\Delta G$ (kcal/mol)\\ \hline
-            WT &                772.00 &              154.13 &         198794.7 &            66765.64 &  0.000\\
-         N348A &                799.33 &              154.36 &         202247.3 &            65175.78 & -2.149\\
-         N348K &                731.36 &              151.01 &         179087.5 &            55799.49 & +3.184\\
-         N348W &                732.13 &              131.21 &         192043.2 &            66916.36 & +3.033\\
-         Y211A &                684.03 &              126.69 &         153517.7 &            40889.12 & +2.526\\ 
-         Y211D &                875.44 &              116.46 &         234909.3 &            50494.11 & -2.760\\ 
-         Y211T &                861.17 &              158.48 &         231078.7 &            97569.29 & +0.876\\
-        vR111A &                754.39 &              103.25 &         185896.8 &            46408.74 & +0.467\\
-   N348A/Y211A &                762.96 &              144.81 &         173059.2 &            59898.10 & +5.203\\
-   N348W/Y211A &                617.21 &              163.33 &         147705.6 &            48548.61 & +8.207\\
-  vR111A/Y211A &                780.71 &              135.65 &         189919.1 &            48868.16 & -2.441\\
+            WT &                734.4856 &              131.6513 &         145460.4 &            60232.26 &  0.000\\
+         N348A &                748.5217 &              137.4864 &         133913.9 &            51078.64 & -2.149\\
+         N348K &                705.0707 &              108.5079 &         141084.4 &            54450.28 & +3.184\\
+         N348W &                697.3642 &              132.6436 &         136886.0 &            53796.44 & +3.033\\
+        vR111A &                713.8081 &              106.7374 &         136103.2 &            52070.85 & +0.466\\ 
+   N348A/Y211A &                703.7027 &              128.5866 &         113464.2 &            57451.62 & +5.203\\ 
+  vR111A/Y211A &                741.0642 &              131.6287 &         130070.6 &            47665.56 & -2.440\\
+         Y211D &                825.2586 &              115.4343 &         158878.7 &            63039.08 & +2.760\\
+         Y211T &                806.8593 &              136.5648 &         167110.7 &            78849.29 & +0.875\\
+         Y211A &                654.1138 &              108.5343 &         108090.0 &            43661.09 & +2.526\\
+   N348W/Y211A &                594.9044 &              134.8233 &         108984.2 &            45451.00 & +8.206\\
     \hline
@@
 \begin {table}[H]
-\caption{\label{tab:MAF_pairwise}Pairwise differences (row variable minus column variable) in mean maximum applied force.
-Bolded values are statistically significant at $p<0.05$.
-Values marked with * and ** are significant at $p<0.01$ and $p<0.001$, respectively.}
+\caption{\label{tab:MAF_pairwise}Pairwise differences (row variable minus column variable) in mean maximum applied force. Bolded values are statistically significant at $p<0.05$.}
 \begin{center}
@@
     \hline
-                & WT                  & N348A              & N348W              & N348K              & Y211A             & Y211D              & Y211T              & vR111A             & N348A/Y211A        & N348W/Y211A        \\ \hline
-    N348A       &  27.33              &                    &                    &                    &                   &                    &                    &                    &                    &                    \\
-    N348W       & -39.87              & \textbf{-67.20}    &                    &                    &                   &                    &                    &                    &                    &                    \\
-    N348K       & -40.64              & \textbf{-67.97}    & -0.77              &                    &                   &                    &                    &                    &                    &                    \\
-    Y211A       & \textbf{-87.97*}    & \textbf{-115.30**} & -48.10             & -47.33             &                   &                    &                    &                    &                    &                    \\
-    Y211D       & \textbf{103.44**}   & \textbf{76.12*}    & \textbf{143.31**}  & \textbf{144.08**}  & \textbf{191.42**} &                    &                    &                    &                    &                    \\
-    Y211T       & \textbf{89.17*}     & \textbf{61.85}     & \textbf{129.04**}  & \textbf{129.81**}  & \textbf{177.15**} & -14.27             &                    &                    &                    &                    \\
-   vR111A       & -17.76              & -44.94             & 22.26              & 23.03              & \textbf{70.36}    & \textbf{-121.05**} & \textbf{-106.78**} &                    &                    &                    \\
-    N348A/Y211A & -9.04               & -36.37             & 30.83              & 31.60              & \textbf{78.93}    & \textbf{-112.48**} & \textbf{-98.21*}   & 8.57               &                    &                    \\
-    N348W/Y211A & \textbf{-154.79**}  & \textbf{-182.12**} & \textbf{-114.92**} & \textbf{-114.15**} & \textbf{-66.2}    & \textbf{-258.23**} & \textbf{-243.96**} & \textbf{-137.18**} & \textbf{-145.75**} &                    \\
-   vR111A/Y211A &  8.71               & -18.12             & 48.58              & 49.35              & \textbf{96.68*}   & \textbf{-94.74*}   & \textbf{-80.46*}   & 26.31              & 0.59               & \textbf{163.50**} \\
+             & WT                      & N348A                   & N348W                   & N348K                   & vR111A                  & N348A/Y211A               & vR111A/Y211A            & Y211D                   & Y211T                   & Y211A  \\ \hline
+    N348A       & +14.036              &                         &                         &                         &                         &                           &                         &                         &                         &        \\
+    N348W       & -29.414              & -43.451                 &                         &                         &                         &                           &                         &                         &                         &        \\
+    N348K       & -37.121              & \textbf{-51.157}        & -7.7060                 &                         &                         &                           &                         &                         &                         &        \\
+   vR111A       & -20.677              & -34.713                 & +8.7370                 & +16.443                 &                         &                           &                         &                         &                         &        \\
+    N348A/Y211A & -30.782              & -44.819                 & -1.3670                 & +6.3380                 & -10.105                 &                           &                         &                         &                         &        \\
+   vR111A/Y211A & +6.5790              & -7.4570                 & +35.993                 & +43.700                 & +27.256                 & +37.361                   &                         &                         &                         &        \\
+    Y211D       & \textbf{+90.772}     & \textbf{+76.736}        & \textbf{+120.19}        & \textbf{+127.89}        & \textbf{+111.45}        & \textbf{+121.56}          & \textbf{+84.194}        &                         &                         &        \\
+    Y211T       & \textbf{+72.373}     & \textbf{+58.337}        & \textbf{+101.79}        & \textbf{+109.50}        & \textbf{+93.051}        & \textbf{+103.16}          & \textbf{+65.795}        & -18.399                 &                         &        \\
+    Y211A       & \textbf{-80.371}     & \textbf{-94.407}        & -50.956                 & -43.250                 & \textbf{-59.694}        & -49.588                   & \textbf{-86.950}        & \textbf{-171.1 4}       & \textbf{-152.75}        &        \\
+    N348W/Y211A & \textbf{-139.58}     & \textbf{-153.62}        & \textbf{-110.17}        & \textbf{+102.46}        & \textbf{-118.903}       & \textbf{-108.80}          & \textbf{+146.16}        & \textbf{+230.35}        & \textbf{-211.95}        & \textbf{-59.209}  \\
     \hline
@@
     \hline
-                & WT                      & N348A                   & N348W                   & N348K                   & Y211A                   & Y211D                     & Y211T                   & vR111A                  & N348A/Y211A             & N348W/Y211A \\ \hline
-    N348A       & 0.35                    &                         &                         &                         &                         &                           &                         &                         &                         &        \\
-    N348W       & 0.22                    & \textbf{0.012}          &                         &                         &                         &                           &                         &                         &                         &        \\
-    N348K       & 0.22                    & \textbf{0.012}          & 0.98                    &                         &                         &                           &                         &                         &                         &        \\
-    Y211A       & \textbf{0.0045}         & \textbf{1.6x10$^{-05}$} & 0.14                    & 0.15                    &                         &                           &                         &                         &                         &        \\
-    Y211D       & \textbf{0.00083}        & \textbf{0.0045}         & \textbf{3.9x10$^{-06}$} & \textbf{3.9x10$^{-06}$} & \textbf{5.7x10$^{-10}$} &                           &                         &                         &                         &        \\
-    Y211T       & \textbf{0.0042}         & \textbf{0.022}          & \textbf{3.0x10$^{-05}$} & \textbf{3.0x10$^{-05}$} & \textbf{1.0x10$^{-08}$} & 0.67                      &                         &                         &                         &        \\
-   vR111A       & 0.59                    & 0.11                    & 0.51                    & 0.51                    & \textbf{0.024}          & \textbf{8.9x10$^{-05}$}   & \textbf{0.00056}        &                         &                         &        \\
-    N348A/Y211A & 0.78                    & 0.20                    & 0.35                    & 0.35                    & \textbf{0.011}          & \textbf{0.00026}          & \textbf{0.0016}         & 0.78                    &                         &        \\
-    N348W/Y211A & \textbf{6.2x10$^{-07}$} & \textbf{9.3x10$^{-12}$} & \textbf{0.00021}        & \textbf{0.00026}        & \textbf{0.032}          & \textbf{$<$ 2x10$^{-16}$} & \textbf{2.4x10$^{-15}$} & \textbf{9.1x10$^{-06}$} & \textbf{2.8x10$^{-06}$} &        \\
-   vR111A/Y211A & 0.78                    & 0.52                    & 0.14                    & 0.14                    & \textbf{0.0019}         & \textbf{0.0024}           & \textbf{0.010}          & 0.44                    & 0.59                    & \textbf{1.6x10$^{-07}$}  \\
+                & WT                      & N348A                   & N348W                   & N348K                   & vR111A                  & N348A/Y211A               & vR111A/Y211A            & Y211D                   & Y211T                   & Y211A  \\ \hline
+    N348A       & 0.60                    &                         &                         &                         &                         &                           &                         &                         &                         &        \\
+    N348W       & 0.31                    & 0.077                   &                         &                         &                         &                           &                         &                         &                         &        \\
+    N348K       & 0.20                    & \textbf{0.038}          & 0.81                    &                         &                         &                           &                         &                         &                         &        \\
+   vR111A       & 0.51                    & 0.16                    & 0.79                    & 0.60                    &                         &                           &                         &                         &                         &        \\
+    N348A/Y211A & 0.29                    & 0.07                    & 0.95                    & 0.81                    & 0.77                    &                           &                         &                         &                         &        \\
+   vR111A/Y211A & 0.82                    & 0.79                    & 0.21                    & 0.13                    & 0.35                    & 0.20                      &                         &                         &                         &        \\
+    Y211D       & \textbf{0.00093}        & \textbf{0.0012}         & \textbf{1.4x10$^{-5}$}  & \textbf{5.0x10$^{-6}$}  & \textbf{5.6x10$^{-5}$}  & \textbf{1.2x10$^{-5}$}    & \textbf{0.0022}         &                         &                         &        \\
+    Y211T       & \textbf{0.01}           & \textbf{0.018}          & \textbf{0.00022}        & \textbf{8.7x10$^{-5}$}  & \textbf{0.0008}         & \textbf{0.0002}           & \textbf{0.021}          & 0.56                    &                         &        \\
+    Y211A       & \textbf{0.0034}         & \textbf{7.2x10$^{-05}$} & 0.074                   & 0.13                    & \textbf{0.035}          & 0.079                     & \textbf{0.0016}         & \textbf{4.2x10$^{-10}$} & \textbf{4.2x10$^{-8}$}  &        \\
+    N348W/Y211A & \textbf{3.9x10$^{-7}$}  & \textbf{1.1x10$^{-10}$} & \textbf{6.5x10$^{-5}$}  & \textbf{0.00021}        & \textbf{1.6x10$^{-5}$}  & \textbf{7.2x10$^{-5}$}    & \textbf{1.3x10$^{-7}$}  & \textbf{< 2x10$^{-16}$} & \textbf{2.0x10$^{-14}$} & \textbf{0.036}  \\
     \hline
@@
     \hline
-                & WT               & N348A                   & N348W           & N348K                   & Y211A                   & Y211D                   & Y211T                   & vR111A          & N348A/Y211A & N348W/Y211A \\ \hline
-    N348A       & 0.77             &                         &                 &                         &                         &                         &                         &                 &             &        \\
-    N348W       & 0.67             & 0.42                    &                 &                         &                         &                         &                         &                 &             &        \\
-    N348K       & 0.17             & 0.056                   & 0.38            &                         &                         &                         &                         &                 &             &        \\
-    Y211A       & \textbf{0.00094} & \textbf{3.5x10$^{-05}$} & \textbf{0.0045} & 0.063                   &                         &                         &                         &                 &             &        \\
-    Y211D       & \textbf{0.0072}  & \textbf{0.0051}         & \textbf{0.0016} & \textbf{4.6x10$^{-05}$} & \textbf{1.2x10$^{-09}$} &                         &                         &                 &             &        \\
-    Y211T       & \textbf{0.016}   & \textbf{0.014}          & \textbf{0.0041} & \textbf{0.00016}        & \textbf{6.1x10$^{-09}$} & 0.77                    &                         &                 &             &        \\
-   vR111A       & 0.38             & 0.18                    & 0.68            & 0.67                    & \textbf{0.016}          & \textbf{0.00032}        & \textbf{0.00094}        &                 &             &        \\
-    N348A/Y211A & 0.061            & \textbf{0.013}          & 0.18            & 0.68                    & 0.17                    & \textbf{5.2x10$^{-06}$} & \textbf{2.1$^{-05}$}    & 0.38            &             &        \\
-    N348W/Y211A & \textbf{0.00017} & \textbf{3.9x10$^{-06}$} & \textbf{0.0011} & \textbf{0.021}          & 0.68                    & \textbf{1.6x10$^{-10}$} & \textbf{6.3x10$^{-10}$} & \textbf{0.0047} & 0.063       &        \\
-   vR111A/Y211A & 0.56             & 0.34                    & 0.86            & 0.47                    & \textbf{0.0072}         & \textbf{0.0010}         & \textbf{0.0025}         & 0.77            & 0.24        & \textbf{0.0019}  \\
+                & WT                      & N348A                   & N348W                   & N348K                   & vR111A                  & N348A/Y211A               & vR111A/Y211A            & Y211D                   & Y211T                   & Y211A  \\ \hline
+    N348A       & 0.33                    &                         &                         &                         &                         &                           &                         &                         &                         &        \\
+    N348W       & 0.76                    & 0.59                    &                         &                         &                         &                           &                         &                         &                         &        \\
+    N348K       & 0.59                    & 0.80                    & 0.76                    &                         &                         &                           &                         &                         &                         &        \\
+   vR111A       & 0.55                    & 0.85                    & 0.76                    & 0.94                    &                         &                           &                         &                         &                         &        \\
+    N348A/Y211A & \textbf{0.017}          & 0.07                    & \textbf{0.031}          & 0.076                   & 0.08                    &                           &                         &                         &                         &        \\
+   vR111A/Y211A & 0.26                    & 0.76                    & 0.46                    & 0.68                    & 0.72                    & 0.22                      &                         &                         &                         &        \\
+    Y211D       & 0.33                    & \textbf{0.029}          & 0.18                    & 0.09                    & 0.08                    & \textbf{0.00046}          & \textbf{0.029}          &                         &                         &        \\
+    Y211T       & 0.09                    & \textbf{0.0056}         & \textbf{0.046}          & \textbf{0.027}          & \textbf{0.023}          & \textbf{4.1x10$^{-5}$}    & \textbf{0.006}          & 0.59                    &                         &        \\
+    Y211A       & \textbf{0.0056}         & \textbf{0.027}          & \textbf{0.016}          & \textbf{0.029}          & \textbf{0.031}          & 0.75                      & 0.09                    & \textbf{8.2x10$^{-5}$}  & \textbf{8.5x10$^{-6}$}  &        \\
+    N348W/Y211A & \textbf{0.006}          & \textbf{0.029}          & \textbf{0.017}          & \textbf{0.032}          & \textbf{0.034}          & 0.76                      & 0.1                     & \textbf{9.4x10$^{-5}$}  & \textbf{8.5x10$^{-6}$}  & 0.94  \\
     \hline

### 2014-01-06 smd_pro/all_analysis.R

@@ -1,0 +1,315 @@
+rm(list = ls())
+
+require(plyr)
+
+scatter.plot <- function(df, name) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=26
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
+                            axis.line=element_line(),
+                            panel.grid.minor=element_blank(),
+                            panel.grid.major=element_blank(),
+                            panel.background=element_blank(),
+                            panel.border=element_blank(),
+                            axis.line=element_line())
+  
+  #print(centers.mclust)
+  g <- ggplot()
+  g <- g + geom_point(data=df, aes(x=distance, y=force, color=id), size=2.5) + scale_colour_manual(values=c("black", "#FF9999"))
+  g <- g + theme(strip.background=element_blank())
+  g <- g + ylab("Force (pN)")
+  g <- g + xlab("Distance (A)")
+  g <- g + scale_y_continuous(breaks=seq(-150, 1500, 150), limits=c(-15, 750))
+  g <- g + scale_x_continuous(breaks=seq(0, 12, 2), limits=c(0, 10))
+  g <- g + theme(panel.border = element_blank(), axis.line = element_line())
+  g <- g + theme(axis.title.x = element_text(size=the_pointsize, vjust=-.2))
+  g <- g + theme(axis.title.y = element_text(angle=90, size=the_pointsize, vjust=.3))
+  g <- g + theme(axis.line = element_line(colour = 'black', size = 1))
+  g <- g + theme(axis.ticks = element_line(colour = 'black', size = 1))
+  g <- g + theme(plot.margin=unit(c(1, 1, 1, 1), "lines"))
+  g <- g + theme(axis.ticks.margin = unit(0.15, "cm"))
+  g <- g + theme(legend.position = "none") 
+  ggsave(g, file=paste("~/Desktop/MACV_SMD/figures/",name,sep=''), width=7, height=5)
+}
+
+diff.plot <- function(diff.df) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=22
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
[…]

### 2014-01-06 smd_pro/data_wt_old/distance/first_analysis.R

@@ -1,6 +1,0 @@
-setwd("~/Desktop/data")
-
-for(i in 0:50) {
-  distance <- read.table(paste("distances/distance_", i, ".dat", sep=""))[2]
-  #force <- read.table(paste("forces/force_", i, ".dat", sep=""))[2]
-}

### 2014-01-06 smd_pro/force_curve.R

@@ -1,0 +1,287 @@
+rm(list = ls())
+
+require(plyr)
+
+scatter.plot <- function(df, name, axes, centers.mclust) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=26
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
+                            axis.line=element_line(),
+                            panel.grid.minor=element_blank(),
+                            panel.grid.major=element_blank(),
+                            panel.background=element_blank(),
+                            panel.border=element_blank(),
+                            axis.line=element_line())
+  
+  #print(centers.mclust)
+  g <- ggplot()
+  g <- g + geom_point(data=df, aes(x=distance, y=force), size=2.5)
+  
+  g <- g + theme(strip.background=element_blank())
+  g <- g + ylab("Force")
+  g <- g + xlab("Distance (Angstrom")
+  g <- g + scale_y_continuous(limits=c(-1, 1))
+  g <- g + scale_x_continuous(limits=c(-1, 1))
+  g <- g + theme(panel.border = element_blank(), axis.line = element_line())
+  g <- g + theme(axis.title.x = element_text(size=the_pointsize, vjust=-.2))
+  g <- g + theme(axis.title.y = element_text(angle=90, size=the_pointsize, vjust=.3))
+  g <- g + theme(axis.line = element_line(colour = 'black', size = 1))
+  g <- g + theme(axis.ticks = element_line(colour = 'black', size = 1))
+  g <- g + theme(plot.margin=unit(c(1, 1, 1, 1), "lines"))
+  g <- g + theme(axis.ticks.margin = unit(0.15, "cm"))
+  g <- g + theme(legend.position = "none") 
+  ggsave(g, file=name, width=7, height=5)
+}
+
+diff.plot <- function(diff.df) {
+  require(ggplot2)
+  require(grid)
+  
+  graphics.off()
+  the_pointsize=22
+  theme_set(theme_bw(base_size=the_pointsize))
+  old_theme <- theme_update(panel.border=element_blank(),
+                            axis.line=element_line(),
+                            panel.grid.minor=element_blank(),
[…]

### 2014-01-06 smd_pro/smd_analysis.R

@@ -1,6 +1,6 @@
 rm(list = ls())
 
-setwd('~/Google Drive/Documents/Biochem/Research/SMD_manuscript/MACV_SMD/smd_pro/')
+setwd('~/Desktop/MACV_SMD/smd_pro/')
 
 diff.plot <- function(diff.df) {
   require(ggplot2)
@@ -199,7 +199,7 @@ df$id <- factor(df$id, c('WT', 'N348A', 'N348K', 'N348W',
                          'R111A', 'N348A/Y211A', 'vR111A/Y211A', 
                          'Y211D', 'Y211T', 'Y211A', 'N348W/Y211A'))
 
-smd.boxplot(df)
+#smd.boxplot(df)
 
 
 #fit <- aov(y1 + y2 ~ id, data = df)

### 2014-01-06 paper.tex
Added citations to fep methods.

@@
 
-Briefly, we used the traditional dual topology approach to FEP.
+Briefly, we used the traditional dual topology approach to FEP citep{Gao2006,Pearlman1994}.
 This involves a thermodynamic cycle where a set of atoms are progressively decoupled from the environment while another set of atoms are progressively coupled.
 To compute the relative free energy difference requires knowing the free energy change when the transformation is carried out for the bound complex and the individual protein.
-Then, one can compute the relative free energy difference between a WT and mutant complex by taking the difference between the energy required to decouple/couple the atoms in solution from the energy required to decouple/couple the atoms in the bound complex.
+Then, one can compute the relative free energy difference between a WT and mutant complex by taking the difference between the energy required to decouple/couple the atoms in solution from the energy required to decouple/couple the atoms in the bound complex citep{Gao2006,Pearlman1994}.
 
@@
 
-The FEP analysis tool in VMD was used for subsequent analysis.
+The VMD mutator tool was used to generate the necessary topology file and the parseFEP tool \citep{Liu2012} in VMD was used for subsequent analysis.
 We used it to perform error analysis and compute the Bennett acceptance ratio as the maximum likelihood free energy difference of the two states under consideration.

### 2014-01-06 namd_config/equilibration/execute_vmd.sh
Adding more scripts.

@@ -1,0 +1,1 @@
+vmd -dispdev text -eofexit < $1

### 2014-01-06 namd_config/equilibration/job.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+for i in {1..20};
+do
+  results="run_${i}"
+
+  cp qsub_eq.sh qsub_eq_current.sh
+
+  sed -i 's/unique_name/'$results'/g' qsub_eq_current.sh
+
+  ./qsub_eq_current.sh
+done

### 2014-01-06 namd_config/equilibration/qsub_eq.sh

@@ -1,0 +1,46 @@
+#!/bin/csh
+
+# Create Working Directory
+set WDIR = $RESULTS/wt_eq/unique_name
+set FDIR = `pwd`
+
+if ( -d $WDIR ) then
+  rm -r $WDIR
+endif
+
+mkdir -p $WDIR
+
+if ( ! -d $WDIR ) then
+  echo $WDIR not created
+  exit
+endif
+
+cd $WDIR
+
+# Copy Data and Config Files
+cp $FDIR/* .
+
+set MPIRUN = /lustre/work/apps/openmpi/bin/mpirun
+set NAMD2  = /lustre/work/apps/NAMD_2.9b3_Source/Linux-x86_64-icc/namd2
+
+echo '#\!/bin/bash' > run.q
+echo '#$ -cwd' >> run.q
+echo '#$ -V' >> run.q
+echo '#$ -S /bin/bash' >> run.q
+echo '#$ -N eq_1' >> run.q
+echo '#$ -j y' >> run.q
+echo '#$ -o wt.log' >> run.q
+echo '#$ -e wt.err' >> run.q
+echo '#$ -q normal' >> run.q
+echo '#$ -pe fill 60' >> run.q
+echo '#$ -P hrothgar\n' >> run.q
+
+echo '# Command to run\n' >> run.q
+echo 'unset SGE_ROOT\n' >> run.q
+
+echo 'ldd /lustre/work/apps/NAMD_2.9b3_Source/Linux-x86_64-icc/namd2\n' >> run.q
+
+echo  $MPIRUN -np 60 -machinefile 'machinefile.$JOB_ID' $NAMD2 wt.namd >> run.q
+
+qsub run.q
+

### 2014-01-06 namd_config/smd/go_cv.py

@@ -1,0 +1,27 @@
+import math, random, subprocess, decimal
+
+def main():
+  for i in range(0, 10):
+    filename = "wt_smd.namd"
+
+    fixed_file = "set_fixed_atom.tcl"
+    eq_run = random.randint(1,20)
+
+    run_to_use = "run_" + str(eq_run)
+
+    command = "cp setup_smd.sh setup_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+    command = "cp qsub_smd.sh qsub_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+    command = "sed -i 's/run_unique/" + run_to_use + "/g' setup_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+    command = "sed -i 's/unique_name/run_" + str(i) + "/g' qsub_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+    command = "sh setup_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+main()

### 2014-01-06 namd_config/smd/qsub_smd.sh

@@ -1,0 +1,50 @@
+#!/bin/csh
+
+# Create Working Directory
+set WDIR = $RESULTS/6_smd/unique_name
+set FDIR = `pwd`
+
+if ( -d $WDIR ) then
+  rm -r $WDIR
+endif
+
+mkdir -p $WDIR
+
+if ( ! -d $WDIR ) then
+  echo $WDIR not created
+  exit
+endif
+
+cd $WDIR
+
+# Copy Data and Config Files
+cp $FDIR/* .
+
+set MPIRUN = /lustre/work/apps/openmpi/bin/mpirun
+set NAMD2  = /lustre/work/apps/NAMD_2.9b3_Source/Linux-x86_64-icc/namd2
+
+## qsub << ENDINPUT
+
+echo '#\!/bin/bash' > run.q
+echo '#$ -cwd' >> run.q
+echo '#$ -V' >> run.q
+echo '#$ -S /bin/bash' >> run.q
+echo '#$ -N smd_6' >> run.q
+echo '#$ -j y' >> run.q
+echo '#$ -o wt_smd.log' >> run.q
+echo '#$ -e wt_smd.err' >> run.q
+echo '#$ -q normal' >> run.q
+echo '#$ -pe fill 60' >> run.q
+echo '#$ -P hrothgar\n' >> run.q
+
+echo '# Command to run\n' >> run.q
+echo 'unset SGE_ROOT\n' >> run.q
+
+echo 'ldd /lustre/work/apps/NAMD_2.9b3_Source/Linux-x86_64-icc/namd2\n' >> run.q
+
+echo  $MPIRUN -np 60 -machinefile 'machinefile.$JOB_ID' $NAMD2 wt_smd.namd >> run.q
+
+qsub run.q
+
+#ENDINPUT
+

### 2014-01-06 namd_config/smd/setup_smd.sh

@@ -1,0 +1,33 @@
+  smd_conf="wt_smd.namd"
+  pull_directory="run_unique";
+  smd_set="set_fixed_atom.tcl";
+  qsub_file="qsub_smd_current.sh";  
+  query_script="query_distance.py";
+
+  if [ -d "$pull_directory" ]; then
+    
+    if [ -f $pull_directory/$qsub_file ]; then
+      rm $pull_directory/$qsub_file
+    fi
+    
+    if [ -f $pull_directory/$smd_conf ]; then
+      rm $pull_directory/$smd_conf
+    fi
+    
+    if [ -f $pull_directory/$smd_set ]; then
+      rm $pull_directory/$smd_set
+    fi    
+
+    cp $smd_conf       $pull_directory
+    cp $qsub_file      $pull_directory
+    cp $smd_set        $pull_directory
+    cp $query_script   $pull_directory
+    
+    cd $pull_directory;
+
+    sh execute_vmd.sh $smd_set
+
+    ./$qsub_file;
+
+    cd ../; 
+  fi

### 2014-01-06 paper.tex
Updated to fix 3 of the 4 remaining changes.

@@
 \usepackage{lineno}
+\usepackage{color}
 
@@
 Then, we measure the force required for dissociation \citep{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
-As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
 Although such biasing techniques are commonly used in protein-ligand binding problems, they are less commonly applied to protein--protein interactions, and almost never to mutational analysis in a protein--protein system.
-This is largely the result of free energy convergence difficulties and computational limitations \citep{Cuendet2008,Cuendet2011}; using maximum force or simple AUC can solve these problems.
+This is largely the result of free energy convergence difficulties and computational limitations \citep{Cuendet2008,Cuendet2011}.
+Using a proxy for relative binding affinity rather than caluclating absolute affinities can solve these problems.
+Here, as proxies, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
 For comparison, we also calculate relative free energy differences using the traditional dual topology FEP paradigm, and we show that the two approaches yield congruent results.
@@
 Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}.
-The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid $\beta$-pleated sheet domain.
+The co-crystal structure shows that the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid $\beta$-pleated sheet domain.
 For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
@@
 Finally, we calculated the system center and dimensions for use in molecular dynamics settings.
+The exact NAMD configuration files are available on github (https://github.com/clauswilke/MACV\_SMD).
 
-The exact NAMD configuration file is made available as a github repository with all necessary set up files is provided.
 We used the Charmm27 \citep{Brooks1983} all-atom force field.
@@
 The center of mass of the $\alpha$-carbons of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation.
-The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms, but rather uses the selection to compute an initial center of mass.
+The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms but rather uses the selection to compute an initial center of mass.
 Then, during the steering run, the single center of mass point is pulled with the parameters desribed below.
@@
 A force constant of 5~kcal/mol/\AA$^2$ was chosen due to its relatively low signal-to-noise ratio.
-This constant is slightly lower than the more common 7~kcal/mol/\AA$^2$ found in several recent studies; that value is commonly selected primarily because it is the force constant found in the NAMD, SMD tutorial available through the developers.
+This constant is slightly lower than the more common 7~kcal/mol/\AA$^2$ found in several recent studies; that value is commonly selected primarily because it is the force constant found in the SMD tutorial available through the NAMD  developers.
 Moreover, the force constant could very likely be set to a range of nearby values with little loss in predictive power.
-In SMD experiments the pulling velocity should be as low as possible for the available computational time \citet{Cuendet2008,Cuendet2011}.
-Thus we choose a velocity of 0.000001~\AA/fs = 1~\AA/ns, and direction down the positive $z$-axis.
-This number could be set to pull faster if the computing time must be reduced, but slower than necessary pulling speeds are not typically considered problematic.
+
+In SMD experiments the pulling velocity should be as low as possible for the available computational time \citep{Cuendet2008,Cuendet2011}.
+We choose a velocity of 0.000001~\AA/fs = 1~\AA/ns, and direction down the positive $z$-axis.
+One could use faster pulling if the computing time must be reduced, but slower than necessary pulling speeds are not typically considered problematic.
 
@@
 
-Briefly, we used the traditional dual topology approach to FEP citep{Gao2006,Pearlman1994}.
+Briefly, we used the traditional dual topology approach to FEP \citep{Gao2006,Pearlman1994}.
 This involves a thermodynamic cycle where a set of atoms are progressively decoupled from the environment while another set of atoms are progressively coupled.
 To compute the relative free energy difference requires knowing the free energy change when the transformation is carried out for the bound complex and the individual protein.
-Then, one can compute the relative free energy difference between a WT and mutant complex by taking the difference between the energy required to decouple/couple the atoms in solution from the energy required to decouple/couple the atoms in the bound complex citep{Gao2006,Pearlman1994}.
+Then, one can compute the relative free energy difference between a WT and mutant complex by taking the difference between the energy required to decouple/couple the atoms in solution from the energy required to decouple/couple the atoms in the bound complex \citep{Gao2006,Pearlman1994}.
 
-Again, the NAMD configuration file is made available via github.
+Again, the NAMD configuration file is made available via github \\ (https://github.com/clauswilke/MACV\_SMD).
 We used a similar configuration to that in equilibration.
@@
 Each of the 50 independent trajectories per mutant produced a fairly noisy force curve.
-The force curves for each mutant were smoothed either by using the smooth.spline and predict functions in R with default settings over all replicates as in Figure~\ref{fig:force_curve}.
-The latter approach was used for statistical analysis.
+The force curves for each mutant were smoothed by using the smooth.spline() and predict() functions in R with default settings over all replicates as in Figure~\ref{fig:force_curve}.
 The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC).
-We used the pairwise.t.test function with the FDR p-value correction to test for significance.
+We tested for signifiant differences in maximum force or AUC by carrying out t tests for all pairwise combinations (each mutant compared to each other mutant), using the pairwise.t.test() function in R.
+We adjusted p values to correct for multiple testing using the False-Discovery-Rate (FDR) method \citep{BenjaminiHochberg1995}.
 The ggplot \citep{ggplot} package was used to generate most of the figures.
+
+Analysis scripts and final data (except MD trajectories) are available on the github repository accompanying this publication (https://github.com/clauswilke/MACV\_SMD).
 
@@
 Despite the fact that viral binding occurs at the site of a flexible loop in the free hTfR structure, our data shows after binding the strand is extremely rigid.
-In the bound conformation, two sites of the loop have root mean squared fluctuation (RMSF) values in the top half of all receptor sites during equilibration (Figure~\ref{fig:rmsf}), and those are almost completely exposed to solvent.
+In the bound conformation, only two sites of the loop have root mean squared fluctuation (RMSF) values in the top half of all receptor sites during equilibration (Figure~\ref{fig:rmsf}), and those are almost completely exposed to solvent.
 This is unsurprising considering the high degree of burial that occurs as a result of viral binding.
@@
 In fact, the RMSD over trajectory looks highly invariant across mutants (Figure~\ref{fig:rmsd}).
-By comparison to the equilibration RMSF calculation, near the end of the SMD trajectory all of the residues in the WT receptor interfacial strand are in the top half of RMSF over all receptor sites (Figure~\ref{fig:rmsf_comparison}).
+In the unbound state, calculated near the end of the SMD trajectory, all of the residues in the WT receptor interfacial strand are in the top half of RMSF over all receptor sites (Figure~\ref{fig:rmsf_comparison}).
 Thus, if sufficient simulation time is not dedicated to allowing this unfolding process, standard free energy techniques may miss the energetic contributions that result from ordering the flexible loop in the hTfR apical domain.
@@
 
-We proceeded to analyze this system using two molecular dynamics techniques.
-First, in SMD by using a known force constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
+We analyzed the GP1/hTfR1 system using two molecular dynamics techniques.
+First, by carrying out SMD using a known force constant and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
 A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}.
-As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants.
-The supplementary movie visually illustrates the separation distance between peptide domains.
-The quantities maximum applied force and AUC were derived from the force versus distances curves.
+As seen in Figure~\ref{fig:force_curve}, the dissociation distance was relatively consistent among mutants.
+{\color{red}\emph{I would still like to see an array of figures with all averaged force curves.}} The supplementary movie visually illustrates the separation distance between peptide domains.
+The quantities maximum applied force and AUC were derived from the force-versus-distances curves.
 Their summary statistics are reported in Table~\ref{tab:summary_statistics}.
@@
 For all 11 complexes tested, the correlation between max force and FEP was $r=-0.795$ at $p=0.0034$ (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was $r=-0.593$ at $p=0.055$.
-Because of this strong correlation, we refer exclusively to the SMD results for the remainder of this work.
+Because of the strong correlation, we refer exclusively to the SMD results for the remainder of this work, focusing primarily on max force.
 
@@
 When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:MAF_pairwise} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}).
-In general, however, maximum applied force seemed to be the more sensitive statistic than AUC.
+In general, however, and consistent with the FEP results, maximum applied force seemed to be the more sensitive statistic than AUC.
 
@@
 Last, we further analyzed a single mutation in GP1, vArg111Ala.
-As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either maximum applied force or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}).
+As mentioned previously, in our simulations this mutant showed no significant change in either maximum applied force or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}).
 This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity.
@@
 
-There are a few additional challenges for investigating host-virus interaction via molecular dynamics simulation.
-As with any atomistic simulation, there is going to be a fairly large noise to signal ratio.
+There are a few additional challenges for investigating host-virus interactions via molecular dynamics simulation.
+As with any atomistic simulation, there is going to be a fairly large noise-to-signal ratio.
 To reduce noise, one could further customize each simulation, e.g.
@@
 Such an approach eventually requires using the Jarzynski inequality \citep{Jar1997} to establish a lower limit for binding energy and would quickly become computationally inefficient for evaluating a large number of mutations in most systems.
-However, considering the strong correlation between FEP and SMD in this system, it may not be important to ensure one's results are path independent for relative affinity calculations.
+However, considering the strong correlation between FEP and SMD in this system, it may not be important to ensure one's results are path independent for relative affinity calculations, as long as the same path is used for all complexes.
 
@@
 That being said, there is no reason to believe this SMD approach to mutagenic studies could not be optimized to reduce computational expense.
-Further analysis would be needed to understand the lower limits of resources required for accurate predictions.
+Further analysis will be needed to understand the lower limits of resources required for accurate predictions.
 
@@
 This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock \\ (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu).
-The authors acknowledge Bryan Sutton for opening access to the Hrothgar cluster.
-The authors also acknowledge the reviewers Ilan Samish and Matteo Masetti for their helpful comments in preparing this manuscript.
+We would like to thank Bryan Sutton for opening access to the Hrothgar cluster and the reviewers Ilan Samish and Matteo Masetti for their helpful comments on this work.
 
-\bibliographystyle{MBE}
+\bibliographystyle{peerj}
 \bibliography{paper}
@@
 The plot shows the flexibility of each site is essentially independent of mutation, and two sites (indices 17 and 18) above 0.72~\AA\ are a part of the flexible loop in the free receptor.
-However, these two residues are not actually found in the protein--protein interface, but rather are almost completely solvent exposed with the virus bound.}
+However, these two residues are not actually found in the protein--protein interface, but rather are almost completely solvent exposed with the virus bound.
+{\color{red}\emph{Please highlight the flexible loop, as in Fig.
+4.}}}
 \end{figure}

### 2014-01-06 paper.tex
One more fix.

@@
 Before systematically applying SMD to the GP1/hTfR1 interaction, we needed to ensure the method was sufficiently sensitive to distinguish between relatively minor point mutations.
-While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
+While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, it is rarely used to parse small energy differences in a protein--protein interaction energy landscape.
 For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.

### 2014-01-06 smd_pro/all_analysis.R
Fixed one force curve color.

@@ -264,6 +264,7 @@ scatter.plot(data.frame(id=rep('1',length(rowMeans(split.distance.nk))), distanc
 scatter.plot(data.frame(id=rep('1',length(rowMeans(split.distance.ra))), distance=rowMeans(split.distance.ra), force=rowMeans(split.force.ra)), "force_curves/ra.pdf")
 scatter.plot(data.frame(id=rep('1',length(rowMeans(split.distance.na.ya))), distance=rowMeans(split.distance.na.ya), force=rowMeans(split.force.na.ya)), "force_curves/na.ya.pdf")
 scatter.plot(data.frame(id=rep('1',length(rowMeans(split.distance.ra.ya))), distance=rowMeans(split.distance.ra.ya), force=rowMeans(split.force.ra.ya)), "force_curves/ra.ya.pdf")
+scatter.plot(data.frame(id=rep('1',length(rowMeans(split.distance.nw.ya))), distance=rowMeans(split.distance.nw.ya), force=rowMeans(split.force.nw.ya)), "force_curves/nw.ya.pdf")
 
 tmp.scatter.wt <- data.frame(id=rep('1',length(rowMeans(split.distance.wt))), distance=rowMeans(split.distance.wt), force=rowMeans(split.force.wt))
 tmp.scatter.ya <- data.frame(id=rep('2',length(rowMeans(split.distance.ya))), distance=rowMeans(split.distance.ya), force=rowMeans(split.force.ya))

### 2014-01-06 paper.tex
copy edit

--- 2014-01-06 paper.tex
@@
 
-Again, the NAMD configuration file is made available via github \\ (https://github.com/clauswilke/MACV\_SMD).
+Again, the NAMD configuration file is made available via github (https://github.com/clauswilke/MACV\_SMD).
 We used a similar configuration to that in equilibration.
@@
 Each of the 50 independent trajectories per mutant produced a fairly noisy force curve.
-The force curves for each mutant were smoothed by using the smooth.spline() and predict() functions in R with default settings over all replicates as in Figure~\ref{fig:force_curve}.
-The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC).
+{\color{red}\emph{This is not entirely clear.
+I thought there is only one approach now.} The force curves for each mutant were smoothed either by using the smooth.spline() and predict() functions in R with default settings over all replicates as in Figure~\ref{fig:force_curve}.
+The latter approach was used for statistical analysis.} The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC).
 We tested for signifiant differences in maximum force or AUC by carrying out t tests for all pairwise combinations (each mutant compared to each other mutant), using the pairwise.t.test() function in R.
@@
 While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
-For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.
+{\color{red}\emph{$\leftarrow$ Is this still a correct statement?
+I know we deleted a novelty claim earlier in the ms.}} For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.
 

### 2014-01-06 paper.tex
Last update before sending to Sara and Andy.

--- 2014-01-06 paper.tex
@@
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
-Then, we measure the force required for dissociation \citep{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
+Then, we measure the force required for dissociation \citep{Lu1999,Is2001A,Is2001B,Park2004,Gumbart2012,Minoetal2013}.
 Although such biasing techniques are commonly used in protein-ligand binding problems, they are less commonly applied to protein--protein interactions, and almost never to mutational analysis in a protein--protein system.
@@
 The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms but rather uses the selection to compute an initial center of mass.
-Then, during the steering run, the single center of mass point is pulled with the parameters desribed below.
+Then, during the steering run, the single center of mass point is pulled with the parameters described below.
 We used the same force field parameters (exclude, cutoff, switching, etc.), the same integrator parameters (time step, rigidbonds on, all molecular being wrapped, etc.), and the same particle mesh ewald parameters as in equilibration.
@@
 Each of the 50 independent trajectories per mutant produced a fairly noisy force curve.
-The force curves for each mutant were smoothed by using the smooth.spline() and predict() functions in R with default settings over all replicates as in Figure~\ref{fig:force_curve}.
+The force curves for each mutant were smoothed over all replicates by using the smooth.spline() and predict() functions in R with default settings.
 The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC).
@@
 First, by carrying out SMD using a known force constant and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
-A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}.
+A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}, and individual images of all averaged force curves are available in the associated github repository, in folder figures/force\_curves.
 As seen in Figure~\ref{fig:force_curve}, the dissociation distance was relatively consistent among mutants.
-{\color{red}\emph{I would still like to see an array of figures with all averaged force curves.}} The supplementary movie visually illustrates the separation distance between peptide domains.
+The supplementary movie visually illustrates the separation distance between peptide domains.
 The quantities maximum applied force and AUC were derived from the force-versus-distances curves.
@@
 \section{Acknowledgements}
-This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock \\ (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu).
+This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu).
 We would like to thank Bryan Sutton for opening access to the Hrothgar cluster and the reviewers Ilan Samish and Matteo Masetti for their helpful comments on this work.
@@
 The plot shows the flexibility of each site is essentially independent of mutation, and two sites (indices 17 and 18) above 0.72~\AA\ are a part of the flexible loop in the free receptor.
-However, these two residues are not actually found in the protein--protein interface, but rather are almost completely solvent exposed with the virus bound.
-{\color{red}\emph{Please highlight the flexible loop, as in Fig.
-4.}}}
+However, these two residues are not actually found in the protein--protein interface, but rather are almost completely solvent exposed with the virus bound.}
 \end{figure}

### 2014-01-07 namd_config/post_processing/post_processing.py
Added everything for smd.

@@ -1,0 +1,131 @@
+import sys, os, math, string, re, gzip, urllib, shutil, fnmatch
+import cStringIO
+from prody import *
+import numpy as np
+from numpy import arange, cos, linspace, pi, sin, random
+
+def main():
+ 
+  forces    = parse_log()
+  distances = parse_dcd()
+   
+## Strips whitespace from list elements
+def striplist(l):
+  return([x.strip() for x in l])
+
+## Python shenanigans to remove empty elements in a list
+## This function strips whitespace as well 
+def stripempties(l):
+  return([x.strip() for x in l if x != ''])
+            
+def parse_dcd():
+
+  pdb_file = ""
+  dcd_file = ""
+  log_file = ""
+        
+  ## Find input file names by pattern matching
+  for file in os.listdir('.'):
+    if fnmatch.fnmatch(file, '*ionized.coor'):
+    ##    print file
+        pdb_file = file
+    if fnmatch.fnmatch(file, '*pulled*.dcd'):
+    ##    print file
+        dcd_file = file
+    if fnmatch.fnmatch(file, '*smd*.log'):
+    ##    print file
+        log_file = file
+
+  structure = parsePDB( pdb_file )
+  traj      = Trajectory( dcd_file )
+  traj.setAtoms( structure )
+  
+  file_out  = "distances.dat"
+
+  tfr1 = structure.select('calpha and resnum 1 to 162')
+
+  macv = structure.select('calpha and resnum 163 to 318')
+ 
+  distances = np.zeros( traj.numFrames() )
+  
+  output_handle = open(file_out, 'w')
+
+  start_dist = 0
+
+  did_they_sep = False
+  sep_time = 0
+
+  for i, frame in enumerate(traj):
+    center_tfr1 = calcCenter(tfr1)
+    center_macv = calcCenter(macv)
+    
+    if( i == 0 ):
+      dist_3d = center_tfr1 - center_macv
+      start_dist = np.sqrt( dist_3d[0]**2 + dist_3d[1]**2 + dist_3d[2]**2 )
+
+    distance_3d = center_tfr1 - center_macv
+     
+    this_distance = np.sqrt( distance_3d[0]**2 + distance_3d[1]**2 + distance_3d[2]**2 )
+    
+    distances[i] = this_distance - start_dist
+    out_line = "%s %s\n" % ( i, distances[i] )
+    output_handle.write( out_line )
+
+    ## Test if the average seperation of the last 10 steps is greater than 5 A
[…]

### 2014-01-07 namd_config/post_processing/process.sh

@@ -1,0 +1,20 @@
+#!/bin/sh
+
+for i in {0..49};
+do
+  FOL=run_$i
+  if [ -d $FOL ]; 
+  then
+    SCP=post_processing.py
+    cp $SCP $FOL
+     
+    cd $FOL
+    
+    python $SCP
+
+    cp force*dat ../data/force/force_$i.dat
+    cp distance*dat ../data/distance/distance_$i.dat
+
+    cd ../
+  fi
+done

### 2014-01-07 namd_config/smd/go.py

@@ -1,0 +1,26 @@
+#!/usr/bin/python
+
+import math, random, subprocess, decimal
+
+def main():
+  for i in range(0, 50):
+    eq_run = random.randint(1,20)
+
+    run_to_use = "run_" + str(eq_run)
+
+    command = "cp setup_smd.sh setup_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+    command = "cp qsub_smd.sh qsub_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+    command = "sed -i 's/run_unique/" + run_to_use + "/g' setup_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+    command = "sed -i 's/unique_name/run_" + str(i) + "/g' qsub_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+    command = "sh setup_smd_current.sh"
+    subprocess.call(command, shell=True)
+
+main()

### 2014-01-07 namd_config/smd/go_cv.py

@@ -1,27 +1,0 @@
-import math, random, subprocess, decimal
-
-def main():
-  for i in range(0, 10):
-    filename = "wt_smd.namd"
-
-    fixed_file = "set_fixed_atom.tcl"
-    eq_run = random.randint(1,20)
-
-    run_to_use = "run_" + str(eq_run)
-
-    command = "cp setup_smd.sh setup_smd_current.sh"
-    subprocess.call(command, shell=True)
-
-    command = "cp qsub_smd.sh qsub_smd_current.sh"
-    subprocess.call(command, shell=True)
-
-    command = "sed -i 's/run_unique/" + run_to_use + "/g' setup_smd_current.sh"
-    subprocess.call(command, shell=True)
-
-    command = "sed -i 's/unique_name/run_" + str(i) + "/g' qsub_smd_current.sh"
-    subprocess.call(command, shell=True)
-
-    command = "sh setup_smd_current.sh"
-    subprocess.call(command, shell=True)
-
-main()

### 2014-01-07 namd_config/smd/qsub_smd.sh

@@ -1,7 +1,7 @@
 #!/bin/csh
 
 # Create Working Directory
-set WDIR = $RESULTS/6_smd/unique_name
+set WDIR = $RESULTS/wt_pro/unique_name
 set FDIR = `pwd`
 
 if ( -d $WDIR ) then
@@ -29,7 +29,7 @@ echo '#\!/bin/bash' > run.q
 echo '#$ -cwd' >> run.q
 echo '#$ -V' >> run.q
 echo '#$ -S /bin/bash' >> run.q
-echo '#$ -N smd_6' >> run.q
+echo '#$ -N hum_smd' >> run.q
 echo '#$ -j y' >> run.q
 echo '#$ -o wt_smd.log' >> run.q
 echo '#$ -e wt_smd.err' >> run.q

### 2014-01-07 namd_config/smd/setup_smd.sh

@@ -2,7 +2,6 @@
   pull_directory="run_unique";
   smd_set="set_fixed_atom.tcl";
   qsub_file="qsub_smd_current.sh";  
-  query_script="query_distance.py";
 
   if [ -d "$pull_directory" ]; then
     
@@ -18,10 +17,9 @@
       rm $pull_directory/$smd_set
     fi    
 
-    cp $smd_conf       $pull_directory
-    cp $qsub_file      $pull_directory
-    cp $smd_set        $pull_directory
-    cp $query_script   $pull_directory
+    cp $smd_conf  $pull_directory
+    cp $qsub_file $pull_directory
+    cp $smd_set   $pull_directory
     
     cd $pull_directory;
 

### 2014-01-09 full_paper.tex
Added version with abstract.

@@
 
-\title{\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}}
+\title{\textbf{Analyzing Machupo virus-receptor binding by molecular dynamics simulations}}
 \author{Austin G. Meyer$^{1,2,3*}$, Sara L. Sawyer$^{2}$, Andrew D. Ellington$^{2}$,\\and Claus O. Wilke$^{1}$}
@@
 \linenumbers
+
 \newpage
 \begin{abstract}
-Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases.
-In particular, the effects of multiple mutations, non-alanine substitutions, and flexible loops are difficult to predict with available tools and protocols.
-We present here a new method to interrogate affinity differences resulting from mutations in a host-virus protein--protein interface.
-Our method is based on extensive non-equilibrium all-atom simulations: We computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1) and estimate affinity using the maximum applied force during a pulling simulation and the area under the force-versus-distance curve.
-We find that these quantities can provide novel biophysical insight into the GP1/hTfR1 interaction.
-First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes.
-Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that one of them may not be important for tight binding.
-Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-resistant hTfR1 mutants.
-Finally, our method provides an elegant framework to compare the effects of multiple mutations, individually and jointly, on protein--protein interactions.
+In many biological applications, we would like to be able to computationally predict mutational effects on affinity in protein-­protein interactions.
+However, many commonly used methods to predict these effects perform poorly in important test cases.
+In particular, the effects of multiple mutations, non­alanine substitutions, and flexible loops are difficult to predict with available tools and protocols.
+We present here an existing method applied in a novel way to a new test case; we interrogate affinity differences resulting from mutations in a host-­virus protein-­protein interface.
+We use steered molecular dynamics (SMD) to computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1).
+We then approximate affinity using the maximum applied force of separation and the area under the force-­versus-­distance curve.
+We find, even without the rigor and planning required for free energy calculations, that these quantities can provide novel biophysical insight into the GP1/hTfR1 interaction.
+First, with no prior knowledge of the system we can differentiate among wild type and mutant complexes.
+Moreover, we show that this simple SMD scheme correlates well with relative free energy differences computed via free energy perturbation.
+Second, although the static co-­crystal structure shows two large hydrogen-­bonding networks in the GP1/hTfR1 interface, our simulations indicate that one of them may not be important for tight binding.
+Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-­resistant hTfR1 mutants.
+Finally, our approach provides a framework to compare the effects of multiple mutations, individually and jointly, on protein­ protein interactions.
 \end{abstract}
+
+\newpage
 
@@
 Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure.
-For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}.
-However, such methods calculate the relative contribution to overall energy for each amino acid in a protein.
-They do not probe the interface directly, but rather calculate the total energy with each substitution \citep{Grant2011,Kortemme2004}.
+For example, end point methods can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}.
+However, such methods can suffer from various simplifying assumptions.
+For example, they generally use an implicit solvent approximation and assume the end state difference with minimal structural rearrangement is sufficient to discriminate energetic differences \citep{Grant2011,Kortemme2004}.
 Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}.
@@
 As such, these methods are poorly suited for most host-virus protein--protein systems.
-By contrast, first principles methods offer the chance to forgo training.
-However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional model to another with no intermediate steps \citep{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
-As a result, the accuracy of the measurement suffers as the two proteins become increasingly dissimilar; therefore, these techniques are largely limited to point mutant comparisons.
+By contrast, first principles methods can forgo training, but currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a transitional model (where one state may be wild-type and the other may be a mutant) to make rigorous free energy calculations  \citep{Gilson1997,Lu2004,Chodera2011,Gumbart2013}.
+While these may be considered two of the gold standard techniques for calculating affinity differences, there are a huge number of theoretical and technical complexities that must all be properly managed to ensure a converged solution \citep{Gumbart2013b}.
+Such considerations quickly come to dominate the protocol, and the necessary book keeping introduces the possibility of human error \citep{Gumbart2013b}.
+Moreover, as the two ending states look ever more dissimilar the chances of convergence fall rapidly.
+To ensure convergence, these techniques are typically limited to small differences (such as point mutant comparisons) with a few, very impressive exceptions \citep{Wang2006,Gumbart2013,Gumbart2013b}.
+For most investigators, larger differences quickly become intractable as the number of intermediate steps required to compute a converged solution grows or the complexity of adding restraining potentials and computing approximations expands \citep{Wang2006,Gumbart2013,Gumbart2013b}.
 
-Here we propose a new method to probe mutational effects in protein--protein interactions directly based on first principles simulation, by pulling two proteins apart that are initially bound in a complex.
+Here we propose that much of these complexities can be avoided if all we are interested in is a relative comparison of the effects of different mutations on protein-protein interactions, rather than measuring an absolute or relative binding affinity with experimentally realistic units.
 We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place.
-Then, we measure the force required for dissociation \citep{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}.
-As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
+Then, we measure the force required for dissociation \citep{Lu1999,Is2001A,Is2001B,Park2004,Gumbart2012,Minoetal2013}.
+Although such biasing techniques are commonly used in protein-ligand binding problems, they are less commonly applied to protein--protein interactions, and almost never to mutational analysis in a protein--protein system.
+This is largely the result of free energy convergence difficulties and computational limitations \citep{Cuendet2008,Cuendet2011}.
+Using a proxy for relative binding affinity rather than caluclating absolute affinities can solve these problems.
+Here, as proxies, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).
+For comparison, we also calculate relative free energy differences using the traditional dual topology FEP paradigm, and we show that the two approaches yield congruent results.
 
-We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}.
-Machupo is an ambisense RNA virus of the arenavirus family \citep{Charrel2003}.
+We used SMD and FEP to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}.
+Machupo virus is an ambisense RNA virus of the arenavirus family \citep{Charrel2003}.
 Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}.
@@
 Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}.
-The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain.
+The co-crystal structure shows that the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid $\beta$-pleated sheet domain.
 For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
@@
 We applied our computational method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes.
+Importantly, the affinity changes predicted using only max force or AUC show a strong correlation with rigorous relative free energy differences computed by FEP.
 At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity may not be as simple as the static structure suggests.
@@
 
+For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDB-ID: 3KAS) \citep{Abraham2010}.
+The apical domain of hTfR1 interacts directly with GP1 while the other two domains are closer to the cell membrane and have essentially no interaction with GP1.
+The biophysical independence of the apical domain allowed us to isolate it without significantly affecting the GP1/hTfR1 interaction.
+
 We used the protein visualization software PyMOL \citep{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1.
 No residues were removed from the viral protein.
-We also removed 10 sugar residues from the native structure.
+Figure~\ref{fig:protein_comparison} shows a model of the initial structure and that of the pared structure.
+Although GP1 has several glycosylatable residues, we opted to use the de-glycosylated protein for this study.
+The complexity of correctly parameterizing diverse sugar moieties is outside of the scope of this paper.
+Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \citep{Abraham2010}.
+In total we removed 10 sugars from the crystal structure for this study.
 
@@
 In addition to the modeled system, for equilibration we generated a configuration file that fixed the $\alpha$-carbon backbone.
-This was accomplished by setting $\beta = 1$.
-Further, we generated a configuration file with fixed $\alpha$-carbon atoms at residues 41-92 (numbered linearly, in this case, starting at 1 for the first amino acid as was required for VMD) in the hTfR1.
+This was accomplished by setting the B-factor column to 1 for the fixed atoms and to zero for all other atoms.
+Further, we generated a configuration file with fixed $\alpha$-carbon atoms at residues 41-92 (numbered linearly, in this case, starting at 1 for the first amino acid as was required for NAMD) in the hTfR1.
 The second file was used to affix a harmonic restraint, thus preventing any unfolding due to system reduction.
-More importantly, the harmonic restraint allowed the protein complex to equilibrate while preventing any drift from its predefined orientation.
+More importantly, the harmonic restraint allowed the protein complex to equilibrate while preventing any drift from its predefined position; the restraint did not constrain the structure of each protein, or the relative position or orientation of the two proteins to each other.
 Finally, we calculated the system center and dimensions for use in molecular dynamics settings.
+The exact NAMD configuration files are available on github (https://github.com/clauswilke/MACV\_SMD).
 
@@
 The initial system temperature was set to 310K.
-Several typical MD settings were used including switching and cutoff.
-In addition, we used a 2 femtosecond time step with rigid bonds (the exact configuration is provided in the supplement).
+Several typical MD settings were used including switching and cutoff distances (see provided configuration files).
+In addition, we used a 2 femtosecond time step with rigid bonds.
 We used periodic boundary conditions with the particle mesh ewald (PME) method of computing full system electrostatics outside of the explicit box.
-Furthermore, we used a group pressure cell, flexible box, and langevin piston during equilibration.
-Finally, a harmonic restraint (called harmonic constraint in VMD) was set as stated previously.
+Furthermore, we used a group pressure cell, flexible box, langevin barostat, and lavegin thermostat during equilibration.
+A harmonic restraint (called harmonic constraint in VMD) was set as stated previously.
 
-To start the simulation, the langevin piston was switched off and the system was minimized for 1000 steps.
+To start the simulation, the barostat was switched off and the system was minimized for 1000 steps.
 Next, the fixed backbone was released, and the system was minimized for an additional 1000 time steps.
 Subsequently, the system was released into all-atom molecular dynamics for 3000 steps.
-Finally, the langevin piston was turned on and the system was simulated for 2 ns (1,000,000 steps) of chemical time.
-For each mutant, twenty independent replicates were run with an identical protocol.
+Finally, the langevin barostat was turned on and the system was simulated for 2 ns (1,000,000 steps) of chemical time.
+For each mutant, twenty independent equilibration replicates were run with an identical protocol.
 
@@
 
+We used the final state from each equilibrated system to restart another MD simulation.
+Our steering protocol is fundamentally similar to \citet{Cuendet2008} with slightly different parameter choices.
+Perhaps the one significant difference lies in our choosing to not use a thermostat or barostat.
+We can make this choice because we are not trying to calculate the binding free energy by any physically rigorous approach (the Jarzynski inequality being one example).
 Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1.
-Also, the $\alpha$-carbon of all residues (163-318 in linear numbering) in the GP1 were given an occupancy of 1, indicating they had an applied force during the simulation.
+These residues were selected as they are far from the binding interface and sufficiently distributed to prevent any orientational motion of the receptor relative to the viral spike protein.
+The center of mass of the $\alpha$-carbons of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation.
+The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms but rather uses the selection to compute an initial center of mass.
+Then, during the steering run, the single center of mass point is pulled with the parameters described below.
+We used the same force field parameters (exclude, cutoff, switching, etc.), the same integrator parameters (time step, rigidbonds on, all molecular being wrapped, etc.), and the same particle mesh ewald parameters as in equilibration.
+Periodic boundary conditions were incorporated as part of the system (as is the convention in NAMD restart) and PME was again used to approximate full system electrostatics.
 
-In every way possible the subsequent steering simulation had the same parameters as the equilibration simulation.
-Thus, the steered molecular dynamics (SMD) \citep{Is2001A,Is2001B} simulation was a typical restart with slightly different settings.
-Periodic boundary conditions were incorporated as part of the system and PME was again used to approximate full system electrostatics.
-In addition, we first tuned the pulling velocity for practical time considerations.
-Subsequently, the spring constant was adjusted to ensure a sufficient signal-to-noise ratio with the given velocity.
-As a result, the spring constant was set to 5~kcal/mol/\AA$^2$, velocity set to 1~\AA/ns, and direction down the positive $z$-axis.
+We ran test simulations at several force constants and visually inspected the results.
+A force constant of 5~kcal/mol/\AA$^2$ was chosen due to its relatively low signal-to-noise ratio.
+This constant is slightly lower than the more common 7~kcal/mol/\AA$^2$ found in several recent studies; that value is commonly selected primarily because it is the force constant found in the SMD tutorial available through the NAMD  developers.
+Moreover, the force constant could very likely be set to a range of nearby values with little loss in predictive power.
 
-Finally, SMD was run for 15 ns (7,500,000 time steps) of chemical time.
+In SMD experiments the pulling velocity should be as low as possible for the available computational time \citep{Cuendet2008,Cuendet2011}.
+We choose a velocity of 0.000001~\AA/fs = 1~\AA/ns, and direction down the positive $z$-axis.
+One could use faster pulling if the computing time must be reduced, but slower than necessary pulling speeds are not typically considered problematic.
+
+SMD was run for 15 ns (7,500,000 time steps) of chemical time.
 For each simulation, we randomly selected one of the equilibration runs for restart.
@@
 
+\subsection{Free Energy Perturbation}
+
+Briefly, we used the traditional dual topology approach to FEP \citep{Gao2006,Pearlman1994}.
+This involves a thermodynamic cycle where a set of atoms are progressively decoupled from the environment while another set of atoms are progressively coupled.
+To compute the relative free energy difference requires knowing the free energy change when the transformation is carried out for the bound complex and the individual protein.
+Then, one can compute the relative free energy difference between a WT and mutant complex by taking the difference between the energy required to decouple/couple the atoms in solution from the energy required to decouple/couple the atoms in the bound complex \citep{Gao2006,Pearlman1994}.
+
+Again, the NAMD configuration file is made available via github (https://github.com/clauswilke/MACV\_SMD).
+We used a similar configuration to that in equilibration.
+One significant difference was to make a cubic water box with a side length equal to the long axis of the complex plus a 10~\AA\  buffer on either side, and simply restrict center of mass motion with the NAMD setting.
+This was done to avoid affecting the system energy while calculating free energy differences.
+
+The transition protocol for bound and free protein systems were identical.
+They started with 1000 steps of minimization and 250,000 steps of equilibration in the starting state for the forward and reverse directions.
+Phase transitions were carried out in steps of $\lambda$=0.05.
+Each transition was carried out for 250,000 steps.
+The first 100,000 steps after phase transition were reserved for equilibration and the final 150,000 steps were used for data collection.
+
+The VMD mutator tool was used to generate the necessary topology file and the parseFEP tool \citep{Liu2012} in VMD was used for subsequent analysis.
+We used it to perform error analysis and compute the Bennett acceptance ratio as the maximum likelihood free energy difference of the two states under consideration.
+Though the larger transitions presented difficulty in a small number of windows, forward and reverse hysteresis was generally in good agreement for all complexes.
+The double mutants were performed by first doing the Y211A mutation followed by the other of the two mutants.
+Then, the $\Delta G$'s were simply added together to get the total energetic difference.
+
 \subsection{Post-processing}
@@
 The starting center-of-mass distance was set to zero and the distance was re-computed at each time step relative to the starting distance.
-The force was output directly to the standard log file.
 
@@
 Each of the 50 independent trajectories per mutant produced a fairly noisy force curve.
-To reduce the noise, we averaged each point with a 50 point interpolation window.
-The window ensured that the averaged maximum applied force was never cut off near the beginning of the simulation due to lost time steps (caused by averaging).
+The force curves for each mutant were smoothed over all replicates by using the smooth.spline() and predict() functions in R with default settings.
 The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC).
-We used the pairwise.t.test function with the FDR p-value correction to test for significance.
+We tested for signifiant differences in maximum force or AUC by carrying out t tests for all pairwise combinations (each mutant compared to each other mutant), using the pairwise.t.test() function in R.
+We adjusted p values to correct for multiple testing using the False-Discovery-Rate (FDR) method \citep{BenjaminiHochberg1995}.
 The ggplot \citep{ggplot} package was used to generate most of the figures.
+
+Analysis scripts and final data (except MD trajectories) are available on the github repository accompanying this publication (https://github.com/clauswilke/MACV\_SMD).
 
@@
 There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV.
-Virtually all of the important sites have been mapped by \textit{in vitro} flow-cytometry based entry assays.
+Many of the important sites have been mapped by \textit{in vitro} flow-cytometry based entry assays.
 The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals).
-In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \citep{Grant2011,Kortemme2004} only marginally useful.
+In addition, much of the binding domain on hTfR1 is on a loop that is flexible prior to viral binding, but organizes to become a strand of a $\beta$-sheet on binding.
+As a result,  many other computational techniques \citep{Grant2011,Kortemme2004} are only marginally useful.
 The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.
 
-For our experiments, we used the experimentally determined GP1/hTfR1 structure (PDBID: 3KAS) \citep{Abraham2010}.
-The apical domain of hTfR1 interacts directly with GP1 while the other two domains are closer to the cell membrane and have essentially no interaction with GP1.
-The biophysical independence of the apical domain allowed us to isolate it without significantly affecting the GP1/hTfR1 interaction.
-Figure~\ref{fig:protein_comparison} shows a model of the initial structure and that of the pared structure.
-Although GP1 has several glycosylatable residues, we opted to use the de-glycosylated protein for this study.
-The complexity of correctly parameterizing diverse sugar moieties is outside of the scope of this paper.
-Furthermore, although it is known that GP1 is glycosylated, and some of those sugars contact hTfR1, the sugars in the available PDB structure are not physiological for mammals \citep{Abraham2010}.
-
 In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}).
-Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
+All of the mutations are within 5~\AA\ of the protein--protein interface.
+Mutations in hTfR1 at site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}.
 Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1.
@@
 
-\subsection{Steered molecular dynamics simulations}
+Despite the fact that viral binding occurs at the site of a flexible loop in the free hTfR structure, our data shows after binding the strand is extremely rigid.
+In the bound conformation, only two sites of the loop have root mean squared fluctuation (RMSF) values in the top half of all receptor sites during equilibration (Figure~\ref{fig:rmsf}), and those are almost completely exposed to solvent.
+This is unsurprising considering the high degree of burial that occurs as a result of viral binding.
+Computing the root mean squared deviation (RMSD) of the entire structure over the trajectory shows that none of the mutations are so deleterious as to cause rapid unbinding.
+In fact, the RMSD over trajectory looks highly invariant across mutants (Figure~\ref{fig:rmsd}).
+In the unbound state, calculated near the end of the SMD trajectory, all of the residues in the WT receptor interfacial strand are in the top half of RMSF over all receptor sites (Figure~\ref{fig:rmsf_comparison}).
+Thus, if sufficient simulation time is not dedicated to allowing this unfolding process, standard free energy techniques may miss the energetic contributions that result from ordering the flexible loop in the hTfR apical domain.
 
-We proceeded to analyze this system using molecular dynamics.
-To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions.
-First, we generated a physically realistic simulation world around the protein complex.
-The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions.
-Next, we carried out successive rounds of energy minimization with a fixed $\alpha$-carbon backbone, followed by heating, and backbone release.
-Last, we performed a short round of all-atom molecular dynamics with minimal restraints \citep{Cuendet2008}.
-We used the final state from each equilibrated system to restart another MD simulation.
-For SMD, at the centroid of several atoms a force was applied to a single point in GP1.
-By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
-A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}.
-The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging.
-The latter approach was used for statistical analysis.
-In addition, to minimize the computational time in unproductive molecular dynamics, we shortened the simulation time such that the interaction just barely dropped to zero near the end of each run.
-As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants.
+\subsection{Molecular dynamics simulations}
+
+We analyzed the GP1/hTfR1 system using two molecular dynamics techniques.
+First, by carrying out SMD using a known force constant and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}.
+A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}, and individual images of all averaged force curves are available in the associated github repository, in folder figures/force\_curves.
+As seen in Figure~\ref{fig:force_curve}, the dissociation distance was relatively consistent among mutants.
 The supplementary movie visually illustrates the separation distance between peptide domains.
-The quantities maximum applied force and AUC were derived from the force versus distances curves.
+The quantities maximum applied force and AUC were derived from the force-versus-distances curves.
 Their summary statistics are reported in Table~\ref{tab:summary_statistics}.
@@
 Before systematically applying SMD to the GP1/hTfR1 interaction, we needed to ensure the method was sufficiently sensitive to distinguish between relatively minor point mutations.
-While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape.
+While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, it is rarely used to parse small energy differences in a protein--protein interaction energy landscape.
 For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.
+
+We proceeded to compare our SMD results to that of the standard dual topology FEP approach to calculate relative free energy differences.
+The correlation between the energetically rigorous FEP and our statistical approach is high.
+For all 11 complexes tested, the correlation between max force and FEP was $r=-0.795$ at $p=0.0034$ (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was $r=-0.593$ at $p=0.055$.
+Because of the strong correlation, we refer exclusively to the SMD results for the remainder of this work, focusing primarily on max force.
 
@@
 
-By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application.
-Traditionally this technique has been either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
+Traditionally SMD has been either applied to compute equilibrium free energies via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}.
 Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}.
 Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions.
-In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application.
-Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation.
-Thus, SMD may open avenues for subsequent experimental work.
+Via SMD, one can separate mutations whose likely effect is altered binding affinity with simple statistics like maximum force of separation.
+Thus, SMD may open avenues for subsequent experimental work in some situations where FEP may be prohibitively difficult.
 
-Our findings may rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}.
+Our findings rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}.
 First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding.
@@
 
-We would like to emphasize here that we cannot expect perfect agreement between our simulations and the available experimental data.
+We would like to emphasize here that we cannot expect perfect agreement between our simulations and the available experimental data, but the correspondence to a well established free energy method bolsters our conclusions.
 While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method.
-First, it is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method.
+First, it is possible that some mutants display measurable phenotypic effects in experiments yet appear identical in simulation.
 More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph).
-Second, our SMD method is fundamentally limited by the accuracy of our starting structure.
+Second, the SMD method is fundamentally limited by the accuracy of our starting structure.
 Third, the available experimental data for the GP1/hTfR1 system were generally obtained from entry assays or whole-cell binding assays rather than molecular binding assays.
@@
 
-For future studies, although our approach offers the simplicity of not requiring prior knowledge about a system of interest (other than a bound model), all atom, unrestrained SMD is likely not the best approach for most relative affinity calculations.
+For future studies, although our approach offers the simplicity of not requiring prior knowledge about a system of interest (other than a bound model), at this point SMD may not the best approach for many relative affinity calculations.
 To ensure one's results are independent of the dissociation path one selects would require computing the work of separation for all likely paths.
 Such an approach eventually requires using the Jarzynski inequality \citep{Jar1997} to establish a lower limit for binding energy and would quickly become computationally inefficient for evaluating a large number of mutations in most systems.
-By contrast, another all atom approach that makes calculations via short simulations of spatially restrained complexes has proven capable of generating relatively accurate binding affinities with less compute time than is required from our steering strategy \citep{Gumbart2013b}.
+However, considering the strong correlation between FEP and SMD in this system, it may not be important to ensure one's results are path independent for relative affinity calculations, as long as the same path is used for all complexes.
 
-\subsection{Notes}
-
-The authors declare no competing financial interest.
+More importantly, with no \textit{a priori} knowledge of the appropriate number of equilibration samples, the best duration of equilibration, the appropriate number of pulling runs, or the best pulling speed means the computational expense in our SMD protocol may not be commensurate with the information provided.
+For example, another all atom approach that makes calculations via short simulations of spatially restrained complexes has proven capable of generating relatively accurate binding affinities with less compute time than is required from our steering strategy \citep{Gumbart2013,Gumbart2013b}.
+That being said, there is no reason to believe this SMD approach to mutagenic studies could not be optimized to reduce computational expense.
+Further analysis will be needed to understand the lower limits of resources required for accurate predictions.
 
 \section{Acknowledgements}
-This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock \\ (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu).
-The authors acknowledge Bryan Sutton for opening access to the Hrothgar cluster.
+This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu).
+We would like to thank Bryan Sutton for opening access to the Hrothgar cluster and the reviewers Ilan Samish and Matteo Masetti for their helpful comments on this work.
 
-This work was supported by the Defense Threat Reduction Agency (HDTRA1-12-C-0007) to A.D.E., S.L.S., and C.O.W., the National Science Foundation (MCB-0943383) and the Welch Foundation (F-1654) to A.D.E., the National Institutes of Health (R01-GM088344) to C.O.W., and the National Institutes of Health (R01-GM093086) to S.L.S.
-
-\bibliographystyle{MBE}
-\bibliography{paper}
+\bibliographystyle{peerj}
+\bibliography{full_paper}
 
@@
 (B) The reduced structure used in SMD simulations.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/rmsfs.pdf}}
+\caption{\label{fig:rmsf}\title{RMSF values during equilibration.} The RMSF values for every site in the bound complex computed during the equilibration phase of the protocol.
+Each color represents the average over 20 trajectories of a single mutant.
+Indices 17-25 are the hTfR flexible loop.
+The plot shows the flexibility of each site is essentially independent of mutation, and two sites (indices 17 and 18) above 0.72~\AA\ are a part of the flexible loop in the free receptor.
+However, these two residues are not actually found in the protein--protein interface, but rather are almost completely solvent exposed with the virus bound.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/rmsds.pdf}}
+\caption{\label{fig:rmsd}\title{RMSD values during equilibration.} The RMSD values over the time of the trajectory computed during the equilibration phase of the protocol.
+Each color represents the average over 20 trajectories of a single mutant.
+The plot shows none of the mutants causes immediate unbinding of the protein--protein complex.
+In addition, the universal upward trend near the end of the equilibration trajectories may indicate the crystal is more tightly packed than would normally occur in solution.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/rmsf_comparison.pdf}}
+\caption{\label{fig:rmsf_comparison}\title{RMSF values of WT hTfR in equilibration and SMD.} The RMSF values for every site in the WT receptor were computed during the equilibration phase and during final 50 frames of the SMD trajectories.
+The black line was computed over equilibration and the red line during SMD.
+The plot shows the solution mobility of the hTfR flexible loop increases more than the average during the unbinding process.}
 \end{figure}
@@
 The large non-overlapping areas indicate a large and statistically significant difference in these three complexes.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/fep_mutants.pdf}}
+\caption{\label{fig:fep}\title{Max force versus free energy perturbation.} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex.
+The WT complex for FEP was simply set to 0.0.
+The correlation between the two is $r=-0.795$ with $p=0.0034$.}
 \end{figure}
@@
 \caption{\label{tab:summary_statistics}Summary statistics for each mutation tested.
-$\mu_\text{MAF}$ is the mean and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations.
-$\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations.}
+$\mu_\text{MAF}$ is the mean in piconewtons and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations.
+$\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations.
+$\Delta G$ is the free energy difference in kcal/mol calculated via FEP by the dual topology paradigm.}
 \begin{center}
-  \resizebox{11cm}{!} {
-    \begin{tabular}{l c c c c}
+  \resizebox{15cm}{!} {
+    \begin{tabular}{l c c c c c}
     \hline
-      Mutation &      $\mu_\text{MAF}$ & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ \\ \hline
-            WT &                772.00 &              154.13 &         198794.7 &            66765.64 \\
-         N348A &                799.33 &              154.36 &         202247.3 &            65175.78 \\
-         N348K &                731.36 &              151.01 &         179087.5 &            55799.49 \\
-         N348W &                732.13 &              131.21 &         192043.2 &            66916.36 \\
-         Y211A &                684.03 &              126.69 &         153517.7 &            40889.12 \\ 
-         Y211D &                875.44 &              116.46 &         234909.3 &            50494.11 \\ 
-         Y211T &                861.17 &              158.48 &         231078.7 &            97569.29 \\
-        vR111A &                754.39 &              103.25 &         185896.8 &            46408.74 \\
-   N348A/Y211A &                762.96 &              144.81 &         173059.2 &            59898.10 \\
-   N348W/Y211A &                617.21 &              163.33 &         147705.6 &            48548.61 \\
-  vR111A/Y211A &                780.71 &              135.65 &         189919.1 &            48868.16 \\
+      Mutation & $\mu_\text{MAF}$ (pN) & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ & $\Delta G$ (kcal/mol)\\ \hline
+            WT &                734.4856 &              131.6513 &         145460.4 &            60232.26 &  0.000\\
+         N348A &                748.5217 &              137.4864 &         133913.9 &            51078.64 & -2.149\\
+         N348K &                705.0707 &              108.5079 &         141084.4 &            54450.28 & +3.184\\
+         N348W &                697.3642 &              132.6436 &         136886.0 &            53796.44 & +3.033\\
+        vR111A &                713.8081 &              106.7374 &         136103.2 &            52070.85 & +0.466\\ 
+   N348A/Y211A &                703.7027 &              128.5866 &         113464.2 &            57451.62 & +5.203\\ 
+  vR111A/Y211A &                741.0642 &              131.6287 &         130070.6 &            47665.56 & -2.440\\
+         Y211D &                825.2586 &              115.4343 &         158878.7 &            63039.08 & +2.760\\
+         Y211T &                806.8593 &              136.5648 &         167110.7 &            78849.29 & +0.875\\
+         Y211A &                654.1138 &              108.5343 &         108090.0 &            43661.09 & +2.526\\
+   N348W/Y211A &                594.9044 &              134.8233 &         108984.2 &            45451.00 & +8.206\\
     \hline

### 2014-01-20 full_paper.tex
Committing some text changes and figures to another structure composite.

@@
 At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity may not be as simple as the static structure suggests.
-For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1.
+For example, the static structure shows a hydrogen-bonding network connected to site N348 in hTfR1.
 According to our simulations, this network may not affect binding affinity directly.
@@
 The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}.
-In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1.
-In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}.
-Finally, an alanine mutation at site 111 in GP1 (mutation vArg111Ala) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}.
+In a manner similar to site 211, N348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1.
+In particular, N348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}.
+Finally, an alanine mutation at site 111 in GP1 (mutation vR111A) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}.
 For notation purposes, the viral site is always referred to with a preceding `v'.
@@
 
-We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:MAF_pairwise}), while the other two did not (Table~\ref{tab:MAF_pairwise}).
+We found that relative to WT, one alanine mutation (Y211A) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:MAF_pairwise}), while the other two did not (Table~\ref{tab:MAF_pairwise}).
 When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:MAF_pairwise} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}).
@@
 
-Although Tyr211Ala appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism.
+Although Y211A appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism.
 Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for change in binding affinity.
 In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding.
-We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}.
+We chose Y211D and Y211T, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}.
 Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity (Figures~\ref{fig:sensitivity} and~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).
 
-We also simulated several point mutations at Asn348 in the hTfR1.
+We also simulated several point mutations at N348 in the hTfR1.
 As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}).
-In addition, neither the Asn348Lys nor the Asn348Trp mutation showed a significant difference from WT.
+In addition, neither the N348Lys nor the N348W mutation showed a significant difference from WT.
 For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT (See Table~\ref{tab:summary_statistics}).
-On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder.
-Moreover, Asn348Trp showed nearly identical results to Asn348Lys.
-The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:MAF_pairwise}).
-To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding.
+On the other hand, there was a detectable difference between N348A and N348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with N348Lys being a weaker binder.
+Moreover, N348W showed nearly identical results to N348Lys.
+The mutations to large amino acids (N348W and N348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (N348A and N348W) produced significantly different affinity changes (Table~\ref{tab:MAF_pairwise}).
+To check the consistency of our results, we hypothesized that the combination of Y211A and N348W, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding.
 As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $, Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study.
-Further, according to maximum applied force (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}).
-We suspect that the effect of Asn348Trp alone is near the limit of detection using our method.
-A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.
+Further, according to maximum applied force (but not AUC), the combination of Y211A and N348W also showed significantly weaker binding than Y211A by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}).
+We suspect that the effect of N348W alone is near the limit of detection using our method.
+A larger number of replicates would possibly have resolved affinity differences between N348W and WT or other mutants more consistently.
 
-Last, we further analyzed a single mutation in GP1, vArg111Ala.
+Last, we further analyzed a single mutation in GP1, vR111A.
 As mentioned previously, in our simulations this mutant showed no significant change in either maximum applied force or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}).
-This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity.
-To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala.
-This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:MAF_pairwise}).
+This result was somewhat surprising, since Y211A, presumably disrupting the same hydrogen-bonding network as vR111A, displayed a significant reduction in affinity.
+To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vR111A/Y211A.
+This double mutant showed affinity indistinguishable from WT and significantly higher than Y211A alone  (Table~\ref{tab:MAF_pairwise}).
 This result shows that the two sites do indeed interact, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction could lead to comparable binding affinity.
@@
 We systematically tested several point mutations to understand their contribution to the binding interaction.
-In the case of Asn348Lys, we have shown that the static structure provides little insight into why this mutation causes loss-of-infectivity \textit{in vivo}.
-While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, change in binding at that site may actually be caused by size and charge restriction.
+In the case of N348Lys, we have shown that the static structure provides little insight into why this mutation causes loss-of-infectivity \textit{in vivo}.
+While N348 appears to be involved in a hydrogen-bonding network in the static structure, change in binding at that site may actually be caused by size and charge restriction.
 We also found that a negatively polar residue at site 211 in hTfR1 seem critical for a tight binding interaction.
-Any non-polar mutation at Tyr211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection.
+Any non-polar mutation at Y211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection.
 
@@
 The ability to hydrogen bond appears to be insignificant.
-This can be inferred from the fact that Tyr211Ala paired with large (Trp) and positively charged (Lys) substitutions at position 348 results in a larger than expected synergistic difference.
-That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually.
-Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
-Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.
+This can be inferred from the fact that Y211A paired with large (W) and positively charged (Lys) substitutions at position 348 results in a larger than expected synergistic difference.
+That is, the double mutant Y211A/N348W caused a much larger decrease in binding than we expected from either mutation individually.
+Third, the GP1 mutation vR111A causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations.
+Although Y211A was the most disruptive single mutant we tested, vR111A in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Y211A alone.
 
@@
 \bibliographystyle{peerj}
-\bibliography{full_paper}
+\bibliography{paper}
 

### 2014-01-20 full_paper.tex
Final changes.

@@
 Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-­resistant hTfR1 mutants.
-Finally, our approach provides a framework to compare the effects of multiple mutations, individually and jointly, on protein­ protein interactions.
+Finally, our approach provides a framework to compare the effects of multiple mutations, individually and jointly, on protein­-protein interactions.
 \end{abstract}
@@
 The co-crystal structure shows that the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid $\beta$-pleated sheet domain.
-For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
+For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010} (Figure~\ref{fig:networks}).
 Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.
@@
 \begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/Networks.pdf}}
+\caption{\label{fig:networks}\title{The two hydrogen bonding networks.} GP1 is shown in blue and hTfR1 is shown in green.
+(A) The first network including Y211 and R111 is shown in white, and the second network containing N348 is shown in pink.
+(B) Near view of the first network with contacts in yellow.
+(C) Near view of the second network with contacts in yellow.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
 \centerline{\includegraphics[width=6in]{figures/protein_comparison.pdf}}

### 2014-01-20 full_paper.tex
Fixed figure order.

@@
 The co-crystal structure shows that the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid $\beta$-pleated sheet domain.
-For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010} (Figure~\ref{fig:networks}).
+For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}.
 Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.
@@
 \subsection{The GP1/hTfR1 system}
-The GP1/hTfR1 interface marks a particularly important and useful test system.
+The GP1/hTfR1 interface (Figure~\ref{fig:networks}) marks a particularly important and useful test system.
 There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV.
@@
 
-In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}).
+In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants}).
 All of the mutations are within 5~\AA\ of the protein--protein interface.
@@
 \subsection{Comparative analysis of the GP1/hTfR1 interface}
-Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction.
+Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface (Figure~\ref{fig:networks}), it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction.
 One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function.
@@
 \begin{figure}[p]
+\centerline{\includegraphics[width=6in]{figures/protein_comparison.pdf}}
+\caption{\label{fig:protein_comparison}\title{The GP1/hTfR1 complex.} GP1 is shown in blue and hTfR1 is shown in green.
+(A) The full, de-glycosylated GP1/hTfR1 co-crystal structure.
+(B) The reduced structure used in SMD simulations.}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
 \centerline{\includegraphics[width=6in]{figures/Networks.pdf}}
@@
 (C) Near view of the second network with contacts in yellow.}
-\end{figure}
-
-\clearpage
-\begin{figure}[p]
-\centerline{\includegraphics[width=6in]{figures/protein_comparison.pdf}}
-\caption{\label{fig:protein_comparison}\title{The GP1/hTfR1 complex.} GP1 is shown in blue and hTfR1 is shown in green.
-(A) The full, de-glycosylated GP1/hTfR1 co-crystal structure.
-(B) The reduced structure used in SMD simulations.}
 \end{figure}
@@
 \begin{figure}[p]
-\centerline{\includegraphics[width=6in]{figures/sensitivity_histogram.pdf}}
-\caption{\label{fig:sensitivity}\title{Distribution of interpolated maximum force for three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left.
-The large non-overlapping areas indicate a large and statistically significant difference in these three complexes.}
+\centerline{\includegraphics[width=6in]{figures/fep_mutants.pdf}}
+\caption{\label{fig:fep}\title{Max force versus free energy perturbation.} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex.
+The WT complex for FEP was simply set to 0.0.
+The correlation between the two is $r=-0.795$ with $p=0.0034$.}
 \end{figure}
@@
 \begin{figure}[p]
-\centerline{\includegraphics[width=6in]{figures/fep_mutants.pdf}}
-\caption{\label{fig:fep}\title{Max force versus free energy perturbation.} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex.
-The WT complex for FEP was simply set to 0.0.
-The correlation between the two is $r=-0.795$ with $p=0.0034$.}
+\centerline{\includegraphics[width=6in]{figures/sensitivity_histogram.pdf}}
+\caption{\label{fig:sensitivity}\title{Distribution of interpolated maximum force for three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left.
+The large non-overlapping areas indicate a large and statistically significant difference in these three complexes.}
 \end{figure}
